### 2014-12-04 intro.tex
plan et introduction

As regulatory phenomena play a crucial role in biological systems, they need to
be studied accurately. Biological Regulatory Networks (BRNs) consist in sets
of either positive or negative mutual effects between the components. With the
purpose of analyzing these systems, they are often modeled as graphs which make
it possible to determine the possible evolutions of all the interacting components of the system. Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently use a new formalism, named the “Process Hitting” (PH) \cite{pauleve2011modelisation}, to model concurrent systems having components with a few qualitative levels. A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system. This particular structure makes the formal analysis of BRNs with hundreds of components tractable. PH is suitable, according to the precision of this information, to model BRNs with different levels of abstraction by capturing the most general dynamics.
The objectives of the work presented in this paper are the following.

Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points \cite{wuensche1998genomic}).
We perform an exhaustive search of the possible states, combination processes, one process from each sort and then check if it is a fixed point.

The second phase of our work consists in computing the dynamics. It consists in determining from a known initial state the possible next states of the PH model. Finally we verify if we can reach a specific state of one or several component (gene or protein). The results are ensured to respect the PH dynamics.

Our contribution is from the results that allowed to determine the stable states, we propose to evaluate the benefits of the Answer Set Programming (ASP) \cite{baral2003knowledge} to compute them. ASP has been proven efficient to tackle models with a large number of components and parameters. Our aim here is to assess its potential w.r.t. the computation of some dynamical properties of the PH model. In this paper, we show that ASP turns out to be effective for these enumerative searches which justifies its use. The benefit of our approach is that it makes possible to get the minimal paths to reach our goal(s) also we can verify if it is possible after a given number of steps. %Furthermore, we claim that we are able to deal with large biological networks (42 sorts).

### 2014-12-10 ASP.tex
ASP definition

In order to develop the nes methods we have used two main frameworks. The first is a logic programming language: the Answer Set Programming (ASP). Precisely it is a logic programming language on which I wrote my bibliographic research report. The second framework is the Process Hitting (PH) a new formalism for the representation of biological regulatory networks. We will see later in this section more details about these two frameworks.

\subsection{Answer Set Programming }

The ASP is a declarative programming paradigm with known semantics as the semantics of answer sets. This paradigm allows the programmer to specify what the problem to be solved and not how to solve it.
ASP programs are written in AnsProlog* (short for "Answer Set Programming in Logic" with superscript *). These programs are composed of a set of facts and a set of rules from which other facts can be derived. A consistent set of facts that can be derived from a program using the rules is called "answer set" for the program. The sets of possible responses to an AnsProlog* program are calculated with a program called a solver.

\subsubsection{Logic program}
Consider an ASP logic program, as each rule is an ordered pair,\cite{baral2003knowledge} :

\begin{equation} \label{eq1ASP}
head \leftarrow body.
\end{equation}
with $Head$ and $Body$ sets of literals.

This simple language has many advantages that make it one of the most effective in terms of knowledge representation, reasoning and declarative problem solving.

\subsubsection{The syntax and semantic of ASP}
\label{sectionSyntaxeASP}
\begin{enumerate}
\item \textbf{L'alphabet}

\noindent{ According \cite{baral2003knowledge}, the alphabet of the axiom (or just the alphabet) of the framework of the answer set is comprised of seven classes of symbols:
\begin{itemize}
\item Variables,
\item Constants,
\item Symbols of functions,
\item Symbols of predicates,
\item Connectors,	
\item Symbols of punctuation, and	
\item Special symbol $\bot$
\end{itemize}
}
All these classes vary from one alphabet to another, but the sets of $ 5^{th} $ and $ 6^{th} $ classes (Connectors and symbols of punctuation) are defined as follows:\\

- Connectors with \{$\neg$, or, $\leftarrow$, not, ‘,’\} \\
- Symbols of punctuation with \{‘(’, ‘)’, ‘.’\} \\

The other classes remain constant as they meet some informal agreement. In general, variables start with a capital letter and contain letters and numbers(X,Y,..). The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter(f,g,a,b...). Sometimes there is the addition of a supplementary agreement which covers the letters used \cite{baral2003knowledge} : \\

What seems a little fuzzy is the concept of predicate. Indeed the word predicate can be an innovation of the new grammar. Let's consider the following sentence: 
\begin{tabbing}
 \textit{Socrates was an athenian.} 
\end{tabbing}
\textit{Socrate} is the subject and \textit{was an athenian} is the predicat. This sentence can be noted in ASP by :
\begin{tabbing}
 \textit{athenian(Socrate) }$\leftarrow ~ \top$. \hspace{2cm} with $\top$ is the symbol of True.
\end{tabbing}

\item \textbf{The rules}
Comme nous l’avons déjà mentionné, une règle est de la forme:

\begin{equation} 
head ~ \leftarrow ~ body.
\tag{\eqref{eq1ASP}}
\end{equation}
\begin{equation} \label{eq3ASP}
A ~ \leftarrow ~ B^+, ~ \textbf{not} ~ B^-.
\end{equation}

This rule is then equivalent to:

\begin{equation} \label{eq2ASP}
L_{0} \leftarrow L_{1},...,L_{m}, \textbf{not} L_{m+1},..., \textbf{not} L_{n}.
\end{equation}

And this rule states:

$$
\left.
    \begin{array}{ll}
        \text{Si} ~ L_{1},...,L_{m} ~ \text{are ~ \textbf{true}}\\
        \text{et} ~ L_{m+1},...,L_{n} ~ \text{are~ \textbf{false}}
    \end{array}
\right \} \text{Then} ~ L_{0} ~ \text{is } \textbf{true} 
$$
with $L_{i}$ literals and k $\geq$ 0, m $\geq$ k and n $\geq$ m.

\underline{\textbf{Head}} is also called conclusion \\
\underline{\textbf{Body}} is also called premise

Special cases rules \cite{Vladimir, baral2003knowledge} :\\

\begin{itemize}
\item A rule is \textbf{constant} if all litterals are constants (noted with \textbf{ground});
\item A \textbf{fact} or a \textbf{reality} : it is a rule with an empty body. It can be written even without the arrow $\leftarrow$ :\\
\begin{equation} 
L_{0}. ~~ or ~~ L_{0} ~ \leftarrow ~  \top.
 \label{eq5ASP}
\end{equation}

\item A \textbf{constraint}:
This false symbol to the head ($head$) is often eliminated and the constraint will be written generally as:
\begin{equation} 
 ~ \leftarrow ~ L_{1},...,L_{m}, ~ \textbf{not} ~ L_{m+1},..., ~\textbf{not} ~L_{n}.
 \label{eq6ASP}
\end{equation}

We say that a set of literals $X$ violates the constraint (\ref{eq6ASP}) if $\{L_{1},...,L_{m}\} \subseteq X$ and $\{L_{m+1},...,L_{n}\} \nsubseteq X$. If we have a program contain this type of rule \ref{eq6ASP}, then $X$ is an answer set of the program $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$, with $r$ such a constraint \ref{eq6ASP} and $X$ does not violate the constraint \ref{eq6ASP}.

\item Cardinality constraints:\\
It is literals extended in the following form:
\begin{equation} 
 ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u.
 \label{eq7ASP}
\end{equation}

with $m \geq 1$, $l$ an integer and $u$ can be an integer or by default the infinite if it does not exist. $l$ and $u$ are the lower and upper limits of the cardinality of the subsets of $ \{q_{1},~ ... ,~ q_{m}\}$ that satisfy the answer sets.
These literals are constrained ($q_{i}$) may occur in the head and body of the rule. Cardinality constraint is satisfied in an answer set $ X $, if the number of atoms of $ \{q_{1},~ ... ,~ q_{m}\}$ belonging to $X$ is between $l$ and $u$. In other words:

\begin{tabbing}
 $l~ \leq ~ | \{q_{1},... ,~ q_{m}\} ~ \cap ~ X | ~ \geq ~u$
\end{tabbing}

with $ \cap $ symbol of intersection and $ | A | $ is the cardinality of the set $ A $.

\end{itemize}

\end{enumerate}

\subsubsection{Modelisation of a problem with ASP}

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}:
\begin{itemize}
\item it is a simplification of a given system
\item it allows an action on the system  
\end{itemize}

Models offer the possibility to provide a solution to a problem identified as such. This concept model can address another angle issues related to representation process.

\begin{enumerate}
\item\textbf{ Modeling steps}
The modeling process can be regarded as a special form of representation whose operations are detailed in \cite{{baral2003knowledge}}. Logical ASP programs follow the strategy of "generate and test". This strategy includes four steps:

\begin{itemize}
\item Enumerate with facts;
\item Explain with the rules;
\item Generate all the possibilities with cardinalities, and finally;
\item Filter with constraints.
\end{itemize}

\item \textbf{ Resolving an ASP program} \\
From \cite{Vladimir} and \cite{AnsPrologAPE}, the couple Grounder and Solver usually work together: the grounder used to remove variables in order to achieve a constant program and the solver computes all anwser sets for stabilized programs generated by the grounder.

Example of Grounder: \texttt{GRINGO}, \texttt{DLV}, \texttt{LPARSE}

Example of ASP Solver : \texttt{SMODELS}, \texttt{DLV}, \texttt{CMODELS}, \texttt{CLASP}...

The combined use of the Grouder and the Solver specifies major programs in a compact, using rules with schematic variables and other abbreviations. Both systems employ grounding algorithms that work quickly and simplify the program. In developping these new methods we worked with \texttt{CLINGO} which is a combination of grouder \texttt{GRINGO} and solver \texttt{clasp}.

\end{enumerate}

### 2014-12-10 PH.tex
PH definition

\pref{def:PH} introduces the Process Hitting~(PH)~\cite{pauleve2011modelisation}
which allows to model
%A Process Hitting (PH) (\pref{def:PH}) gathers 
a finite number of local levels,
called \emph{processes},
grouped into a finite set of components, called \emph{sorts}.
A process is noted $a_i$, where $a$ is the sort's name,
and $i$ is the process identifier within sort $a$.
At any time, exactly one process of each sort is \emph{active},
and the set of active processes is called a \emph{state}.

The concurrent interactions between processes are defined by a set of \emph{actions}.
Actions describe the replacement of a process by another of the same sort
conditioned by the presence of at most one other process in the current state.
An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
``$a_i$ \emph{hits} $b_j$ to make it bounce to $b_k$'',
where $a_i,b_j,b_k$ are processes of sorts $a$ and $b$,
called respectively \emph{hitter}, \emph{target} and
\emph{bounce} of the action.
We also call a \emph{self-hit} any action whose hitter and target sorts are the same,
that is, of the form: $\PHfrappe{a_i}{a_i}{a_k}$.

\begin{definition}[Process Hitting]\label{def:PH}
  A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
  \begin{itemize}
    \item[--]  $\PHs = \{a,b,\dots\}$ is the finite set of \emph{sorts};
    \item[--]  $\PHl = \prod_{a\in\PHs} \PHl_a$ is the set of states with
      $\PHl_a = \{a_0,\dots,a_{l_a}\}$
      the finite set of \emph{processes} of sort $a\in\Sigma$
      and $l_a$ a positive integer, with $a\neq b\Rightarrow \PHl_a \cap \PHl_b = \emptyset$;
    \item[--]  $\PHa = \{ \PHfrappe{a_i}{b_j}{b_k} \in \PHl_a \times \PHl_b^2 \mid
      (a,b) \in \PHs^2 \wedge b_j\neq b_k \wedge a=b\Rightarrow a_i=b_j\}$
      is the finite set of \emph{actions}.
  \end{itemize}
\end{definition}

\begin{definition}[Action]
\label{def:PhAction}
An action is noted $\PHfrappe{a_i}{b_j}{b_k}$ where $a_i$ is a process
of sort $a$ and $b_j$, $b_k$ two processes of sort $b$. When $a_i = b_j$ , such an action is refered as a self-action and $a_i$
is called a self-hitting process.
\end{definition}

\begin{example*}
\pref{fig:ph} represents a PH $(\PHs,\PHl,\PHa)$ with three sorts
($\PHs = \{a, b, c\}$) and:
$\PHl_a = \{a_0, a_1\}$,
$\PHl_b = \{b_0, b_1, b_2\}$,
$\PHl_c = \{c_0, c_1\}$.
\begin{figure}[ht]
\centering
\begin{tikzpicture}%[font=\scriptsize]
%\path[use as bounding box] (0,-1) rectangle (4,4);

\TSort{(0,0)}{z}{3}{l}
\TSort{(2,4)}{b}{2}{t}
\TSort{(4,1)}{a}{2}{r}
\THit{b_0}{}{z_1}{.east}{z_2}
\THit{b_1}{}{z_0}{.north east}{z_2}
\THit{a_0}{}{b_1}{.south}{b_0}
\THit{a_1}{out=60,in=0,selfhit}{a_1}{.east}{a_0}

\path[bounce,bend right]
\TBounce{z_1}{}{z_2}{.south}
\TBounce{z_0}{bend right=50}{z_2}{.south east}
;
\path[bounce,bend left]
\TBounce{a_1}{}{a_0}{.north}
\TBounce{b_1}{}{b_0}{.south}
;

 \THit{z_0}{}{a_0}{.west}{a_1}

\path[bounce,bend left]
\TBounce{a_0}{}{a_1}{.south}
;
\TState{a_0,b_0,z_1}
\end{tikzpicture}
\caption{\label{fig:ph} 
A PH model example with three sorts: $a$, $b$ and $z$($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2).
Circles represent the processes, boxes represent the sorts, and the actions are drawn by pairs of arrows in solid and dotted lines.The grayed processes stand for a possible initial state.
}
\end{figure}

\end{example*}

\begin{definition}[Next state]
\label{def:NextState}
Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
one of its states. The set of the next possible states for $s$ are computed as follows:
\begin{center}
$next(s) = {s[b_k /b_j ]| \exists (a i , b j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa}$
\end{center}

\end{definition}

\begin{definition}[Stable state or fixed point]
\label{def:FixPoint}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be a state, $s$ is a stable state for $PH$ if and only if $next(s) = \emptyset $.
\end{definition}

\begin{definition}[Reachability]
\label{def:Reachability}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be a state OF $PH$, $s$ is a stable state for $PH$ if and only if $next(s) = \emptyset $. (i.r there is no playable action at the state $s$).
\end{definition}

\begin{definition} [Playable action]
\label{def:playableAction}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$. We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is playable at the state $s$ if and only if $hitter(h)=a_i ~ \in s$ and $target(h)=b_j \in s$ (i.e $s[a]=a_i$ et $s[b]=b_j$ ) \\
The resulting state after playing the action $h$ at $s$ is denoted by $(s \play h)$ or $(s \play h)[b]=b_k$ and $\forall c \in \PHs, ~ c \neq b, (s \play h)[c]=s[c].$
\end{definition}

### 2014-12-11 PH.tex
search of fixed point

@@
 \begin{definition}[Stable state or fixed point]
-\label{def:FixPoint}
+\label{def:FixPoint1}
 Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
@@
 \end{definition}
+\begin{definition}[Lemme: Stable state or fixed point]
+\label{def:FixPoint2}
+Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
+$s \in \PHl$ be a state, $s$ is a stable state for $PH$ there is no playable action at the state $s$: 
+\begin{center}
+$\forall h \in \PHa$, $frappeur(h) ~ \notin s ~ \vee ~ cible(h) \notin s$)
+\end{center}
+\end{definition}
 
-\begin{definition}[Reachability]
-\label{def:Reachability}
-Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
-$s \in \PHl$ be a state OF $PH$, $s$ is a stable state for $PH$ if and only if $next(s) = \emptyset $. (i.r there is no playable action at the state $s$).
+\begin{definition}[Reachability question]
+\label{def:reachability}
+
+ If $\ctx \in \PHl$ is a state and $A \in \Proc$ a set of processes, we note $\mathcal{P}(\ctx, A)$ pour \emph{la question d'accessibilité}:
+ \begin{center}
+  $\exists? \delta \in \Sce(\ctx), A \subseteq \PHget(\ctx \play \delta) $ (i.e $\forall a_i \in A, \PHget{(\ctx \play \delta)}{a}=a_i$).
+ \end{center}
+With $\Sce(\ctx)$ 
+is the set of successively playable actions from the state $\ctx$. 
 \end{definition}

### 2014-12-11 fixPoint.tex
search of fixed point

%Section of Fix point implementation
The study of fixed points (or basins of attraction) provides an important understanding of the different behaviors of a BRN (Biological Regulatory Network)\cite{wuensche1998genomic}.  
The fixed point is a stable state of the BRN in wich it is not possible any more to have new changes.
Let $(\Sigma, L, H)$ be a Process Hitting .
 It has been shown that a state $s \in L$ is a fixed point of the Process Hitting if and only if $s$ is has no next state \cite{pauleve2011modelisation} i.e. there is no playable action at this state. In fact a steady state od a Process Hitting network is a set of processes with exactly one process of each sort as well as every process has no hit with the other selected ones (process with self-hit cannot be apart of a stable state).

\subsection{Process Hitting network traduction }
In order to handle a PH biological network, it was necessary first to present it with ASP. To do this we chose the predicates: $sort$, $process$ and $action$. Below is an example of PH network shown in ASP.

\begin{exemple} [Example PH network with ASP]
If we try to present the network of Figure \ref{fig:expPH} we will have:
\begin{lstlisting}
sort("a"). sort("b"). sort("z"). %\label{ASPsort}
process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
action("a",0,"b",1,0). action("a",1,"a",1,0). action("b",1,"z",0,2). %\label{actions1}
action("b",0,"z",1,2). action("z",0,"a",0,1). %\label{actions2}
\end{lstlisting}
Line \ref{ASPsort} shows every sort of network with the predicate comes out and in parentheses the name of the sort. In line \ref{ASPprocess} there is a list of processes corresponding to each $sort$ for example the sort "$z$" has 3 process numbered from 0 to 2, this numbering is provided by the $2^{nd}$ parameter of the predicate $process("z", 0..2)$. Finally we find all network actions that are defined in lines \ref{actions1} and \ref{actions2}, for example the first action $action("a",0,"b",1,0)$ is an action from $a_0$ to $b_1$ to bound so $b$ from $b_1$ into $b_0$.
\end{exemple}

\subsection{Search of fixed points }
Parler du point fixe et ses interrêts. \\
Decrire la nouvelle méthode qui permettent de determiner les états stables d'un réseau

In fact we have to translate the definition of a stable state in a method developped in ASP. So first we eliminate all processes with self-hit and we save them in the predicate "$shownProcess$":
\begin{lstlisting}
hiddenProcess(A,I) :-  action(A,I,B,J,K), A=B ,process(A,I), process(B,J), process(B,K). %\label{hiddenProcess}
shownProcess(A,I) :-  not hiddenProcess(A,I), process(A,I). %\label{shownProcess}
\end{lstlisting}
Then we have to browse this graph and extract all possible combinations of shown processes by choosing a process from each sort.
\begin{lstlisting}
1 { selectProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
\end{lstlisting}
It now remains to check each combination of processes whether it is a fix point. For this we use a special type of ASP rules: a \textbf{constraint}. The idea of constraints is based on the fact that a solution would be eliminated if it does not satisfy the constraint. For our problem a combination is eliminated if there is an action between two of the selected process:
\begin{lstlisting}
:- hit(A,I,B,J), selectProcess(A, I), selectProcess(B, J), A!= B. %\label{contraintFix}
\end{lstlisting}

\begin{example*}
Considering the last graphical presentation of a precess hitting \pref{fig:ph}, there are 3 sorts $a$, $b$ have 2 levels and $z$ has 3 so we can find $2*2*3 = 12$ states (whatever they can be reached or not). If we verify wether there exist fixed points we deduce that we have only one: $<b\_0, z\_2, a\_0>$. It is clear in \pref{fig:ph} that there is no action between each two processes. Besides our new ASP method proof this and returns also the same answer:
\begin{tabbing}
 \texttt{
 Answer 1 : fixProcess(a, 0), fixProcess(b, 0), fixProcess(z, 2)
 }
\end{tabbing}
\end{example*}

### 2014-12-12 reachability.tex
PH network evolution

%Section: Dynamic network evolution

In this section, we will present firstly how to determine using ASP the possible evolution of a biological network after a finite number of steps. Then we answer to the reachability question; which evolutions that allow the achievement of goals (future active processes) from a known initial state?

\subsection{Future states identification }
From an initial known state, a PH network can evolve into several new states after a few steps. The predicate $"time (0 .. n)"$ sets the number of steps we want to play. For example if the biologist wants to know which states are reachable after 10 stages, it has only to replace $n$ by 10, and he will have $"time (0 .. 10)"$. To compute the future states it is necessary to define an initial state of the network. So for initializing the active processes we add in the ASP script representing the PH network the following rule for each sort. We note that this rules are generated automatically at the time of translating PH network to ASP and by default the level of all active processes is $0$.\\
\begin{lstlisting}
init(activeProcess("a",0)).
\end{lstlisting}
with $a$ is the name of the sort and $0$ the index of the active process.\\
The dynamic of a network is realised thanks to its actions. So to identify the future states we should start by identifying the playable actions for each state. We recall that an action is playable when both precesses: hitter and taget are active (Definition \ref{def:playableAction}). In ASP the predicate $"playableAction(A,I,B,J,K,T)"$ it is true when the processes $(A,I)$ and $(B,J)$ are active at the step $T$.
Our approch is to study the asynchronous dynamic so only one action is playable at the same step ($T$). That's why between two successive states there will be one change of one sort. We present this by the predicate $"activeFromTo(B,J,K,T)"$. It means that in the sort $B$ the active process change from index $J$ to $K$ at time step $T$.
At the same step we can find many possible changes , that's why the rule (line \ref{e2}) offers a set of all these possible changes thanks to braces.
In fact the rule is encoded with a count atom at its head, which makes it a choice rule. Rule in line \ref{e3} filters any answer with more than 1 change at the same time $T$ to ensure the asynchronous evolution.
\begin{lstlisting}
{activeFromTo(B,J,K,T)} :-  playableAction(A,I,B,J,K,T), instate(activeProcess(A,I),T),
instate(activeProcess(B,J),T), J!=K, time(T). %\label{e2}
:- 2{ activeFromTo(B,J,K,T)}, time(T). %\label{e3}
\end{lstlisting}

In order to determinate the next active processes at $T+1$ we use the following rules :
\begin{lstlisting}
instate(activeProcess(B,K),T+1) :-  activeFromTo(B,J,K,T), time(T). %\label{e4}
instate(activeProcess(A,I),T+1) :-  instate(activeProcess(A,I),T), activeFromTo(B,J,K,T),
A!=B, time(T). %\label{e5}
\end{lstlisting}

At the next step $T+1$ we find the new active process resulted from the predicate $activeFromTo$ (line \ref{e4}) as well as all the unchanged processes that correspond to the other sorts (line \ref{e5}).\\
As a result it will be displayed all possible evolutions of the netwoks.
\subsection{Reachability verification}
\begin{itemize}

\item \textbf{known steps number } 
Parler de la méthode qui vérifie l'atteignabilité après un nombre fini de changements
\item \textbf{unknown steps number: incremental approch (iclingo)}
Parler de la méthode qui vérifie l'atteignabilité après un nombre indéfini de changements et qui retourne le chemin le plus court s'il existe pour atteindre ces objectifs
\end{itemize}

### 2014-12-16 fixPoint.tex
PH network evolution

@@
-%Section of Fix point implementation
-The study of fixed points (or basins of attraction) provides an important understanding of the different behaviors of a BRN (Biological Regulatory Network)\cite{wuensche1998genomic}.  
+The study of fixed points (or basins of attraction) provides an important understanding of the different behaviors of a BRN (Biological Regulatory Network) \cite{wuensche1998genomic}.  
 The fixed point is a stable state of the BRN in wich it is not possible any more to have new changes.
@@
 
-\begin{exemple} [Example PH network with ASP]
-If we try to present the network of Figure \ref{fig:expPH} we will have:
+\begin{example*} [Example PH network with ASP]
+If we try to present the network of Figure \ref{fig:ph} we will have:
 \begin{lstlisting}
@@
 Finally we find all network actions that are defined in lines \ref{actions1} and \ref{actions2}, for example the first action $action("a",0,"b",1,0)$ is an action from $a_0$ to $b_1$ to bound so $b$ from $b_1$ into $b_0$.
-\end{exemple}
+\end{example*}
 
@@
 \begin{lstlisting}
-hiddenProcess(A,I) :-  action(A,I,B,J,K), A=B ,process(A,I), process(B,J), process(B,K). %\label{hiddenProcess}
+hiddenProcess(A,I) :-  action(A,I,B,J,K), A=B, process(A,I),process(B,J),
+                       process(B,K). %\label{hiddenProcess}
 shownProcess(A,I) :-  not hiddenProcess(A,I), process(A,I). %\label{shownProcess}
@@
 
-\begin{example*}
+\begin{example}
 Considering the last graphical presentation of a precess hitting \pref{fig:ph}, there are 3 sorts $a$, $b$ have 2 levels and $z$ has 3 so we can find $2*2*3 = 12$ states (whatever they can be reached or not).
@@
 \end{tabbing}
-\end{example*}
+\end{example}

### 2014-12-16 comparaison.tex

In this section, we show the effectiveness of our approach on some examples. All
computations were performed on a Pentium V, 3.2 GHz with 4 GB RAM.

\subsection{Evaluation}
It should be noted that the only reachability analysis developed so far on the Process Hitting networks was implemented in the software Pint, and consists in an approximation: it is possible that it terminates but remains inconclusive (although this is rare). Moreover, it currently does not give us the path to activate the goal. But we also note that there are other methods or softwares for dynamical analysis applied to other biological network models.

We define the models used for each methos, logical regulatory network for GINsim (Gène Interaction Network
Simulation) \cite{gonzalez2006ginsim}, \cite{naldi2009logical} and \cite{naldi2007decision}, Instantiable Transition Systems for libddd (Library of Data Decision
Diagrams) \cite{thierry2009hierarchical} and \cite{colange2013towards} and state transition network for Rocca's et al. method \cite{roccaasp}.

For our comparative study we chose organic network of different sizes. As mentioned each method is applied on a specific model and translation into these models from the PH network or viceversa is done thanks to PINT \cite{pauleve2011modelisation}.
The following table prensents the comparisons run for 4 reachability analysis methods  applied to large examples:
\begin{center}

\begin{figure}
\label{tab:reachability}
\noindent
\begin{tabular}{|l||c|c|c||c|c|c|c|}
  \hline
   Model& \#procs & \#states & target & Rocca & libddd & GINsim & ASP \\
  \hline
  tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & process & 7m17.01s & XX & XX & 0m01.90s \\
  \hline
  ERBB\_G1-S \cite{Samaga2009}  & 152 & $2^{70}$ & state & - &1m55.38s & 2m01.64s & 0m11.84s \\
  \hline
  ERBB\_G1-S & 52 & $2^{70}$ & process & - &1m54.96s & - & 0m05.02s \\
  \hline
  TCRsig40 \cite{Klamt06} & 156 & $2^{73}$ & process & - & out & out & 0m05.02s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP. For each test we find the name of the model, the number of its processes and its states then the target(s). The target can be a network state (set of processes) or one or some processes. "out" marks an execution asking too much time or memory and "-" marks that is not possible to do the test.
}
\end{figure}
\end{center}

This table shows the effectiveness of our method compared to existing. In fact if we start by talking about Rocca's method, actually it offers the possibility to verify all model checking properties (LTL et CTL) for transition state networks. But we note that it is very difficult to write a transition state network in ASP, for example the network of tetard \cite{khalis2009smbionet} only 42 processes, take 3 minutes to be translated from PH.

Second the \textsc{libddd}

Third the \textsc{GINsim} is an analysis software and simulation of genetic interaction networks. It computes the stable states instantly as welle as our ASP method do it instantly.The difference that \textsc{GINsim} does not return all stable states but only reachable ones. Regarding the reachability problem, its approach is to calculate all the transition state graph and then check if there is a path between two given states. Thus it is necessary to have the entire network transition states and both initial and final states to determine the path. It was indicated in the third line of the table \ref{tab:reachability} that GINsim could not solve the problem because it is not possible to check the reachability of just one or a few components of the network indeed must be given a final state.

%We propose to evaluate the over-approximation
%due to our way of handling hybrid dynamics. For that purpose, we consider
%the 5-dimensional linear differential

%\subsection{Benchmarks }

\subsection{Limitation}

### 2014-12-16 reachability.tex
CLNS format

@@
 \begin{lstlisting}
-{activeFromTo(B,J,K,T)} :-  playableAction(A,I,B,J,K,T), instate(activeProcess(A,I),T),
-instate(activeProcess(B,J),T), J!=K, time(T). %\label{e2}
+{activeFromTo(B,J,K,T)} :-  playableAction(A,I,B,J,K,T),
+          instate(activeProcess(A,I),T), instate(activeProcess(B,J),T),
+          J!=K, time(T). %\label{e2}
 :- 2{ activeFromTo(B,J,K,T)}, time(T). %\label{e3}
@@
 instate(activeProcess(B,K),T+1) :-  activeFromTo(B,J,K,T), time(T). %\label{e4}
-instate(activeProcess(A,I),T+1) :-  instate(activeProcess(A,I),T), activeFromTo(B,J,K,T),
-A!=B, time(T). %\label{e5}
+instate(activeProcess(A,I),T+1) :-  instate(activeProcess(A,I),T),
+          activeFromTo(B,J,K,T), A!=B, time(T). %\label{e5}
 \end{lstlisting}
@@
 \subsection{Reachability verification}
-\begin{itemize}
+In this section, we focus on the reachability of a process which corresponds to the question \cite{def:reachability} :\\
+“Is it possible, starting from a given initial state, to play a number of actions so that a given processes are active in the resulting state?”
 
-\item \textbf{known steps number } 
-Parler de la méthode qui vérifie l'atteignabilité après un nombre fini de changements
-\item \textbf{unknown steps number: incremental approch (iclingo)}
-Parler de la méthode qui vérifie l'atteignabilité après un nombre indéfini de changements et qui retourne le chemin le plus court s'il existe pour atteindre ces objectifs
-\end{itemize}
+\noindent Now we try to adapt the code of network evolution to resolve the reachability problem.
+First we define a predicate for the objective processes we call it $"goal"$, we add a rule with this predicate to the script defining the PH :
+\begin{lstlisting}
+goal(activeProcess("a",1)). %\label{c1}
+\end{lstlisting}
+The rule line \ref{c1} verifies if after the network evolution, its state satisfies the goals at step $T$. Else the answer will be eliminated.
+\begin{lstlisting}
+satisfiable(F,T) :-  goal(F), instate(F,T). %\label{c2}
+\end{lstlisting}
+The limitation of this method is that the user has to choose the number of steps of the evolution.
+It is an disadvantage because a search in N steps will find no solution if the shortest path to solve the rechability requires N+1 steps.
+The solution is to use the incremental computation mode ( \textsc{iclingo} \cite{gebser2008user}).
+So we have almost the same program expect for incremental step numbers.
+In each step t, the program computes the playable actions $playableAction(A,I,B,J,K,t)$,the possible change $activeFromTo(B,J,K,t-1)$ and the new active processes for the next step $instate(activeProcess(A,I),t+1)$.
+Regarding the part of local steps we use a special constraint (rule \textbf{c4}) that means that the program should continue to the next step if it's not satisfiable so that we eliminate responses that do not meet the goals.
+\begin{lstlisting}
+notSatisfiable(t) :-  goal(F), not instate(F,t). %\label{c3}
+:- notSatisfiable(t). %\label{c4}
+\end{lstlisting}

### 2014-12-17 comparaison.tex
comparaison2

@@
 \subsection{Evaluation}
-It should be noted that the only reachability analysis developed so far on the Process Hitting networks was implemented in the software Pint, and consists in an approximation: it is possible that it terminates but remains inconclusive (although this is rare).
-Moreover, it currently does not give us the path to activate the goal.
-But we also note that there are other methods or softwares for dynamical analysis applied to other biological network models.
+To know the effectiveness of our new approach, we had to position ourselves against existing methods dealing with different biological network models.
+We have chosen the more known methods: GINsim (Gene Interaction Network
+Simulation)\cite{gonzalez2006ginsim}, \cite{naldi2009logical} and \cite{naldi2007decision}, LIBDDD (Library of Data Decision
+Diagrams) \cite{thierry2009hierarchical} and \cite{colange2013towards}, Rocca's et al. method \cite{roccaasp} and PINT \cite{pauleve2011modelisation}.
+Each method use a specific model, we find logical regulatory network for GINsim , Instantiable Transition Systems for LIBDDD  and state transition network for Rocca's et al. method \cite{roccaasp}.\\
+For our comparative study we chose organic network of different sizes.
+As mentioned each method is applied on a specific model and translation into these models from the PH network or viceversa is done thanks to PINT.
 
-We define the models used for each methos, logical regulatory network for GINsim (Gène Interaction Network
-Simulation) \cite{gonzalez2006ginsim}, \cite{naldi2009logical} and \cite{naldi2007decision}, Instantiable Transition Systems for libddd (Library of Data Decision
-Diagrams) \cite{thierry2009hierarchical} and \cite{colange2013towards} and state transition network for Rocca's et al. method \cite{roccaasp}.
-
-For our comparative study we chose organic network of different sizes.
-As mentioned each method is applied on a specific model and translation into these models from the PH network or viceversa is done thanks to PINT \cite{pauleve2011modelisation}.
-The following table prensents the comparisons run for 4 reachability analysis methods  applied to large examples:
 \begin{center}
@@
   \hline
+%   & \multicolumn{7}{l|}{Reachability}\\
+%  \hline
   tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & process & 7m17.01s & XX & XX & 0m01.90s \\
@@
   \hline
-  ERBB\_G1-S & 52 & $2^{70}$ & process & - &1m54.96s & - & 0m05.02s \\
+  ERBB\_G1-S & 152 & $2^{70}$ & process & - &1m54.96s & - & 0m05.02s \\
   \hline
@@
 \end{center}
+This table prensents the comparisons run for 4 dynamic analysis methods applied to examples of biological regulatory networks.
+The models of these networks where chosen as being representative of different scales, large networks and small networks.
+This table shows the effectiveness of our method compared to existing. In fact we develop in the sequel that our method provides more compared to other
 
-This table shows the effectiveness of our method compared to existing.
-In fact if we start by talking about Rocca's method, actually it offers the possibility to verify all model checking properties (LTL et CTL) for transition state networks.
-But we note that it is very difficult to write a transition state network in ASP, for example the network of tetard \cite{khalis2009smbionet} only 42 processes, take 3 minutes to be translated from PH.
+\begin{itemize}
 
-Second the \textsc{libddd}
+\item[--] \textbf{Rocca's method :} It offers the possibility to verify model-checking CTL properties of transition state networks.
+But we note that it is very difficult to write a transition state network in ASP, for example the network of tetard \cite{khalis2009smbionet} with only 42 processes, take 3 minutes to be translated from PH network.
 
-Third the \textsc{GINsim} is an analysis software and simulation of genetic interaction networks.
+\item[--] \textbf{\textsc{GINsim} :} It is an analysis software and simulation of genetic interaction networks.
 It computes the stable states instantly as welle as our ASP method do it instantly.The difference that \textsc{GINsim} does not return all stable states but only reachable ones.
 Regarding the reachability problem, its approach is to calculate all the transition state graph and then check if there is a path between two given states.
-Thus it is necessary to have the entire network transition states and both initial and final states to determine the path.
+But the large network transition state cannot be displayed.
+In order to determine the path it is necessary to have the entire network transition states and both initial and final states.
 It was indicated in the third line of the table \ref{tab:reachability} that GINsim could not solve the problem because it is not possible to check the reachability of just one or a few components of the network indeed must be given a final state.
 
-%We propose to evaluate the over-approximation
-%due to our way of handling hybrid dynamics. For that purpose, we consider
-%the 5-dimensional linear differential
+\item[--] \textbf{\textsc{libddd} :}
+This is a library for model-checking (CTL, LTL and reachability).
+It returns the answer "True" or "False" after verfying the reachability.
+However our approach is used to return the activation path of the affected components.
+In addition it is clear that \textsc{libddd} put much time to respond (23 times slower to example line 3 of the table).
+That's why for the big examples ($2^{73}$ states) we don't have a response even after 12 minutes but we get it in less than 2 minutes by using the ASP method.
+Also the \textsc{libddd} do not compute the stable states of a network.
 
-%\subsection{Benchmarks }
+\item[--] \textbf{PINT : }It should be noted that the only reachability analysis developed so far on the Process Hitting networks was implemented in the software Pint, and consists in an approximation: it is possible that it terminates but remains inconclusive (although this is rare).
+Moreover, it currently does not give us the path to activate the goal.
+\end{itemize}
 
 \subsection{Limitation}
+
+We developped new methods for the dynamic properties, 
+identifying stable states and finding all possible paths to reach states.
+Comparing with other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}) is relatively faster and also permit to edit larger networks (about $2^{73}$ states) but we had also some inconclusif cases.
+In fact the program is inconclusif if the state is not reachable and there is some loops in the graph so the program may turn indefinitely in one of these loops since it is not possible to achieve the goals or to stop the calculation.
+We then proposed in this case to limit the increment to a maximum number that can be reached in the case of an endless loop and which may not be in the case where the targets are verified.
+\begin{tabbing}
+ \texttt{ iClingo find\_path.lp <network\_name>.lp --imax=n}
+\end{tabbing}
+Should be given some interesting values for this integer $ n $, which sets the maximum number of steps.
+For example, the total number of states is a maximum (it will never be exceeded by a minimum path).
+However, it can reduce the number of sorts if it is assumed that one type will be changed by step (which is often the case of Boolean networks) or the number of shares with the same assumption.
+Note that if in a way the goal is reached after a time $n$, the calculation stops and displays this path as solution.

### 2014-12-18 PH.tex
definitions

@@
 An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
-``$a_i$ \emph{hits} $b_j$ to make it bounce to $b_k$'',
+``$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$'',
 where $a_i,b_j,b_k$ are processes of sorts $a$ and $b$,
@@
 
-\begin{definition}[Process Hitting]\label{def:PH}
+\begin{definition}[Process Hitting]\label{def:PH} \\
   A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
@@
 
-\begin{definition}[Action]
-\label{def:PhAction}
-An action is noted $\PHfrappe{a_i}{b_j}{b_k}$ where $a_i$ is a process
-of sort $a$ and $b_j$, $b_k$ two processes of sort $b$. When $a_i = b_j$ , such an action is refered as a self-action and $a_i$
-is called a self-hitting process.
-\end{definition}
+%\begin{definition}[Action]
+%\label{def:PhAction}
+%An action is noted $\PHfrappe{a_i}{b_j}{b_k}$ where $a_i$ is a process
+%of sort $a$ and $b_j$, $b_k$ two processes of sort $b$. When $a_i = b_j$ , such an action is refered as a self-action and $a_i$
+%is called a self-hitting process.
+%\end{definition}
 
@@
 \caption{\label{fig:ph} 
-A PH model example with three sorts: $a$, $b$ and $z$($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2).
-Circles represent the processes, boxes represent the sorts, and the actions are drawn by pairs of arrows in solid and dotted lines.The grayed processes stand for a possible initial state.
+A PH model example with three sorts: $a$, $b$ and $z$ ($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2).
+Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 4 \emph{actions} that assure the dynamic behavior are drawn by pairs of arrows in solid and dotted lines.
+The grayed processes stand for a possible initial state.
 }
 \end{figure}
+\end{example*}
+A network state is a set of active processes and there is a single process for each kind.
+In a state $ s \in \PHl$, we say that a process $ a_i \in s$ with $ a $ the name of the sort and $ i $ the level of this sort at this state, and we denote by $s[a]= a_i$.
 
-\end{example*}
+\begin{definition} [Playable action]
+\label{def:playableAction}\\
+Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
+We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is playable at the state $s$ if and only if $hitter(h)=a_i ~ \in s$ and $target(h)=b_j \in s$ (i.e $s[a]=a_i$ et $s[b]=b_j$ ) \\
+
+\end{definition}
+
+The resulting state after playing an action $h$ at an initial state $s$ is denoted by $(s \play h)$ or $(s \play h)[b]=b_k$ and $\forall c \in \PHs, ~ c \neq b, (s \play h)[c]=s[c].$
+
+\subsection*{Dynamic properties}
+
+The study of the dynamics of biological networks was the focus of many researchers, justifying the diversity of network modelisations and the different methods developed in order to verify the dynamic properties.
+In this paper we present 2 properties: the stable state and the reachability.
+We develop in this section these properties and how they could be verified in PH network.
+
+The dynamic of a network or its evolution is realised through its actions.
+In fact when an action $h = \PHfrappe{a_i}{b_j}{b_k}$ is played at a state $s$, the network evolves to the next state results $s'$.
+We found so in the state $s'$ all processes of $s$ with the replacement of process $b_j=cible(h)$ by $b_k=bond(h)$.
 
 \begin{definition}[Next state]
-\label{def:NextState}
+\label{def:NextState} \\
 Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
@@
 \end{center}
-
 \end{definition}
 
-\begin{definition}[Stable state or fixed point]
+We call the set of succession of actions ($h_0 \play h_1 \play h_2...$) from a state $ s_0 $ a \emph{scenario} noted $\Sce(s_0)$ corresponding to the succession of states ($s_0 \mapsto s_1 \mapsto s_2 ...$).
+
+\begin{definition}[Stable state or fixed point] \\
 \label{def:FixPoint1}
 Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
-$s \in \PHl$ be a state, $s$ is a stable state for $PH$ if and only if $next(s) = \emptyset $.
+$ \{s_0,..., s_n\} \in \PHl$ be state, $s_n$ is a stable state for $PH$ if and only if $next(s_n) = \emptyset $.
+\begin{center}
+$s_0 \mapsto s_1 ... \mapsto s_n \mapsto \emptyset $
+\end{center}
 \end{definition}
-\begin{definition}[Lemme: Stable state or fixed point]
+
+\begin{definition}
+[Lemme: Stable state or fixed point]
 \label{def:FixPoint2}
@@
 
- If $\ctx \in \PHl$ is a state and $A \in \Proc$ a set of processes, we note $\mathcal{P}(\ctx, A)$ pour \emph{la question d'accessibilité}:
+ If $s \in \PHl$ is a state and $A \in \Proc$ a set of processes, we note $\mathcal{P}(s, A)$ for the \emph{the reachability question}:
  \begin{center}
-  $\exists? \delta \in \Sce(\ctx), A \subseteq \PHget(\ctx \play \delta) $ (i.e $\forall a_i \in A, \PHget{(\ctx \play \delta)}{a}=a_i$).
+  $\exists? \delta \in \Sce(s), A \subseteq \PHget(s \play \delta) $ (i.e $\forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$).
  \end{center}
-With $\Sce(\ctx)$ 
-is the set of successively playable actions from the state $\ctx$. 
+with $\Sce(s)$ 
+is the set of successively playable actions from the state $s$. 
 \end{definition}
-
-\begin{definition} [Playable action]
-\label{def:playableAction}
-Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
-We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is playable at the state $s$ if and only if $hitter(h)=a_i ~ \in s$ and $target(h)=b_j \in s$ (i.e $s[a]=a_i$ et $s[b]=b_j$ ) \\
-The resulting state after playing the action $h$ at $s$ is denoted by $(s \play h)$ or $(s \play h)[b]=b_k$ and $\forall c \in \PHs, ~ c \neq b, (s \play h)[c]=s[c].$
-\end{definition}

### 2015-01-08 ASP.tex
CLNS format
mise à jour

@@
-In order to develop the nes methods we have used two main frameworks.
-The first is a logic programming language: the Answer Set Programming (ASP).
-Precisely it is a logic programming language on which I wrote my bibliographic research report.
+In order to develop these methods we have used two main frameworks.
+The first is a logic programming language: the Answer Set Programming (ASP), precisely it is a logic programming language.
 The second framework is the Process Hitting (PH) a new formalism for the representation of biological regulatory networks.
@@
 \begin{enumerate}
-\item \textbf{L'alphabet}
+\item \textbf{The alphabet}
 
@@
 
-Special cases rules \cite{Vladimir, baral2003knowledge} :\\
+Special cases rules \cite{Vladimir} and \cite{baral2003knowledge} :\\
 

### 2015-01-08 PH.tex
mise à jour

@@
 An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
-``$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$'',
-where $a_i,b_j,b_k$ are processes of sorts $a$ and $b$,
+"$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$",
+where $a_i$, $b_j$, $b_k$ are processes of sorts $a$ and $b$,
 called respectively \emph{hitter}, \emph{target} and
@@
 
-\begin{definition}[Stable state or fixed point] \\
+\begin{definition}[Stable state or fix point] \\
 \label{def:FixPoint1}
@@
 \begin{definition}
-[Lemme: Stable state or fixed point]
+[Lemma: Stable state or fix point] \\
 \label{def:FixPoint2}

### 2015-01-08 comparaison.tex
mise à jour

@@
 \noindent
-\begin{tabular}{|l||c|c|c||c|c|c|c|}
+\begin{tabular}{|l||c|c|c||c|c|c|c|c|}
   \hline
-   Model& \#procs & \#states & target & Rocca & libddd & GINsim & ASP \\
+   Model& \#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
   \hline
-%   & \multicolumn{7}{l|}{Reachability}\\
-%  \hline
-  tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & process & 7m17.01s & XX & XX & 0m01.90s \\
+  tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & process & 7m17.01s & \todo{XX} & \todo{XX} & \todo{XX} & 0m01.90s \\
   \hline
-  ERBB\_G1-S \cite{Samaga2009}  & 152 & $2^{70}$ & state & - &1m55.38s & 2m01.64s & 0m11.84s \\
+  ERBB\_G1-S \cite{Samaga2009}  & 152 & $2^{70}$ & state & - & \todo{XX} &1m55.38s & 2m01.64s & 0m11.84s \\
   \hline
-  ERBB\_G1-S & 152 & $2^{70}$ & process & - &1m54.96s & - & 0m05.02s \\
+  ERBB\_G1-S & 152 & $2^{70}$ & process & - & 0m0.02s &1m54.96s & - & 0m05.02s \\
   \hline
-  TCRsig40 \cite{Klamt06} & 156 & $2^{73}$ & process & - & out & out & 0m05.02s \\
+  TCRsig40 \cite{Klamt06} & 156 & $2^{73}$ & process & - & \todo{XX} & out & out & 0m05.02s \\
   \hline
@@
 \end{tabbing}
-Should be given some interesting values for this integer $ n $, which sets the maximum number of steps.
+Should be given some interesting values for this integer \texttt{"n"}, which sets the maximum number of steps.
 For example, the total number of states is a maximum (it will never be exceeded by a minimum path).
 However, it can reduce the number of sorts if it is assumed that one type will be changed by step (which is often the case of Boolean networks) or the number of shares with the same assumption.
-Note that if in a way the goal is reached after a time $n$, the calculation stops and displays this path as solution.
+Note that if in a way the goal is reached after a time \texttt{"n"}, the calculation stops and displays this path as a solution.

### 2015-01-08 fixPoint.tex
mise à jour

@@
-The study of fixed points (or basins of attraction) provides an important understanding of the different behaviors of a BRN (Biological Regulatory Network) \cite{wuensche1998genomic}.  
+The study of fix points (or basins of attraction) provides an important understanding of the different behaviors of a BRN (Biological Regulatory Network) \cite{wuensche1998genomic}.  
 The fixed point is a stable state of the BRN in wich it is not possible any more to have new changes.
@@
 In order to handle a PH biological network, it was necessary first to present it with ASP.
-To do this we chose the predicates: $sort$, $process$ and $action$.
-Below is an example of PH network shown in ASP.
+To do this we chose the predicates: \texttt{sort}, \texttt{process} and \texttt{action}.
+Below is an example of PH network written in ASP.
 
@@
 Line \ref{ASPsort} shows every sort of network with the predicate comes out and in parentheses the name of the sort.
-In line \ref{ASPprocess} there is a list of processes corresponding to each $sort$ for example the sort "$z$" has 3 process numbered from 0 to 2, this numbering is provided by the $2^{nd}$ parameter of the predicate $process("z", 0..2)$.
-Finally we find all network actions that are defined in lines \ref{actions1} and \ref{actions2}, for example the first action $action("a",0,"b",1,0)$ is an action from $a_0$ to $b_1$ to bound so $b$ from $b_1$ into $b_0$.
+In line \ref{ASPprocess} there is a list of processes corresponding to each \texttt{sort}, for example the sort \texttt{"z"} has 3 processes numbered from \texttt{0} to \texttt{2}, this numbering is provided by the $2^{nd}$ parameter of the predicate \texttt{process("z", 0..2)}.
+Finally we find all network actions that are defined in lines \ref{actions1} and \ref{actions2}, for example the first action \texttt{action("a",0,"b",1,0)} is an action from $a_0$ to $b_1$ to bound so $b$ from $b_1$ into $b_0$.
 \end{example*}
 
-\subsection{Search of fixed points }
-Parler du point fixe et ses interrêts. \\
-Decrire la nouvelle méthode qui permettent de determiner les états stables d'un réseau
+\subsection{Search of fix points }
+\todo{ Parler du point fixe et ses interrêts. \\
+Décrire la nouvelle méthode qui permet de déterminer les états stables d'un réseau }
 
-In fact we have to translate the definition of a stable state in a method developped in ASP. So first we eliminate all processes with self-hit and we save them in the predicate "$shownProcess$":
+In fact we have to translate the definition of a stable state in a method developped in ASP. So first we eliminate all processes with self-hit and we save them in the predicate \texttt{shownProcess}:
 \begin{lstlisting}
@@
 \begin{example}
-Considering the last graphical presentation of a precess hitting \pref{fig:ph}, there are 3 sorts $a$, $b$ have 2 levels and $z$ has 3 so we can find $2*2*3 = 12$ states (whatever they can be reached or not).
+Considering the last graphical presentation of a precess hitting \pref{fig:ph}, there are 3 sorts \texttt{"a"}, \texttt{"b"} have 2 levels and \texttt{"z"} has 3, so we can find $2*2*3 = 12$ states (whatever they can be reached or not).
 If we verify wether there exist fixed points we deduce that we have only one: $<b\_0, z\_2, a\_0>$.

### 2015-01-08 reachability.tex
mise à jour

@@
 From an initial known state, a PH network can evolve into several new states after a few steps.
-The predicate $"time (0 ..
-n)"$ sets the number of steps we want to play.
-For example if the biologist wants to know which states are reachable after 10 stages, it has only to replace $n$ by 10, and he will have $"time (0 ..
-10)"$.
+The predicate \texttt{"time (0 ..
+n)"} sets the number of steps we want to play.
+For example if the biologist wants to know which states are reachable after 10 stages, it has only to replace \texttt{"n"} by 10, and he will have \texttt{"time (0 ..
+10)"}.
 To compute the future states it is necessary to define an initial state of the network.
@@
 \end{lstlisting}
-with $a$ is the name of the sort and $0$ the index of the active process.\\
+with \texttt{"a"} is the name of the sort and\texttt{"0"} the index of the active process.\\
 The dynamic of a network is realised thanks to its actions.
@@
 We recall that an action is playable when both precesses: hitter and taget are active (Definition \ref{def:playableAction}).
-In ASP the predicate $"playableAction(A,I,B,J,K,T)"$ it is true when the processes $(A,I)$ and $(B,J)$ are active at the step $T$.
-Our approch is to study the asynchronous dynamic so only one action is playable at the same step ($T$).
+In ASP the predicate \texttt{"playableAction(A,I,B,J,K,T)"} it is true when the processes \texttt{(A,I)} and \texttt{(B,J)} are active at the step \texttt{T}.
+Our approch is to study the asynchronous dynamic so only one action is playable at the same step ( \texttt{T}).
 That's why between two successive states there will be one change of one sort.
-We present this by the predicate $"activeFromTo(B,J,K,T)"$.
-It means that in the sort $B$ the active process change from index $J$ to $K$ at time step $T$.
+We present this by the predicate  \texttt{"activeFromTo(B,J,K,T)"}.
+It means that in the sort \texttt{B} the active process change from index \texttt{J} to \texttt{K} at time step \texttt{T}.
 At the same step we can find many possible changes , that's why the rule (line \ref{e2}) offers a set of all these possible changes thanks to braces.
 In fact the rule is encoded with a count atom at its head, which makes it a choice rule.
-Rule in line \ref{e3} filters any answer with more than 1 change at the same time $T$ to ensure the asynchronous evolution.
+Rule in line \ref{e3} filters any answer with more than 1 change at the same time \texttt{T} to ensure the asynchronous evolution.
 \begin{lstlisting}
@@
 
-In order to determinate the next active processes at $T+1$ we use the following rules :
+In order to determinate the next active processes at \texttt{T+1} we use the following rules :
 \begin{lstlisting}
@@
 
-At the next step $T+1$ we find the new active process resulted from the predicate $activeFromTo$ (line \ref{e4}) as well as all the unchanged processes that correspond to the other sorts (line \ref{e5}).\\
+At the next step \texttt{T+1} we find the new active process resulted from the predicate \texttt{activeFromTo} (line \ref{e4}) as well as all the unchanged processes that correspond to the other sorts (line \ref{e5}).\\
 As a result it will be displayed all possible evolutions of the netwoks.
 \subsection{Reachability verification}
-In this section, we focus on the reachability of a process which corresponds to the question \cite{def:reachability} :\\
+In this section, we focus on the reachability of a process which corresponds to the question \ref{def:reachability} :\\
 “Is it possible, starting from a given initial state, to play a number of actions so that a given processes are active in the resulting state?”
@@
 \noindent Now we try to adapt the code of network evolution to resolve the reachability problem.
-First we define a predicate for the objective processes we call it $"goal"$, we add a rule with this predicate to the script defining the PH :
+First we define a predicate for the objective processes we call it \texttt{"goal"}, we add a rule with this predicate to the script defining the PH :
 \begin{lstlisting}
@@
 \end{lstlisting}
-The rule line \ref{c1} verifies if after the network evolution, its state satisfies the goals at step $T$. Else the answer will be eliminated.
+The rule line \ref{c1} verifies if after the network evolution, its state satisfies the goals at step \texttt{T}. Else the answer will be eliminated.
 \begin{lstlisting}
@@
 So we have almost the same program expect for incremental step numbers.
-In each step t, the program computes the playable actions $playableAction(A,I,B,J,K,t)$,the possible change $activeFromTo(B,J,K,t-1)$ and the new active processes for the next step $instate(activeProcess(A,I),t+1)$.
+In each step t, the program computes the playable actions \texttt{playableAction(A,I,B,J,K,t)},the possible change \texttt{activeFromTo(B,J,K,t-1)} and the new active processes for the next step \texttt{instate(activeProcess(A,I),t+1)}.
 Regarding the part of local steps we use a special constraint (rule \textbf{c4}) that means that the program should continue to the next step if it's not satisfiable so that we eliminate responses that do not meet the goals.

### 2015-01-09 comparaison.tex
modif comparaison

@@
   \hline
-  tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & process & 7m17.01s & \todo{XX} & \todo{XX} & \todo{XX} & 0m01.90s \\
+  tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
   \hline
-  ERBB\_G1-S \cite{Samaga2009}  & 152 & $2^{70}$ & state & - & \todo{XX} &1m55.38s & 2m01.64s & 0m11.84s \\
+  ERBB\_G1-S \cite{Samaga2009}  & 152 & $2^{70}$ & state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
   \hline
-  ERBB\_G1-S & 152 & $2^{70}$ & process & - & 0m0.02s &1m54.96s & - & 0m05.02s \\
+  ERBB\_G1-S & 152 & $2^{70}$ & process & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
   \hline
-  TCRsig40 \cite{Klamt06} & 156 & $2^{73}$ & process & - & \todo{XX} & out & out & 0m05.02s \\
+  TCRsig40 \cite{Klamt06} & 156 & $2^{73}$ & process & - & 0m0.014s & out & - & 0m05.02s \\
   \hline
@@
 \item[--] \textbf{PINT : }It should be noted that the only reachability analysis developed so far on the Process Hitting networks was implemented in the software Pint, and consists in an approximation: it is possible that it terminates but remains inconclusive (although this is rare).
-Moreover, it currently does not give us the path to activate the goal.
+Moreover, it currently does not give us the path to activate the goal but only the "True" or "False" of the reachability.
+In the first example \textsc{PINT} did not give an answer because of the number of goals is greater than 5.
 \end{itemize}

### 2015-01-15 ASP.tex
fix broken reference

@@
 \item\textbf{ Modeling steps}
-The modeling process can be regarded as a special form of representation whose operations are detailed in \cite{{baral2003knowledge}}.
+The modeling process can be regarded as a special form of representation whose operations are detailed in \cite{baral2003knowledge}.
 Logical ASP programs follow the strategy of "generate and test".

### 2015-01-15 intro.tex
mise à jour
proofread intro

@@
+\todo{à revoir} \\
 As regulatory phenomena play a crucial role in biological systems, they need to
@@
 it possible to determine the possible evolutions of all the interacting components of the system.
-Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently use a new formalism, named the “Process Hitting” (PH) \cite{pauleve2011modelisation}, to model concurrent systems having components with a few qualitative levels.
+Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{pauleve2011modelisation}, to model concurrent systems having components with a few qualitative levels.
 A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.
-This particular structure makes the formal analysis of BRNs with hundreds of components tractable.
-PH is suitable, according to the precision of this information, to model BRNs with different levels of abstraction by capturing the most general dynamics.
+This particular structure makes the formal analysis of BRNs with hundreds of components tractable \modMF{by using approached methods computing approximations of the dynamics~\cite{PMR12-MSCS}}.
+Furthermore, PH is suitable, according to the precision of the available information, to model BRNs with different levels of abstraction by capturing the most general dynamics.
+
 The objectives of the work presented in this paper are the following.
-
-Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points \cite{wuensche1998genomic}).
-We perform an exhaustive search of the possible states, combination processes, one process from each sort and then check if it is a fixed point.
-
+Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points \cite{wuensche1998genomic} \annotMF{Est-ce que cette ref est pertinente ?
+Elle me semble plutôt axée sur la notion d'attracteur}).
+For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fixed points.
 The second phase of our work consists in computing the dynamics.
 It consists in determining from a known initial state the possible next states of the PH model.
-Finally we verify if we can reach a specific state of one or several component (gene or protein).
-The results are ensured to respect the PH dynamics.
+Then we verify if we can reach a specific state for one or several component (genes or proteins).
+The results are ensured to respect strictly the PH dynamics.
 
@@
 %Furthermore, we claim that we are able to deal with large biological networks (42 sorts).
+
+\annotMF{Je pense que dans toute l'intro, il faut remettre ASP au centre car c'est bien lui qui rend notre contribution particulière.
+Il ne faudrait pas séparer Objectifs/Contributions car les deux se recoupent, ou trouver un moyen de parler d'ASP avant de présenter le travail sur les points fixes et la dynamique.}
+
+\modMF{%
+Section \ref{defs} defines the paradigm of Answer Set Programming,
+and the Process Hitting framework.
+We then use ASP in section \ref{fixpoint} to tackle the enumeration
+of all fixed points in a model,
+and in section \ref{dynamics} in order to compute the dynamics of a model
+and check reachability properties.
+Finally, a comparison of the methods developed in this paper is
+proposed in section \ref{comparison},
+and in section \ref{ccl} we conclude and talk about future work.
+}%

### 2015-01-15 intro.tex
change refs to PMR10 and PMR12

@@
 it possible to determine the possible evolutions of all the interacting components of the system.
-Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{pauleve2011modelisation}, to model concurrent systems having components with a few qualitative levels.
+Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels.
 A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.

### 2015-01-15 ASP.tex
proofread ASP section

@@
-In order to develop these methods we have used two main frameworks.
-The first is a logic programming language: the Answer Set Programming (ASP), precisely it is a logic programming language.
-The second framework is the Process Hitting (PH) a new formalism for the representation of biological regulatory networks.
-We will see later in this section more details about these two frameworks.
+In order to develop these methods we have used two main frameworks. The first is both a programming paradigm and a declarative programming language: the Answer Set Programming (ASP).
+The second framework is the Process Hitting (PH) a previously introduced formalism for the representation of biological regulatory networks.
 
@@
 
-The ASP is a declarative programming paradigm with known semantics as the semantics of answer sets.
-This paradigm allows the programmer to specify what the problem to be solved and not how to solve it.
-ASP programs are written in AnsProlog* (short for "Answer Set Programming in Logic" with superscript *).
-These programs are composed of a set of facts and a set of rules from which other facts can be derived.
-A consistent set of facts that can be derived from a program using the rules is called "answer set" for the program.
+\emph{Answer Set Programming} (ASP) is a declarative programming paradigm with semantics known as the \emph{semantics of answer sets}.
+This paradigm allows the programmer to specify what the problem to be solved is, and not how to solve it.
+ASP programs are written in AnsProlog* (short for "Answer Set Programming in Logic" with superscript:~*).
+These programs are composed of a set of \emph{facts} and a set of \emph{rules}, as defined below, from which other facts can be derived.
+A consistent set of facts that can be derived from a program using the rules is called \emph{answer set} for the program.
 The sets of possible responses to an AnsProlog* program are calculated with a program called a solver.
@@
 \subsubsection{Logic program}
-Consider an ASP logic program, as each rule is an ordered pair,\cite{baral2003knowledge} :
-
+An ASP program is composed of rules, each one being an ordered pair~\cite{baral2003knowledge}:
 \begin{equation} \label{eq1ASP}
@@
 \end{equation}
-with $Head$ and $Body$ sets of literals.
+with $head$ and $body$ two sets of literals.
 
 This simple language has many advantages that make it one of the most effective in terms of knowledge representation, reasoning and declarative problem solving. 
+\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”}
 
@@
 
-\noindent{ According \cite{baral2003knowledge}, the alphabet of the axiom (or just the alphabet) of the framework of the answer set is comprised of seven classes of symbols:
+\noindent{ According to \cite{baral2003knowledge}, the alphabet of the axioms (or simply called the \emph{alphabet}) of ASP is composed of seven classes of symbols:
 \begin{itemize}
@@
 \item Connectors,	
-\item Symbols of punctuation, and	
-\item Special symbol $\bot$
+\item Symbols of punctuation,
+\item The special symbol $\bot$
 \end{itemize}
 }
-All these classes vary from one alphabet to another, but the sets of $ 5^{th} $ and $ 6^{th} $ classes (Connectors and symbols of punctuation) are defined as follows:\\
+All these classes vary from one alphabet to another, but the sets of $ 5^{th} $ and $ 6^{th} $ classes (Connectors and symbols of punctuation) are defined as follows:
+\begin{itemize}
+  \item Connectors with \{$\neg$, or, $\leftarrow$, not, ‘,’\}
+  \item Symbols of punctuation with \{‘(’, ‘)’, ‘.’\}
+\end{itemize}
 
-- Connectors with \{$\neg$, or, $\leftarrow$, not, ‘,’\} \\
-- Symbols of punctuation with \{‘(’, ‘)’, ‘.’\} \\
+\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
 
 The other classes remain constant as they meet some informal agreement.
-In general, variables start with a capital letter and contain letters and numbers(X,Y,..).
-The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter(f,g,a,b...).
-Sometimes there is the addition of a supplementary agreement which covers the letters used \cite{baral2003knowledge} : \\
+In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
+The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
+String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).
 
+\annotMF{Je ne suis pas sûr que l'explication sur les prédicats soit pertinente. Si on la garde, il faudrait la reformuler.}
 What seems a little fuzzy is the concept of predicate. Indeed the word predicate can be an innovation of the new grammar. Let's consider the following sentence: 
@@
 \item \textbf{The rules}
-Comme nous l’avons déjà mentionné, une règle est de la forme:
-
-\begin{equation} 
-head ~ \leftarrow ~ body.
-\tag{\eqref{eq1ASP}}
+As previously mentioned, a rule has the following form:
+%\begin{equation} 
+%head ~ \leftarrow ~ body.
+%\tag{\eqref{eq1ASP}}
+%\end{equation}
+\begin{equation} \label{eq3ASP}
+A ~ \leftarrow ~ B^+, ~ \ASPnot B^-.
 \end{equation}
-\begin{equation} \label{eq3ASP}
-A ~ \leftarrow ~ B^+, ~ \textbf{not} ~ B^-.
+where the part on the left of the arrow is called $head$
+and the right part is called $body$.
+This rule is then equivalent to:
+\begin{equation} \label{eq2ASP}
+L_{0} \leftarrow L_{1},...,L_{m}, \ASPnot L_{m+1},..., \ASPnot L_{n}.
 \end{equation}
-
-This rule is then equivalent to:
-
-\begin{equation} \label{eq2ASP}
-L_{0} \leftarrow L_{1},...,L_{m}, \textbf{not} L_{m+1},..., \textbf{not} L_{n}.
-\end{equation}
-
 And this rule states:
-
-$$
+\[
 \left.
     \begin{array}{ll}
-        \text{Si} ~ L_{1},...,L_{m} ~ \text{are ~ \textbf{true}}\\
-        \text{et} ~ L_{m+1},...,L_{n} ~ \text{are~ \textbf{false}}
+        \text{If} ~ L_{1},...,L_{m} ~ \text{are \textbf{true}}\\
+        \text{and} ~ L_{m+1},...,L_{n} ~ \text{are \textbf{false}}
     \end{array}
 \right \} \text{Then} ~ L_{0} ~ \text{is } \textbf{true} 
-$$
-with $L_{i}$ literals and k $\geq$ 0, m $\geq$ k and n $\geq$ m.
+\]
+with $L_{i}$ literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
+Therefore, $head$ is also called the \emph{conclusion} of the rule,
+and $body$ is its \emph{premise}.
 
-\underline{\textbf{Head}} is also called conclusion \\
-\underline{\textbf{Body}} is also called premise
-
-Special cases rules \cite{Vladimir} and \cite{baral2003knowledge} :\\
-
+There are some special cases of rules \cite{Vladimir,baral2003knowledge}:
 \begin{itemize}
-\item A rule is \textbf{constant} if all litterals are constants (noted with \textbf{ground});
-\item A \textbf{fact} or a \textbf{reality} : it is a rule with an empty body. It can be written even without the arrow $\leftarrow$ :\\
+\item A rule is said \textbf{constant} or \emph{ground} if all of its literals are constants;
+\item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
+  It can be written without the central arrow ($\leftarrow$) or with a body equal to $\top$:
 \begin{equation} 
-L_{0}. ~~ or ~~ L_{0} ~ \leftarrow ~  \top.
+L_{0} \leftarrow \top. \qquad \text{or} \qquad L_{0}.
  \label{eq5ASP}
@@
 
-\item A \textbf{constraint}:
-This false symbol to the head ($head$) is often eliminated and the constraint will be written generally as:
+\item A \textbf{constraint} is a rule where the head equals false ($\bot$).
+  In this case, the head is often not represented and the constraint will be written generally as:
 \begin{equation} 
- ~ \leftarrow ~ L_{1},...,L_{m}, ~ \textbf{not} ~ L_{m+1},..., ~\textbf{not} ~L_{n}.
+  \leftarrow L_{1},...,L_{m}, \ASPnot L_{m+1},..., \ASPnot L_{n}.
   \label{eq6ASP}
@@
 
-We say that a set of literals $X$ violates the constraint (\ref{eq6ASP}) if $\{L_{1},...,L_{m}\} \subseteq X$ and $\{L_{m+1},...,L_{n}\} \nsubseteq X$.
-If we have a program contain this type of rule \ref{eq6ASP}, then $X$ is an answer set of the program $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$, with $r$ such a constraint \ref{eq6ASP} and $X$ does not violate the constraint \ref{eq6ASP}.
+We say that a set of literals $X$ violates the constraint \eqref{eq6ASP} if $\{L_{1},...,L_{m}\} \subseteq X$ and $\{L_{m+1},...,L_{n}\} \nsubseteq X$.
+If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.
 
-\item Cardinality constraints:\\
-It is literals extended in the following form:
+\item A \emph{cardinality constraint} is a rule of the following form:
 \begin{equation} 
- ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u.
+ ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u \leftarrow body.
  \label{eq7ASP}
 \end{equation}
-
-with $m \geq 1$, $l$ an integer and $u$ can be an integer or by default the infinite if it does not exist.
-$l$ and $u$ are the lower and upper limits of the cardinality of the subsets of $ \{q_{1},~ ...
-,~ q_{m}\}$ that satisfy the answer sets.
-These literals are constrained ($q_{i}$) may occur in the head and body of the rule.
-Cardinality constraint is satisfied in an answer set $ X $, if the number of atoms of $ \{q_{1},~ ...
-,~ q_{m}\}$ belonging to $X$ is between $l$ and $u$.
-In other words:
-
-\begin{tabbing}
- $l~ \leq ~ | \{q_{1},... ,~ q_{m}\} ~ \cap ~ X | ~ \geq ~u$
-\end{tabbing}
-
-with $ \cap $ symbol of intersection and $ | A | $ is the cardinality of the set $ A $.
-
+with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$).
+The literals $q_i$ can also appear in the body of the rule or in other rules,
+and $l$ and $u$ are the lower and upper bounds of the number of these literals
+that can appear in the answer sets.
+%cardinality of the subsets of $ \{q_{1}, ... , q_{m}\}$ that satisfy the answer sets.
+Therefore, a cardinality constraint is satisfied in an answer set $X$, if the number of atoms of $\{q_{1}, ... , q_{m}\}$ belonging to $X$ is between $l$ and $u$,
+or, in other words, if:
+\begin{equation}
+  l \leq | \{q_{1},... , q_{m}\} \cap X | \leq u
+\end{equation}
+where $\cap$ is the symbol of the intersection of sets
+and $|A|$ is the cardinality of the set $A$.
+\modMF{%
+In the following, if they are not explicitly given,
+we consider that $l$ defaults to $0$ and $u$ defaults to $\infty$.
+\annotMF{Les cardinalités ne sont pas seulement utilisées de cette manière ! Dans notre code, on les trouve dans la tête et dans le corps des règles.}
+\todo{À compléter : cardinalités dans le corps des règles}
+}%
 \end{itemize}
@@
 
-\subsubsection{Modelisation of a problem with ASP}
+\subsubsection{Modeling of a problem with ASP}
 
@@
 \begin{itemize}
-\item it is a simplification of a given system
-\item it allows an action on the system  
+\item it is a simplification of a given system,
+\item it allows an action on the system.
+\annotMF{Quel genre d'action ?}
 \end{itemize}
-
-Models offer the possibility to provide a solution to a problem identified as such. This concept model can address another angle issues related to representation process.
+Models offer the possibility to provide a solution to a problem identified as such.
+\reformuler{This concept model can address another angle issues related to representation process.}
 
 \begin{enumerate}
-\item\textbf{ Modeling steps}
+\item\textbf{Modeling steps}
 The modeling process can be regarded as a special form of representation whose operations are detailed in \cite{baral2003knowledge}.
-Logical ASP programs follow the strategy of "generate and test".
+Logical ASP programs follow the strategy of “generate and test”.
 This strategy includes four steps:
-
 \begin{itemize}
 \item Enumerate with facts;
-\item Explain with the rules;
-\item Generate all the possibilities with cardinalities, and finally;
+\item Explain with rules;
+\item Generate all the possibilities with cardinalities; and finally,
 \item Filter with constraints.
@@
 
-\item \textbf{ Resolving an ASP program} \\
-From \cite{Vladimir} and \cite{AnsPrologAPE}, the couple Grounder and Solver usually work together: the grounder used to remove variables in order to achieve a constant program and the solver computes all anwser sets for stabilized programs generated by the grounder.
+\item \textbf{Solving an ASP program}
+\modMF{%
+The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
+that usually work together:
+}%
+the grounder is used to remove variables in order to achieve an equivalent but constant program,
+and the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
+Amongst all available grounders for ASP, we can name
+\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
+and the solvers include
+\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...
 
-Example of Grounder: \texttt{GRINGO}, \texttt{DLV}, \texttt{LPARSE}
-
-Example of ASP Solver : \texttt{SMODELS}, \texttt{DLV}, \texttt{CMODELS}, \texttt{CLASP}...
-
-The combined use of the Grouder and the Solver specifies major programs in a compact, using rules with schematic variables and other abbreviations.
+\reformuler{
+The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations.
 Both systems employ grounding algorithms that work quickly and simplify the program.
-In developping these new methods we worked with \texttt{CLINGO} which is a combination of grouder \texttt{GRINGO} and solver \texttt{clasp}.
+}
+For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.
 

### 2015-01-15 PH.tex
change refs to PMR10 and PMR12
proofread PH section

--- 2014-12-11 PH.tex
@@
-\pref{def:PH} introduces the Process Hitting~(PH)~\cite{pauleve2011modelisation}
+\pref{def:PH} introduces the Process Hitting~(PH)~\cite{PMR10-TCSB}
 which allows to model
@@
 An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
-``$a_i$ \emph{hits} $b_j$ to make it bounce to $b_k$'',
-where $a_i,b_j,b_k$ are processes of sorts $a$ and $b$,
+“$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$”,
+where $a_i$, $b_j$, $b_k$ are processes of sorts $a$ and $b$,
 called respectively \emph{hitter}, \emph{target} and
@@
 \begin{definition}[Process Hitting]\label{def:PH}
-  A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
+  A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$ where:
   \begin{itemize}
-    \item[--]  $\PHs = \{a,b,\dots\}$ is the finite set of \emph{sorts};
-    \item[--]  $\PHl = \prod_{a\in\PHs} \PHl_a$ is the set of states with
+    \item  $\PHs = \{a,b,\dots\}$ is the finite set of \emph{sorts};
+    \item  $\PHl = \prod_{a\in\PHs} \PHl_a$ is the set of \emph{states} where
       $\PHl_a = \{a_0,\dots,a_{l_a}\}$
-      the finite set of \emph{processes} of sort $a\in\Sigma$
-      and $l_a$ a positive integer, with $a\neq b\Rightarrow \PHl_a \cap \PHl_b = \emptyset$;
-    \item[--]  $\PHa = \{ \PHfrappe{a_i}{b_j}{b_k} \in \PHl_a \times \PHl_b^2 \mid
+      is the finite set of \emph{processes} of sort $a\in\Sigma$
+      and $l_a$ is a positive integer, with $a\neq b\Rightarrow \PHl_a \cap \PHl_b = \emptyset$;
+    \item  $\PHa = \{ \PHfrappe{a_i}{b_j}{b_k} \in \PHl_a \times \PHl_b^2 \mid
       (a,b) \in \PHs^2 \wedge b_j\neq b_k \wedge a=b\Rightarrow a_i=b_j\}$
@@
 \caption{\label{fig:ph} 
-A PH model example with three sorts: $a$, $b$ and $z$($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2).
-Circles represent the processes, boxes represent the sorts, and the actions are drawn by pairs of arrows in solid and dotted lines.The grayed processes stand for a possible initial state.
+A PH model example with three sorts: $a$, $b$ and $z$ ($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2).
+Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 5 \emph{actions} that model the dynamic behavior are depicted by pairs of arrows in solid and dotted lines.
+The grayed processes stand for the possible initial state: $\PHstate{a_1, v_0, z_1}$.
 }
 \end{figure}
+\end{example*}
+A state of the networks is a set of active processes containing a single process of each sort.
+\modMF{%
+The active process of a given sort $a \in \PHs$ in a state $s \in \PHl$
+is noted $\PHget{s}{a}$.
+For any given process $a_i$ we also note: $a_i \in s$ if and only if $\PHget{s}{a} = a_i$.
+}%
 
-\end{example*}
+\begin{definition} [Playable action]
+\label{def:playableAction}
+Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
+We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is \emph{playable in state $s$} if and only if $a_i \in s$ and $b_j \in s$ (\ie $\PHget{s}{a} = a_i$ and $\PHget{s}{b}=b_j$).
+\end{definition}
+The resulting state after playing an action $h$ in a state $s$ where it is playable is denoted by $(s \play h)$, where $\PHget{(s \play h)}{b} = b_k$ and $\forall c \in \PHs, c \neq b, \PHget{(s \play h)}{c}=\PHget{s}{c}$.
 
-\begin{definition}[Next state]
+\subsection*{Dynamic properties}
+
+The study of the dynamics of biological networks was the focus of many works, explaining the diversity of network modelings and the different methods developed in order to check dynamic properties.
+In this paper we focus on 2 main properties: the stable states and the reachability.
+We formally define these properties in the following and explain how they could be verified in a PH network.
+
+The dynamics of a network or its evolution is specified by its actions.
+Indeed, when an action $h = \PHfrappe{a_i}{b_j}{b_k}$ is played at a state $s$,
+the network evolves to a resulting next state $s'$ where exactly one active process
+is modified, namely: the active process $b_j$ changes into $b_k$.
+
+\todo{Glu pour la définition \ref{def:NextState}}
+
+\begin{definition}[Next states]
 \label{def:NextState}
@@
 \begin{center}
-$next(s) = {s[b_k /b_j ]| \exists (a i , b j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa}$
+$next(s) = \{s[b_k /b_j ]| \exists (a_i , b_j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa\}$
 \end{center}
-
+\todo{Définir la notation $s[b_k /b_j ]$}
+\annotMF{Est-ce que $next(s)$ est vraiment utile ? On ne s'en ressert pas après.}
 \end{definition}
 
-\begin{definition}[Stable state or fixed point]
+A sequence of successively playable actions ($h_0 \play h_1 \play h_2...$)
+from an initial state $s_0$ is called a \emph{scenario in $s_0$}.
+We note $\Sce(s_0)$ the set of all scenarios in $s_0$.
+
+\todo{Glu pour la définition \ref{def:FixPoint1} et le lemme \ref{def:FixPoint2}}
+
+\begin{definition}[Stable state or fix point]
 \label{def:FixPoint1}
 Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
-$s \in \PHl$ be a state, $s$ is a stable state for $PH$ if and only if $next(s) = \emptyset $.
+$s \in \PHl$ be state;
+$s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
+\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?}
 \end{definition}
-\begin{definition}[Lemme: Stable state or fixed point]
+
+\begin{definition}
+[Lemma: Stable state or fix point]
 \label{def:FixPoint2}
 Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
-$s \in \PHl$ be a state, $s$ is a stable state for $PH$ there is no playable action at the state $s$: 
+$s \in \PHl$ be a state;
+$s$ is a stable state for $PH$ if and only if there is no playable action in state $s$: 
 \begin{center}
-$\forall h \in \PHa$, $frappeur(h) ~ \notin s ~ \vee ~ cible(h) \notin s$)
+$\forall h \in \PHa$, $frappeur(h) \notin s \vee cible(h) \notin s$
 \end{center}
 \end{definition}
+
+\todo{Glu pour la définition \ref{def:reachability}}
 
@@
 
- If $\ctx \in \PHl$ is a state and $A \in \Proc$ a set of processes, we note $\mathcal{P}(\ctx, A)$ pour \emph{la question d'accessibilité}:
+ If $s \in \PHl$ is a state and $A \subseteq \Proc$ a set of processes,
+ we note $\mathcal{P}(s, A)$ the \emph{the reachability question}:
  \begin{center}
-  $\exists? \delta \in \Sce(\ctx), A \subseteq \PHget(\ctx \play \delta) $ (i.e $\forall a_i \in A, \PHget{(\ctx \play \delta)}{a}=a_i$).
+  $\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$
  \end{center}
-With $\Sce(\ctx)$ 
-is the set of successively playable actions from the state $\ctx$. 
+\todo{Il faut définir $\Proc$}
+\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}
 \end{definition}
-
-\begin{definition} [Playable action]
-\label{def:playableAction}
-Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
-We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is playable at the state $s$ if and only if $hitter(h)=a_i ~ \in s$ and $target(h)=b_j \in s$ (i.e $s[a]=a_i$ et $s[b]=b_j$ ) \\
-The resulting state after playing the action $h$ at $s$ is denoted by $(s \play h)$ or $(s \play h)[b]=b_k$ and $\forall c \in \PHs, ~ c \neq b, (s \play h)[c]=s[c].$
-\end{definition}

### 2015-01-15 fixPoint.tex
change refs to PMR10 and PMR12
begin proofread dynamics section

@@
-The study of fix points (or basins of attraction) provides an important understanding of the different behaviors of a BRN (Biological Regulatory Network) \cite{wuensche1998genomic}.  
-The fixed point is a stable state of the BRN in wich it is not possible any more to have new changes.
-Let $(\Sigma, L, H)$ be a Process Hitting .
- It has been shown that a state $s \in L$ is a fixed point of the Process Hitting if and only if $s$ is has no next state \cite{pauleve2011modelisation} i.e.
-there is no playable action at this state.
-In fact a steady state od a Process Hitting network is a set of processes with exactly one process of each sort as well as every process has no hit with the other selected ones (process with self-hit cannot be apart of a stable state).
+The study of fix points (and, more generally, basins of attraction) provides an important understanding of the different behaviors of a Biological Regulatory Network BRN~\cite{wuensche1998genomic}.
+\modMF{%
+Indeed, a system will always eventually end in a basin of attraction,
+and this may depend on biological switch or other complex phenomena.
+}%
+A fixed point is a state of the BRN in which it is not possible any more to have new changes;
+in other words, it is a basin of attraction that is composed of only one state.
 
-\subsection{Process Hitting network traduction }
-In order to handle a PH biological network, it was necessary first to present it with ASP.
-To do this we chose the predicates: \texttt{sort}, \texttt{process} and \texttt{action}.
-Below is an example of PH network written in ASP.
+In the following, we consider a Process Hitting $(\Sigma, \PHl, \PHh)$.
+It has been shown that a state $s \in L$ is a fixed point of the Process Hitting if and only if $s$ has no next state~\cite{PMR10-TCSB} \ie if there is no playable action in this state.
+Therefore, a steady state in a Process Hitting network is a state so that
+every process does not hit or is not hit by another process in the same state.
+We note that given this result, processes involved in a self-hit cannot be part of a stable state.
 
-\begin{example*} [Example PH network with ASP]
-If we try to present the network of Figure \ref{fig:ph} we will have:
+\subsection{Process Hitting translation in ASP}
+Before writing any piece of ASP,
+we first need to present the studied PH network in ASP.
+To do this we use the following self-describing predicates:
+\texttt{sort}, \texttt{process} and \texttt{action}.
+\annotMF{Décrire rapidement le sens de ces prédicats ?}
+
+\begin{example}[Representation of a PH network in ASP]
+\label{ex:asp-ph}
+The representation of the PH network of Figure \ref{fig:ph} in ASP is the following:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Line \ref{ASPsort} shows every sort of network with the predicate comes out and in parentheses the name of the sort.
-In line \ref{ASPprocess} there is a list of processes corresponding to each \texttt{sort}, for example the sort \texttt{"z"} has 3 processes numbered from \texttt{0} to \texttt{2}, this numbering is provided by the $2^{nd}$ parameter of the predicate \texttt{process("z", 0..2)}.
-Finally we find all network actions that are defined in lines \ref{actions1} and \ref{actions2}, for example the first action \texttt{action("a",0,"b",1,0)} is an action from $a_0$ to $b_1$ to bound so $b$ from $b_1$ into $b_0$.
-\end{example*}
+Line \ref{ASPsort} enumerates every sort of network.
+In line \ref{ASPprocess} we create the list of processes corresponding to each sort,
+for example the sort \texttt{"z"} has 3 processes numbered from \texttt{0} to \texttt{2};
+this specific predicate will in fact expand into the three following predicates:
+\texttt{process("z", 0)}, \texttt{process("z", 1)}, \texttt{process("z", 2)}.
+Finally, all the actions of the network are defined in lines \ref{actions1} and \ref{actions2};
+for example, the first predicate \texttt{action("a",0,"b",1,0)} represents the action
+$\PHfrappe{a_1}{b_1}{b_0}$.
+\end{example}
 
-\subsection{Search of fix points }
+\subsection{Search of fix points}
 \todo{ Parler du point fixe et ses interrêts. \\
@@
 
-In fact we have to translate the definition of a stable state in a method developped in ASP. So first we eliminate all processes with self-hit and we save them in the predicate \texttt{shownProcess}:
+The enumeration of fix points requires to translate the definition of a stable state
+in a set of ASP rules.
+The first step consists of eliminating all processes involved in a self-hit;
+the others are recorded by the predicate \texttt{shownProcess}:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Then we have to browse this graph and extract all possible combinations of shown processes by choosing a process from each sort.
+Then, we have to browse all remaining processes of this graph
+in order to generate all possible states,
+that is, all possible combinations of processes by choosing one process from each sort:
 \begin{lstlisting}
@@
 \end{lstlisting}
-It now remains to check each combination of processes whether it is a fix point.
-For this we use a special type of ASP rules: a \textbf{constraint}.
-The idea of constraints is based on the fact that a solution would be eliminated if it does not satisfy the constraint.
-For our problem a combination is eliminated if there is an action between two of the selected process:
+\modMF{%
+The previous line creates an many candidate answer sets as there are possible states
+to take into account.
+The last step consists of filtering each state that is not a fixed points,
+or, in other words, eliminating all candidate answer sets that do not
+represent a fixed point.
+For this, we use a constraint:
+indeed, any solution that satisfy the body of a constraint will be removed from the answer set.
+}%
+Regarding our problem, a state is eliminated if there exists an action linking two processes:
 \begin{lstlisting}
-:- hit(A,I,B,J), selectProcess(A, I), selectProcess(B, J), A!= B. %\label{contraintFix}
+:- hit(A,I,B,J), selectProcess(A, I), selectProcess(B, J), A != B. %\label{contraintFix}
 \end{lstlisting}
@@
 \begin{example}
-Considering the last graphical presentation of a precess hitting \pref{fig:ph}, there are 3 sorts \texttt{"a"}, \texttt{"b"} have 2 levels and \texttt{"z"} has 3, so we can find $2*2*3 = 12$ states (whatever they can be reached or not).
-If we verify wether there exist fixed points we deduce that we have only one: $<b\_0, z\_2, a\_0>$.
-It is clear in \pref{fig:ph} that there is no action between each two processes.
-Besides our new ASP method proof this and returns also the same answer:
-\begin{tabbing}
- \texttt{
+The PH model of figure \pref{fig:ph} contains 3 sorts:
+$a$ and $b$ have 2 processes and $z$ has 3, therefore, the whole model has $2*2*3 = 12$ states (whether they can be reached or not from a given initial state).
+We can check that this model contains only one fixed point: $\PHstate{b_0, z_2, a_0}$.
+Indeed, there is no action between each two of the processes contained is this state;
+moreover, no other state verifies this property.
+
+If we execute the ASP program detailed below,
+alongside with the description of the PS model given in example~\ref{ex:asp-ph},
+we obtain one answer set, which matches the expected result:
+\begin{lstlisting}[numbers=none]
 Answer 1 : fixProcess(a, 0), fixProcess(b, 0), fixProcess(z, 2)
- }
-\end{tabbing}
+\end{lstlisting}
 \end{example}

### 2015-01-15 reachability.tex
begin proofread dynamics section

@@
 
-In this section, we will present firstly how to determine using ASP the possible evolution of a biological network after a finite number of steps.
-Then we answer to the reachability question; which evolutions that allow the achievement of goals (future active processes) from a known initial state?
+In this section, firstly present how to determine the possible evolution of a PH model after a finite number of steps with an ASP program.
+Then we tackle the reachability question: are there scenarios in a given initial state
+that allow to reach a given goal (under the form of a set of processes)?
 
-\subsection{Future states identification }
-From an initial known state, a PH network can evolve into several new states after a few steps.
-The predicate \texttt{"time (0 ..
-n)"} sets the number of steps we want to play.
-For example if the biologist wants to know which states are reachable after 10 stages, it has only to replace \texttt{"n"} by 10, and he will have \texttt{"time (0 ..
-10)"}.
-To compute the future states it is necessary to define an initial state of the network.
-So for initializing the active processes we add in the ASP script representing the PH network the following rule for each sort.
-We note that this rules are generated automatically at the time of translating PH network to ASP and by default the level of all active processes is $0$.\\
+\subsection{Future states identification}
+\reformuler{From an initial known state, a PH network can evolve into several new states after a few steps.}
+\modMF{%
+Firstly, we focus on the evolutions of the model in a limited number of steps.
+}%
+The predicate \texttt{time(0..n)} sets the number of steps we want to play.
+\modMF{%
+The value of \texttt{n} can be arbitrarily chosen.
+For example, \texttt{time(0..10)} will be used to compute the 11 first steps,
+including the initial state.
+In order to specify such an initial state, we add several facts of the following form
+to list the processes included in this state:
+}%
 \begin{lstlisting}
@@
 \end{lstlisting}
-with \texttt{"a"} is the name of the sort and\texttt{"0"} the index of the active process.\\
-The dynamic of a network is realised thanks to its actions.
-So to identify the future states we should start by identifying the playable actions for each state.
-We recall that an action is playable when both precesses: hitter and taget are active (Definition \ref{def:playableAction}).
-In ASP the predicate \texttt{"playableAction(A,I,B,J,K,T)"} it is true when the processes \texttt{(A,I)} and \texttt{(B,J)} are active at the step \texttt{T}.
-Our approch is to study the asynchronous dynamic so only one action is playable at the same step ( \texttt{T}).
-That's why between two successive states there will be one change of one sort.
-We present this by the predicate  \texttt{"activeFromTo(B,J,K,T)"}.
-It means that in the sort \texttt{B} the active process change from index \texttt{J} to \texttt{K} at time step \texttt{T}.
-At the same step we can find many possible changes , that's why the rule (line \ref{e2}) offers a set of all these possible changes thanks to braces.
-In fact the rule is encoded with a count atom at its head, which makes it a choice rule.
-Rule in line \ref{e3} filters any answer with more than 1 change at the same time \texttt{T} to ensure the asynchronous evolution.
+where \texttt{"a"} is the name of the sort and\texttt{"0"} the index of the active process.
+The dynamics of a network is described by its actions,
+identifying the future states requires to first identify the playable actions for each state.
+We recall that an action is playable in a state when both its hitter and target are active in this state (see Definition \ref{def:playableAction}).
+Therefore, we define an ASP predicate \texttt{playableAction(A,I,B,J,K,T)} that is true
+when the processes \texttt{(A,I)} and \texttt{(B,J)} are active at step \texttt{T}.
+\annotMF{Peut-être écrire : $\texttt{A}_\texttt{I}$ et $\texttt{B}_\texttt{J}$}
+It is also needed to enforce the strictly asynchronous dynamic
+which state that exactly one process can change between two steps.
+We thus represent the change of the active process of a sort
+by the predicate \texttt{"activeFromTo(B,J,K,T)"}
+which means that in sort \texttt{B}, the active process changes from index \texttt{J} to \texttt{K} between steps \texttt{T} and \texttt{T+1}.
+\annotMF{Peut-être écrire : $\texttt{B}_\texttt{J}$ et $\texttt{B}_\texttt{K}$}
+\modMF{%
+The cardinality rule of line \ref{e2} then
+created as many answer sets as there are possible evolutions from the current step,
+thus recreating all possible branchings in the possible evolutions of the model.
+This allows to filter all scenarios where two actions have been played between
+two steps, by using the constraint of \ref{e3}.
+Thus, the remaining scenarios of the answer set all strictly follow
+the asynchronous dynamics of the PH.
+}%
 \begin{lstlisting}
@@
 \end{lstlisting}
+\annotMF{N'est-il pas possible de réunir ces deux règles avec des bornes 1--1 de la façon suivante : 1 \{activeFromTo(B,J,K,T)\} 1 $\leftarrow$ ... ?}
 

### 2015-01-16 reachability.tex
finish proofread dynamics section

@@
 This allows to filter all scenarios where two actions have been played between
-two steps, by using the constraint of \ref{e3}.
+two steps, by using the constraint of line \ref{e3}.
 Thus, the remaining scenarios of the answer set all strictly follow
@@
 \end{lstlisting}
-\annotMF{N'est-il pas possible de réunir ces deux règles avec des bornes 1--1 de la façon suivante : 1 \{activeFromTo(B,J,K,T)\} 1 $\leftarrow$ ... ?}
+\annotMF{N'est-il pas possible de réunir ces deux règles avec des bornes 1--1 de la façon suivante : 1 \{activeFromTo(B,J,K,T)\} 1 $\leftarrow$ ...
+?
+Sinon, voir à peut-être enlever la cardinalité pour une règle plus simple à comprendre.}
 
-In order to determinate the next active processes at \texttt{T+1} we use the following rules :
+Finally, the active processes at step \texttt{T+1},
+that represent the next state depending on the chosen bounce,
+can be computed by the following rules:
 \begin{lstlisting}
@@
 \end{lstlisting}
+In other words, the state of step \texttt{T+1} contains one new active process
+resulting from the predicate \texttt{activeFromTo} (line \ref{e4})
+as well as all the unchanged processes that correspond to the other sorts (line \ref{e5}).
 
-At the next step \texttt{T+1} we find the new active process resulted from the predicate \texttt{activeFromTo} (line \ref{e4}) as well as all the unchanged processes that correspond to the other sorts (line \ref{e5}).\\
-As a result it will be displayed all possible evolutions of the netwoks.
+\modMF{%
+The overall result of the pieces of program presented in this subsection
+is an answer set containing one answer for each
+possible evolution in \texttt{n} time steps,
+and starting from a given initial state.
+}%
+
 \subsection{Reachability verification}
-In this section, we focus on the reachability of a process which corresponds to the question \ref{def:reachability} :\\
-“Is it possible, starting from a given initial state, to play a number of actions so that a given processes are active in the resulting state?”
+In this section, we focus on the reachability of set of processes which corresponds to the question (see definition \ref{def:reachability}):
+“Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
+\annotMF{En fin de rédaction, il faudrait voir à peut-être décaler les définitions de la section “Dynamic properties” afin de faciliter la lecture.}
 
-\noindent Now we try to adapt the code of network evolution to resolve the reachability problem.
-First we define a predicate for the objective processes we call it \texttt{"goal"}, we add a rule with this predicate to the script defining the PH :
+We now want to adapt the code of the dynamics computation of previous section in order to resolve this reachability problem.
+For this, we first use a predicate to list the processes we want to reach, called \texttt{goal}, and we add as many rules of the following form as there are objective processes:
 \begin{lstlisting}
@@
 \end{lstlisting}
-The rule line \ref{c1} verifies if after the network evolution, its state satisfies the goals at step \texttt{T}. Else the answer will be eliminated.
+\modMF{%
+Then, the literal \texttt{satisfiable(F,T)}
+checks if a given process \texttt{F} of the goal
+is contained in the state of step \texttt{T},
+as defined in the rule of line \ref{c1}.
+}%
+Else the answer will be eliminated.
+\todo{Il manque la contrainte qui élimine les mauvais parcours !}
 \begin{lstlisting}
@@
 \end{lstlisting}
-The limitation of this method is that the user has to choose the number of steps of the evolution.
-It is an disadvantage because a search in N steps will find no solution if the shortest path to solve the rechability requires N+1 steps.
-The solution is to use the incremental computation mode ( \textsc{iclingo} \cite{gebser2008user}).
-So we have almost the same program expect for incremental step numbers.
-In each step t, the program computes the playable actions \texttt{playableAction(A,I,B,J,K,t)},the possible change \texttt{activeFromTo(B,J,K,t-1)} and the new active processes for the next step \texttt{instate(activeProcess(A,I),t+1)}.
-Regarding the part of local steps we use a special constraint (rule \textbf{c4}) that means that the program should continue to the next step if it's not satisfiable so that we eliminate responses that do not meet the goals.
+
+However, the limitation of the method above is that the user has to decide upstream
+the number of computed steps that should be sufficient.
+It is a main disadvantage because a search in $N$ steps will find no solution
+if the shortest path to the goal requires $N+1$ steps.
+\annotMF{Est-ce que ça ne risque pas aussi de rallonger anormalement la résolution si le plus court chemin est en $n$ étapes avec $n << N$ ?}
+One solution is to use an incremental computation mode,
+which is especially tackled by by the incremental solver \textsc{iclingo}~\cite{gebser2008user}.
+\annotMF{Peut-on citer le manuel d'utilisation de Clingo comme référence ?}
+\todo{Expliquer brièvement en quoi consiste la résolution incrémentale en ASP, notamment les trois parties distinctes : définitions, corps d'itération, fin d'itération.}
+\modMF{%
+Thus, the obtained program is almost identical,
+except that step numbers \texttt{N}
+are replaced by constant placeholders \texttt{t}
+that will be incremented throughout the execution.
+}%
+In each step \texttt{t}, the program computes the playable actions \texttt{playableAction(A,I,B,J,K,t)}, the possible bounces \texttt{activeFromTo(B,J,K,t-1)}
+and the new states \texttt{instate(activeProcess(A,I),t+1)}
+in the same fashion than previously.
+Regarding the part of local steps,
+\todo{Définir “the part of local steps”}
+we use a special constraint (line \textbf{c4})
+which states that the program should continue to the next step
+while it is not satisfiable,
+so that all answers that do not meet the goals are eliminated:
 \begin{lstlisting}

### 2015-01-16 comparaison.tex
change refs to PMR10 and PMR12
proofread comparison section

@@
-In this section, we show the effectiveness of our approach on some examples. All
-computations were performed on a Pentium V, 3.2 GHz with 4 GB RAM.
+In this section, we show the effectiveness of our approach on some examples,
+\modMF{%
+and compare it to other existing approaches.
+}%
+All computations were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
 
 \subsection{Evaluation}
-To know the effectiveness of our new approach, we had to position ourselves against existing methods dealing with different biological network models.
-We have chosen the more known methods: GINsim (Gene Interaction Network
-Simulation)\cite{gonzalez2006ginsim}, \cite{naldi2009logical} and \cite{naldi2007decision}, LIBDDD (Library of Data Decision
-Diagrams) \cite{thierry2009hierarchical} and \cite{colange2013towards}, Rocca's et al. method \cite{roccaasp} and PINT \cite{pauleve2011modelisation}.
-Each method use a specific model, we find logical regulatory network for GINsim , Instantiable Transition Systems for LIBDDD  and state transition network for Rocca's et al. method \cite{roccaasp}.\\
-For our comparative study we chose organic network of different sizes.
-As mentioned each method is applied on a specific model and translation into these models from the PH network or viceversa is done thanks to PINT.
+To evaluate the effectiveness of our new approach,
+we position ourselves with respect to existing methods dealing with different biological models.
+We have chosen the following methods, that are detailed below:
+GINsim (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
+LIBDDD (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
+Rocca's et al. method \cite{roccaasp}
+and PINT \cite{PMR12-MSCS}.
+\annotMF{Il y a une confusion entre méthodes et outils ici je pense.}
+\annotMF{Il faut trouver un autre nom pour “Rocca's method”.}
+Each method use a specific kind of representation:
+logical regulatory network for GINsim,
+Instantiable Transition Systems for LIBDDD,
+state transition network for Rocca's et al. method
+and PH for Pint and our method.
 
+\modMF{%
+For this comparative study, we focus biological network of different sizes:
+a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
+an ERBB receptor-regulated G1/S transition (ERBB) model with 20 components~\cite{Samaga2009}
+and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.
+}%
+These models were chosen to be of different sizes:
+from small (12 components) to large (40 components).
+We note however that the PH models considered may contain more sorts than
+the original number of biological components, due to the use of
+“cooperative sorts”, which allow to model Booelan gates but do not necessarily
+have a biological meaning.
+The different model representations required have been obtained by translations
+from the PH
+ensuring the conservation of the dynamical properties.
+All these translations are available with Pint.
+\todo{URL vers Pint et les autres outils}
+All results alongside with more detailed specifications of the models
+are given in table \ref{tab:reachability}.
+The methods and the results provided by each of them are detailed in the following.
+The overall results shows that our method is efficient in computing reachability
+from a given initial state;
+furthermore, it sometimes provides more informations that the other existing ones.
+
+\begin{figure}[htp]
 \begin{center}
-
-\begin{figure}
 \label{tab:reachability}
-\noindent
+\noindent%
 \begin{tabular}{|l||c|c|c||c|c|c|c|c|}
@@
   \hline
-  tetard \cite{khalis2009smbionet} & 42 & $2^{19}$ & state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
+  TTR & 42 & $2^{19}$ & state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
   \hline
-  ERBB\_G1-S \cite{Samaga2009}  & 152 & $2^{70}$ & state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
+  ERBB & 152 & $2^{70}$ & state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
   \hline
-  ERBB\_G1-S & 152 & $2^{70}$ & process & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
+  ERBB & 152 & $2^{70}$ & process & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
   \hline
-  TCRsig40 \cite{Klamt06} & 156 & $2^{73}$ & process & - & 0m0.014s & out & - & 0m05.02s \\
+  TCR & 156 & $2^{73}$ & process & - & 0m0.014s & out & - & 0m05.02s \\
   \hline
 \end{tabular}
-\caption{Compared performances of Rocca et al.
-method, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
-For each test we find the name of the model, the number of its processes and its states then the target(s).
-The target can be a network state (set of processes) or one or some processes.
-"out" marks an execution asking too much time or memory and "-" marks that is not possible to do the test.
+\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
+For each test, this table gives the considered model,
+the number of its processes and its states, the type of goal
+(either a whole state or a sub-state)
+and the computation time of the different methods used for the tests,
+where “out” marks an execution taking too much time or memory
+and “-” indicates that is not possible to do the test.
 }
+\end{center}
 \end{figure}
-\end{center}
-This table prensents the comparisons run for 4 dynamic analysis methods applied to examples of biological regulatory networks.
-The models of these networks where chosen as being representative of different scales, large networks and small networks.
-This table shows the effectiveness of our method compared to existing. In fact we develop in the sequel that our method provides more compared to other
+
+\todo{Rajouter le nombre de sortes dans le tableau ou la description}
+\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”. Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.}
+
+\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau}
 
@@
 
-\item[--] \textbf{Rocca's method :} It offers the possibility to verify model-checking CTL properties of transition state networks.
-But we note that it is very difficult to write a transition state network in ASP, for example the network of tetard \cite{khalis2009smbionet} with only 42 processes, take 3 minutes to be translated from PH network.
+\item \textbf{Rocca's method} offers the possibility to model-check CTL properties
+of state-transition networks.
+\modMF{%
+We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
+Therefore, using this method first requires an exponential computation to
+express the state-transition graph of the studied model into ASP.
+Indeed, for the tadpole tail resorption model comprising only 42 processes,
+this computation took 3 minutes, which was not included in the solving time
+reported in table~\ref{tab:reachability}.
+Afterwards, this method is able to check any kind of CTL formula
+(and not only the EF form that we focused on in this paper),
+but this checking may be time-consuming.
+}%
 
-\item[--] \textbf{\textsc{GINsim} :} It is an analysis software and simulation of genetic interaction networks.
-It computes the stable states instantly as welle as our ASP method do it instantly.The difference that \textsc{GINsim} does not return all stable states but only reachable ones.
-Regarding the reachability problem, its approach is to calculate all the transition state graph and then check if there is a path between two given states.
-But the large network transition state cannot be displayed.
-In order to determine the path it is necessary to have the entire network transition states and both initial and final states.
-It was indicated in the third line of the table \ref{tab:reachability} that GINsim could not solve the problem because it is not possible to check the reachability of just one or a few components of the network indeed must be given a final state.
+\item \textbf{\textsc{GINsim}} is a software for the edition, simulation and analysis
+of gene interaction networks.
+It allows to compute all reachable stable states instantly;
+however, it is not possible to compute all stable states independently from the initial state.
+Regarding the reachability problem, \textsc{GINSim} only allows to check
+full states, because of its approach which consists in computing
+all the state-transition graph
+\todo{Seulement la partie accessible ou l'ensemble du graphe ?}
+and then check if there exists a path between two given states.
+Therefore, it was not possible to perform reachability checks on partial states.
+Small state-transition graphs can also be displayed by this tool.
 
-\item[--] \textbf{\textsc{libddd} :}
-This is a library for model-checking (CTL, LTL and reachability).
-It returns the answer "True" or "False" after verfying the reachability.
-However our approach is used to return the activation path of the affected components.
-In addition it is clear that \textsc{libddd} put much time to respond (23 times slower to example line 3 of the table).
-That's why for the big examples ($2^{73}$ states) we don't have a response even after 12 minutes but we get it in less than 2 minutes by using the ASP method.
-Also the \textsc{libddd} do not compute the stable states of a network.
+\item \textbf{\textsc{libddd}}
+\modMF{%
+is a library for symbolic model-checking of CTL \& LTL properties.
+It can thus especially be used to check reachability properties;
+however, as opposed to our method, it does not output an execution path
+solving this reachability.
+In addition, it relies on the construction of the state-transition graph
+which is then stored under the form of a binary decision diagram for a more efficient analysis.
+}%
+This computation explains why \textsc{libddd} takes more time to respond,
+and does not answer in less than 12 minutes for the biggest example
+(which contains $2^{73}$ states).
+Finally, \textsc{libddd} is not able to compute the stable states of a network.
 
-\item[--] \textbf{PINT : }It should be noted that the only reachability analysis developed so far on the Process Hitting networks was implemented in the software Pint, and consists in an approximation: it is possible that it terminates but remains inconclusive (although this is rare).
-Moreover, it currently does not give us the path to activate the goal but only the "True" or "False" of the reachability.
+\item \textbf{PINT}
+\modMF{%
+is a library gathering tools and translators related to the PH.
+It should be noted that the only reachability analysis developed so far on the Process Hitting
+was implemented in the software Pint, and consists in an approximation:
+by avoiding to compute the state-transition graph, it is ensured to be really efficient,
+at the cost of possibly terminating without being conclusive (although this is rare).
+}%
+Moreover, in the case of a positive answer,
+it currently does not return the execution the path achieving the desired reachability,
+but only outputs its conclusion.
 In the first example \textsc{PINT} did not give an answer because of the number of goals is greater than 5.
@@
 
-We developped new methods for the dynamic properties, 
-identifying stable states and finding all possible paths to reach states.
-Comparing with other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}) is relatively faster and also permit to edit larger networks (about $2^{73}$ states) but we had also some inconclusif cases.
-In fact the program is inconclusif if the state is not reachable and there is some loops in the graph so the program may turn indefinitely in one of these loops since it is not possible to achieve the goals or to stop the calculation.
-We then proposed in this case to limit the increment to a maximum number that can be reached in the case of an endless loop and which may not be in the case where the targets are verified.
-\begin{tabbing}
- \texttt{ iClingo find\_path.lp <network\_name>.lp --imax=n}
-\end{tabbing}
-Should be given some interesting values for this integer \texttt{"n"}, which sets the maximum number of steps.
-For example, the total number of states is a maximum (it will never be exceeded by a minimum path).
-However, it can reduce the number of sorts if it is assumed that one type will be changed by step (which is often the case of Boolean networks) or the number of shares with the same assumption.
-Note that if in a way the goal is reached after a time \texttt{"n"}, the calculation stops and displays this path as a solution.
+We developed new methods to check dynamical properties,
+namely identifying stable states and finding all possible paths to reach a given goal.
+Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
+our method is relatively faster and also permits to study larger networks
+(up to $2^{73}$ states in our tests) despite some inconclusive cases.
+\modMF{%
+Indeed, our ASP program is non-conclusive in the cases where the given goal is not
+reachable and the model contains loops in its dynamics
+(which happens in almost every biological model).
+In this case, the program will compute all infinite paths in these loops,
+and never reach a goal or a fixed point.
+}%
+It is still possible, however, to limit the number of iterations to an arbitrary
+maximum which will be eventually reached in the case of an endless loop.
+\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
+\begin{lstlisting}[numbers=none]
+iclingo find\_path.lp <network\_name>.lp --imax=n
+\end{lstlisting}
+\annotMF{Cette ligne de commande ne me semble pas nécessaire. Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.}
+Some interesting values can be given to this integer \texttt{n},
+which sets the maximum number of steps.
+For example, the total number of states is a maximum which is obvious,
+as it will never be exceeded by a minimum path,
+but too hight to be very interesting.
+The total number of sorts can also be an interesting value,
+under the hypothesis that each one will change its active process at most once,
+which is often the case for Boolean networks,
+or, by using the same reasoning, the total number of processes.
+\modMF{%
+In this case, however, a termination with no solution cannot be considered as a formal
+negative answer, unless one can prove that the chosen value \texttt{n}
+is bigger than the longest possible path in the state-transition graph.
+Given our implementation, if the step \texttt{n} is reached,
+the computation stops and displays paths that are not valid executions
+to the goal.
+}%

### 2015-01-22 ASP.tex
correction1

@@
 
-This simple language has many advantages that make it one of the most effective in terms of knowledge representation, reasoning and declarative problem solving. 
-\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”}
+This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving. 
+\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”} \repEB{OK}
 
@@
 \annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
-
+\repEB{T’as raison au début j’ai écrit tout le code en mode $math$ et puis j’ai changé juste la partie du code et j’ai oublié de faire la mise à jour de cette partie de définition d’ASP}
 The other classes remain constant as they meet some informal agreement.
@@
 \begin{tabbing}
- \textit{athenian(Socrate) }$\leftarrow ~ \top$. \hspace{2cm} with $\top$ is the symbol of True.
+ \texttt{athenian(Socrate) $\leftarrow ~ \top$. }\hspace{2cm} with \texttt{$\top$}  is the symbol of True.
 \end{tabbing}
@@
 As previously mentioned, a rule has the following form:
-%\begin{equation} 
-%head ~ \leftarrow ~ body.
-%\tag{\eqref{eq1ASP}}
-%\end{equation}
-\begin{equation} \label{eq3ASP}
+\texttt{
+\begin{equation*} \label{eq3ASP}
  A ~ \leftarrow ~ B^+, ~ \ASPnot B^-.
-\end{equation}
-where the part on the left of the arrow is called $head$
+\end{equation*}
+}
+
+where the part on the left of the arrow is called $ head$
 and the right part is called $body$.
 This rule is then equivalent to:
-\begin{equation} \label{eq2ASP}
-L_{0} \leftarrow L_{1},...,L_{m}, \ASPnot L_{m+1},..., \ASPnot L_{n}.
-\end{equation}
+\texttt{
+\begin{equation*} \label{eq2ASP}
+l_{0} \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
+\end{equation*}
+}
 And this rule states:
@@
     \begin{array}{ll}
-        \text{If} ~ L_{1},...,L_{m} ~ \text{are \textbf{true}}\\
-        \text{and} ~ L_{m+1},...,L_{n} ~ \text{are \textbf{false}}
+        \text{If} ~ \texttt{$l_{1},...,l_{m}$}  ~ \text{are \textbf{true}}\\
+        \text{and} ~ \texttt{$l_{m+1},...,l_{n}$} ~ \text{are \textbf{false}}
     \end{array}
-\right \} \text{Then} ~ L_{0} ~ \text{is } \textbf{true} 
+\right \} \text{Then} ~ \texttt{$ l_{0} $} ~ \text{is } \textbf{true} 
 \]
-with $L_{i}$ literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
+with \texttt{$l_{i}$} literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
 Therefore, $head$ is also called the \emph{conclusion} of the rule,
-and $body$ is its \emph{premise}.
+and \texttt{body} is its \emph{premise}.
 
@@
   It can be written without the central arrow ($\leftarrow$) or with a body equal to $\top$:
+  \texttt{
 \begin{equation} 
-L_{0} \leftarrow \top. \qquad \text{or} \qquad L_{0}.
+l_{0} \leftarrow \top. \qquad \text{or} \qquad l_{0}.
  \label{eq5ASP}
 \end{equation}
+}
 
@@
 \begin{equation} 
-  \leftarrow L_{1},...,L_{m}, \ASPnot L_{m+1},..., \ASPnot L_{n}.
+  \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
   \label{eq6ASP}
@@
 
-We say that a set of literals $X$ violates the constraint \eqref{eq6ASP} if $\{L_{1},...,L_{m}\} \subseteq X$ and $\{L_{m+1},...,L_{n}\} \nsubseteq X$.
+We say that a set of literals $X$ violates the constraint \eqref{eq6ASP} if $\{l_{1},...,l_{m}\} \subseteq X$ and $\{l_{m+1},...,l_{n}\} \nsubseteq X$.
 If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.
@@
 with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$).
-The literals $q_i$ can also appear in the body of the rule or in other rules,
+\repEB{ This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$.
+But that the number of this atoms $q_i$ should be between $l$ and $m$.
+That means that it is possible to find the same atom repeated many times at one answer set.}  \\
+In addition the literals $q_i$ can also appear in the body of the rule,
 and $l$ and $u$ are the lower and upper bounds of the number of these literals
@@
 \item it is a simplification of a given system,
-\item it allows an action on the system.
+\item it allows an action on the system \repEB{like resolving problems, searching for a path, identifying states ... }
 \annotMF{Quel genre d'action ?} 
 \end{itemize}
-Models offer the possibility to provide a solution to a problem identified as such.
-\reformuler{This concept model can address another angle issues related to representation process.}
+This concept model can address another angle issues related to representation process.
 
@@
 }%
-the grounder is used to remove variables in order to achieve an equivalent but constant program,
-and the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
+first the grounder is used to remove variables in order to achieve an equivalent but constant program,
+then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
 Amongst all available grounders for ASP, we can name
@@
 }
+\repEB{c'est à ajouter dans le paragraphe ou à reformuler?}
 For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-22 PH.tex
correction1

@@
 
-\todo{Glu pour la définition \ref{def:NextState}}
+\todo{Glu pour la définition \ref{def:NextState}} \repEB{ça veut dire quoi Glu?}
 
@@
 \todo{Définir la notation $s[b_k /b_j ]$}
-\annotMF{Est-ce que $next(s)$ est vraiment utile ? On ne s'en ressert pas après.}
+\annotMF{Est-ce que $next(s)$ est vraiment utile ?
+On ne s'en ressert pas après.} \repEB{Je crois que si, pour le définition du point fixe et pour comprendre l'évolution du réseau d'un état à un autre}
 \end{definition}
@@
 $s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
-\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?}
+\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?} \repEB{Je pense garder fix point}
 \end{definition}
@@
 
-\begin{definition}[Reachability question]
+\begin{definition}[Reachability property]
 \label{def:reachability}
 
- If $s \in \PHl$ is a state and $A \subseteq \Proc$ a set of processes,
- we note $\mathcal{P}(s, A)$ the \emph{the reachability question}:
+ If $s \in \PHl$ is a state and $A \subseteq \Proc$ \repEB{ with $\Proc$ is a set of processes that we find at most one process of each sort $\{a_i, b_j,..., z_k\}$ }
+ we note $\mathcal{P}(s, A)$ the \emph{the reachability property}:
  \begin{center}
@@
 \todo{Il faut définir $\Proc$}
-\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}
+\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}  \repEB{ Je l'ai changé} 
 \end{definition}

### 2015-01-22 comparaison.tex
correction1

@@
 we position ourselves with respect to existing methods dealing with different biological models.
-We have chosen the following methods, that are detailed below:
-GINsim (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
-LIBDDD (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
-Rocca's et al. method \cite{roccaasp}
-and PINT \cite{PMR12-MSCS}.
-\annotMF{Il y a une confusion entre méthodes et outils ici je pense.}
-\annotMF{Il faut trouver un autre nom pour “Rocca's method”.}
+We have chosen the following tools, that are detailed below: 
+\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
+\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al.
+\cite{roccaasp} which was developed ALSO on ASP.
+\annotMF{Il y a une confusion entre méthodes et outils ici je pense.} \repEB{J'ai modifié}
+\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
 Each method use a specific kind of representation:
@@
 state transition network for Rocca's et al. method
-and PH for Pint and our method.
+and Process Hitting (PH) for Pint as well as for our method.
 
@@
 ensuring the conservation of the dynamical properties.
-All these translations are available with Pint.
+All these translations are available with \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models.
+It available online \url{http://loicpauleve.name/pint/} }
 \todo{URL vers Pint et les autres outils}
@@
 \noindent%
-\begin{tabular}{|l||c|c|c||c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
   \hline
-   Model& \#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
+   Model&  \#sorts &\#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
   \hline
-  TTR & 42 & $2^{19}$ & state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
+  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
   \hline
-  ERBB & 152 & $2^{70}$ & state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
+  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
   \hline
-  ERBB & 152 & $2^{70}$ & process & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
+  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
   \hline
-  TCR & 156 & $2^{73}$ & process & - & 0m0.014s & out & - & 0m05.02s \\
+  TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
   \hline
@@
 
-\todo{Rajouter le nombre de sortes dans le tableau ou la description}
-\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”. Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.}
+\todo{Rajouter le nombre de sortes dans le tableau ou la description} \repEB{avec ou sans les sortes coopératives? car pour PH y'en a plus de composants que pour les autres modèles}
+\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”.
+Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.} \modifEB{Je préfère ta proposition mais tu vois ça prend plus d'espace et le tableau dépasse les limites }
 
-\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau}
+\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau} \repEB{En fait pour la méthode de Rocca le script qu'il m'ont envoyé génére quelques erreurs lors de la compilation.
+pour libddd y'avait pas l'option (--fixpoint) ne génère rien comme résultat et GINsim répond très rapidement mais il faudrait calculer tout le graphe de transition avant.
+}
 
@@
 all the state-transition graph
-\todo{Seulement la partie accessible ou l'ensemble du graphe ?}
+\todo{Seulement la partie accessible ou l'ensemble du graphe ?} \repEB{Je pense que oui mais je n'arrive pas à le m'assurer car il m'a fallut les états initiaux pour tester l'atteignabilité à ces états stables}
 and then check if there exists a path between two given states.
@@
 It is still possible, however, to limit the number of iterations to an arbitrary
-maximum which will be eventually reached in the case of an endless loop.
+maximum which will be eventually reached in the case of an endless loop. This is possible with the option \texttt{"--imax=n"}.
 \annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
-\begin{lstlisting}[numbers=none]
-iclingo find\_path.lp <network\_name>.lp --imax=n
-\end{lstlisting}
-\annotMF{Cette ligne de commande ne me semble pas nécessaire. Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.}
+%%
+%\begin{lstlisting}[numbers=none]
+%iclingo find\_path.lp <network\_name>.lp -- imax=n
+%\end{lstlisting}
+%%
+\annotMF{Cette ligne de commande ne me semble pas nécessaire.
+Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.} \modifEB{ok t'as raison je l'ai enlevée et remplacée par juste l'option \texttt{"--imax=n"}}
 Some interesting values can be given to this integer \texttt{n},

### 2015-01-22 intro.tex
correction1

@@
 The objectives of the work presented in this paper are the following.
-Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points \cite{wuensche1998genomic} \annotMF{Est-ce que cette ref est pertinente ?
+Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points \cite{wuensche1998genomic} \annotMF{Est-ce que cette ref est pertinente ?
 Elle me semble plutôt axée sur la notion d'attracteur}).
-For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fixed points.
+\modifEB{Je l’ai mise car en effet je pense que le point fixe est un cas spécial des attracteurs.}
+For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fix points.
 The second phase of our work consists in computing the dynamics.
@@
 We then use ASP in section \ref{fixpoint} to tackle the enumeration
-of all fixed points in a model,
+of all fix points in a model,
 and in section \ref{dynamics} in order to compute the dynamics of a model

### 2015-01-22 reachability.tex
correction1

@@
 \end{lstlisting}
-where \texttt{"a"} is the name of the sort and\texttt{"0"} the index of the active process.
+where \texttt{"a"} is the name of the sort and \texttt{"0"} the index of the active process.
 The dynamics of a network is described by its actions,
@@
 Therefore, we define an ASP predicate \texttt{playableAction(A,I,B,J,K,T)} that is true
-when the processes \texttt{(A,I)} and \texttt{(B,J)} are active at step \texttt{T}.
-\annotMF{Peut-être écrire : $\texttt{A}_\texttt{I}$ et $\texttt{B}_\texttt{J}$}
+when the processes $\texttt{A}_\texttt{I}$ and $\texttt{B}_\texttt{J}$ are active at step \texttt{T}.
+\annotMF{Peut-être écrire : $\texttt{A}_\texttt{I}$ et $\texttt{B}_\texttt{J}$} \repEB{c'est fait}
 It is also needed to enforce the strictly asynchronous dynamic
@@
 by the predicate \texttt{"activeFromTo(B,J,K,T)"}
-which means that in sort \texttt{B}, the active process changes from index \texttt{J} to \texttt{K} between steps \texttt{T} and \texttt{T+1}.
-\annotMF{Peut-être écrire : $\texttt{B}_\texttt{J}$ et $\texttt{B}_\texttt{K}$}
+which means that in sort \texttt{B}, the active process changes from index $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
+\annotMF{Peut-être écrire : $\texttt{B}_\texttt{J}$ et $\texttt{B}_\texttt{K}$} \repEB{c'est fait}
 \modMF{%
@@
 ?
-Sinon, voir à peut-être enlever la cardinalité pour une règle plus simple à comprendre.}
+Sinon, voir à peut-être enlever la cardinalité pour une règle plus simple à comprendre.} \repEB{ En fait non ce n'est pas possible vu que la première est une règle normale et la deuxième est une contrainte.D'ailleurs si je mets comme tu me demande je n'aurai pas le meme résultat.
+En effet la première règle donne l'ensemble de tous les changements possibles pour un instant T  en plus il faut que ça soit un ensemble pour trouver tous les chemins.
+Je peux les enlever pour la clarté mais pour le resultat des deux scripts n'est pas le même}
 
@@
 instate(activeProcess(B,K),T+1) :-  activeFromTo(B,J,K,T), time(T). %\label{e4}
-instate(activeProcess(A,I),T+1) :-  instate(activeProcess(A,I),T),
-          activeFromTo(B,J,K,T), A!=B, time(T). %\label{e5}
+instate(activeProcess(A,I),T+1) :-  instate(activeProcess(A,I),T), %\label{e5}
+          activeFromTo(B,J,K,T), A!=B, time(T). %\label{e5-a}
 \end{lstlisting}
@@
 resulting from the predicate \texttt{activeFromTo} (line \ref{e4})
-as well as all the unchanged processes that correspond to the other sorts (line \ref{e5}).
+as well as all the unchanged processes that correspond to the other sorts (line \ref{e5} and \ref{e5-a}).
 
@@
 \subsection{Reachability verification}
-In this section, we focus on the reachability of set of processes which corresponds to the question (see definition \ref{def:reachability}):
+In this section, we focus on the reachability of set of processes which corresponds to the reachability question (see definition \ref{def:reachability}):
 “Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
 \annotMF{En fin de rédaction, il faudrait voir à peut-être décaler les définitions de la section “Dynamic properties” afin de faciliter la lecture.}
-
+\repEB{Ok}
 We now want to adapt the code of the dynamics computation of previous section in order to resolve this reachability problem.
@@
 \modMF{%
-Then, the literal \texttt{satisfiable(F,T)}
+Then, the literal \texttt{satisfiable(F, T)} 
 checks if a given process \texttt{F} of the goal
 is contained in the state of step \texttt{T},
-as defined in the rule of line \ref{c1}.
+as defined in the rule of line \ref{c2}.
 }%
-Else the answer will be eliminated.
-\todo{Il manque la contrainte qui élimine les mauvais parcours !}
+\modifEB{Else the answer will be eliminated by the constraint \ref{c2-a} which verify if all processes of the goal are satisfied.}
+\todo{Il manque la contrainte qui élimine les mauvais parcours !} \repEB{Je les ai pas mises car ils s'agit de 3 lignes et encore plus d'éxplication et je ne vois pas si c'est important de les mettre car il faudrait les expliquer aussi}
 \begin{lstlisting}
-satisfiable(F,T) :-  goal(F), instate(F,T). %\label{c2}
+satisfiable(F, T) :-  goal(F), instate(F, T). %\label{c2}
+:- X={ satisfaible(F, T+1) : goal(F) }, getNbreGoals(Y), X=Y, time(T). %\label{c2-a}
 \end{lstlisting}
@@
 It is a main disadvantage because a search in $N$ steps will find no solution
-if the shortest path to the goal requires $N+1$ steps.
-\annotMF{Est-ce que ça ne risque pas aussi de rallonger anormalement la résolution si le plus court chemin est en $n$ étapes avec $n << N$ ?}
+if the shortest path to the goal requires $K$ steps with $K > N$. \modifEB{It also may lengthen the resolution if the shortest path requires $n$ steps with $n << N$}
+\annotMF{Est-ce que ça ne risque pas aussi de rallonger anormalement la résolution si le plus court chemin est en $n$ étapes avec $n << N$ ?} \repEB{Si ça risque...}
 One solution is to use an incremental computation mode,
-which is especially tackled by by the incremental solver \textsc{iclingo}~\cite{gebser2008user}.
-\annotMF{Peut-on citer le manuel d'utilisation de Clingo comme référence ?}
+which is especially tackled by the incremental solver \textsc{iclingo}~\cite{gebser2008user}. \modifEB{In fact \textsc{iclingo} further includes the following guidelines:}
+
+\begin{tabbing}
+\texttt{\#base.}  \\
+... (fix code for definitions) ... \\
+\texttt{\#cumulative constant.}  \\
+... (code depending on constant: body iteration) ... \\
+\texttt{\#volatile constant.} \\
+... (code for testing the satisfiyability property : stop computing iteration) ...
+\end{tabbing}
+
+Via \texttt{"\#base."} the subsequent part of a logic program is declared as static, that is, it is processed only once at the beginning of an incremental computation.
+In contrast, \texttt{"\#cumulative constant."} and \texttt{"\#volatile constant."}  are used to declare a (symbolic) constant as a placeholder for incremental step numbers.
+In the following our constant is denoted by \texttt{"t"} indicate the number of a step in the evolution of the network.
+
+\annotMF{Peut-on citer le manuel d'utilisation de Clingo comme référence ?} \repEB{Effectivement la réference \cite{gebser2008user} est un manuel d'utilisation pour Clingo ainsi que IClingo}
 \todo{Expliquer brièvement en quoi consiste la résolution incrémentale en ASP, notamment les trois parties distinctes : définitions, corps d'itération, fin d'itération.}
@@
 except that step numbers \texttt{N}
-are replaced by constant placeholders \texttt{t}
-that will be incremented throughout the execution.
+are replaced by constant placeholders \texttt{"t"}
+that will be incremented throughout the execution (body iteration).
 }%
@@
 and the new states \texttt{instate(activeProcess(A,I),t+1)}
-in the same fashion than previously.
-Regarding the part of local steps,
+in the same way than previously.
+Regarding the part of local steps, \modifEB{that is, all rules, facts, and integrity constraints computed in one step are dismissed before the next incremental step},
 \todo{Définir “the part of local steps”}
-we use a special constraint (line \textbf{c4})
+we use a special constraint (line \ref{c4})
 which states that the program should continue to the next step

### 2015-01-26 ASP.tex
update

@@
 This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving.
-\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”} \repEB{OK}
 
@@
 \annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
-\repEB{T’as raison au début j’ai écrit tout le code en mode $math$ et puis j’ai changé juste la partie du code et j’ai oublié de faire la mise à jour de cette partie de définition d’ASP}
+\repEB{à revoir}
 The other classes remain constant as they meet some informal agreement.
@@
 with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$).
-\repEB{ This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$.
+This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$.
 But that the number of this atoms $q_i$ should be between $l$ and $m$.
-That means that it is possible to find the same atom repeated many times at one answer set.}  \\
+That means that it is possible to find the same atom repeated many times at one answer set.\\
 In addition the literals $q_i$ can also appear in the body of the rule,
@@
 and $|A|$ is the cardinality of the set $A$.
-\modMF{%
 In the following, if they are not explicitly given,
 we consider that $l$ defaults to $0$ and $u$ defaults to $\infty$.
-\annotMF{Les cardinalités ne sont pas seulement utilisées de cette manière ! Dans notre code, on les trouve dans la tête et dans le corps des règles.}
-\todo{À compléter : cardinalités dans le corps des règles}
-}%
+
 \end{itemize}
@@
 \item it is a simplification of a given system,
-\item it allows an action on the system \repEB{like resolving problems, searching for a path, identifying states ... }
-\annotMF{Quel genre d'action ?} 
+\item it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties... 
 \end{itemize}
@@
 \item \textbf{Solving an ASP program}
-\modMF{%
 The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
 that usually work together:
-}%
 first the grounder is used to remove variables in order to achieve an equivalent but constant program,
@@
 }
-\repEB{c'est à ajouter dans le paragraphe ou à reformuler?}
 For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-26 PH.tex
update

--- 2015-01-15 PH.tex
@@
 \pref{def:PH} introduces the Process Hitting~(PH)~\cite{PMR10-TCSB}
-which allows to model
-%A Process Hitting (PH) (\pref{def:PH}) gathers 
-a finite number of local levels,
+which allows to model a finite number of local levels,
 called \emph{processes},
@@
 The concurrent interactions between processes are defined by a set of \emph{actions}.
-Actions describe the replacement of a process by another of the same sort
+Each action is responsible for the replacement of one process by another of the same sort
 conditioned by the presence of at most one other process in the current state.
@@
 A state of the networks is a set of active processes containing a single process of each sort.
-\modMF{%
 The active process of a given sort $a \in \PHs$ in a state $s \in \PHl$
@@
 For any given process $a_i$ we also note: $a_i \in s$ if and only if $\PHget{s}{a} = a_i$.
-}%
 
@@
 
-\begin{definition}[Stable state or fix point]
+\begin{definition}[Stable state or fixed point]
 \label{def:FixPoint1}
@@
 $s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
-\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?}
 \end{definition}
@@
 
-\begin{definition}[Reachability question]
+\begin{definition}[Reachability property]
 \label{def:reachability}
 
- If $s \in \PHl$ is a state and $A \subseteq \Proc$ a set of processes,
- we note $\mathcal{P}(s, A)$ the \emph{the reachability question}:
+ If $s \in \PHl$ is a state and $A \subseteq \Proc$ with $\Proc$ is a set of processes that we find at most one process of each sort $\{a_i, b_j,..., z_k\}$ 
+ we note $\mathcal{P}(s, A)$ the \emph{the reachability property}:
  \begin{center}
@@
 \todo{Il faut définir $\Proc$}
-\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}
 \end{definition}

### 2015-01-26 comparaison.tex
update tableau et abstract
update

@@
 In this section, we show the effectiveness of our approach on some examples,
-\modMF{%
 and compare it to other existing approaches.
-}%
 All computations were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
@@
 \texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al.
-\cite{roccaasp} which was developed ALSO on ASP.
-\annotMF{Il y a une confusion entre méthodes et outils ici je pense.} \repEB{J'ai modifié}
+\cite{roccaasp} which was developed also on ASP.
 \annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
-Each method use a specific kind of representation:
-logical regulatory network for GINsim,
-Instantiable Transition Systems for LIBDDD,
+Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
+logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
+Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
 state transition network for Rocca's et al. method
-and Process Hitting (PH) for Pint as well as for our method.
+and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models.
+It available online \url{http://loicpauleve.name/pint/} } as well as for our method.
 
-\modMF{%
 For this comparative study, we focus biological network of different sizes:
@@
 and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.
-}%
+
 These models were chosen to be of different sizes:
@@
 ensuring the conservation of the dynamical properties.
-All these translations are available with \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models.
-It available online \url{http://loicpauleve.name/pint/} }
-\todo{URL vers Pint et les autres outils}
+All these translations are available with \textsc{PINT} 
 All results alongside with more detailed specifications of the models
@@
   \hline
-  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
+  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
   \hline
-  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
+  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
   \hline
-  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
+  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
+  \hline
+  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & $ \lbrack 5mn ; \infty \rbrack$ & 4m27.93s \\
   \hline
@@
 
-\todo{Rajouter le nombre de sortes dans le tableau ou la description} \repEB{avec ou sans les sortes coopératives? car pour PH y'en a plus de composants que pour les autres modèles}
-\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”.
-Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.} \modifEB{Je préfère ta proposition mais tu vois ça prend plus d'espace et le tableau dépasse les limites }
+\todo{Diviser en 2 tableaux}
 
@@
 of state-transition networks.
-\modMF{%
 We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
@@
 but this checking may be time-consuming.
-}%
 
@@
 \item \textbf{\textsc{libddd}}
-\modMF{%
 is a library for symbolic model-checking of CTL \& LTL properties.
@@
 which is then stored under the form of a binary decision diagram for a more efficient analysis.
-}%
+
 This computation explains why \textsc{libddd} takes more time to respond,
@@
 \item \textbf{PINT}
-\modMF{%
 is a library gathering tools and translators related to the PH.
@@
 at the cost of possibly terminating without being conclusive (although this is rare).
-}%
 Moreover, in the case of a positive answer,
@@
 (up to $2^{73}$ states in our tests) despite some inconclusive cases.
-\modMF{%
 Indeed, our ASP program is non-conclusive in the cases where the given goal is not
@@
 and never reach a goal or a fixed point.
-}%
 It is still possible, however, to limit the number of iterations to an arbitrary
@@
 %%
-\annotMF{Cette ligne de commande ne me semble pas nécessaire.
-Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.} \modifEB{ok t'as raison je l'ai enlevée et remplacée par juste l'option \texttt{"--imax=n"}}
 Some interesting values can be given to this integer \texttt{n},
@@
 or, by using the same reasoning, the total number of processes.
-\modMF{%
 In this case, however, a termination with no solution cannot be considered as a formal
@@
 to the goal.
-}%

### 2015-01-26 fixPoint.tex
correction1
update

@@
-The study of fix points (and, more generally, basins of attraction) provides an important understanding of the different behaviors of a Biological Regulatory Network BRN~\cite{wuensche1998genomic}.
-\modMF{%
+The study of fixed points (and, more generally, basins of attraction) provides an important understanding of the different behaviors of a Biological Regulatory Network BRN~\cite{wuensche1998genomic}.
 Indeed, a system will always eventually end in a basin of attraction,
 and this may depend on biological switch or other complex phenomena.
-}%
 A fixed point is a state of the BRN in which it is not possible any more to have new changes;
@@
 In the following, we consider a Process Hitting $(\Sigma, \PHl, \PHh)$.
-It has been shown that a state $s \in L$ is a fixed point of the Process Hitting if and only if $s$ has no next state~\cite{PMR10-TCSB} \ie if there is no playable action in this state.
-Therefore, a steady state in a Process Hitting network is a state so that
+It has been shown that a state $s \in L$ is a fixed point or a stable state of the Process Hitting if and only if $s$ has no next state~\cite{PMR10-TCSB} \ie if there is no playable action in this state.
+Therefore, a stable state in a Process Hitting network is a state so that
 every process does not hit or is not hit by another process in the same state.
-We note that given this result, processes involved in a self-hit cannot be part of a stable state.
+We note that given this result, processes involved in a self-hit (a process hit its self) cannot be part of a stable state.
 
 \subsection{Process Hitting translation in ASP}
-Before writing any piece of ASP,
-we first need to present the studied PH network in ASP.
+Before writing any piece of scripts on ASP to analyse the dynamic of the network,
+we first need to translat the concerned PH network in ASP  \footnote{You can find our new methods, translation from PH to ASP and the dynamic verification on line: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
+
 To do this we use the following self-describing predicates:
-\texttt{sort}, \texttt{process} and \texttt{action}.
-\annotMF{Décrire rapidement le sens de ces prédicats ?}
+\texttt{"sort"} to define sorts, \texttt{"process"} for the processes and \texttt{"action"} for the network actions.
+We will see in example \ref{ex:asp-ph} how a PH network is defined with these predicates.
+We note that all the ASP scripts are
 
@@
 \label{ex:asp-ph}
-The representation of the PH network of Figure \ref{fig:ph} in ASP is the following:
+The representation of the PH network of figure \ref{fig:ph} in ASP is the following:
 \begin{lstlisting}
-sort("a"). sort("b"). sort("z"). %\label{ASPsort}
+sort(X) :- process(X,) %\label{ASPsort}
 process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
@@
 
-\subsection{Search of fix points}
+\subsection{Search of fixed points}
 \todo{ Parler du point fixe et ses interrêts. \\
@@
 
-The enumeration of fix points requires to translate the definition of a stable state
+The enumeration of fixed points requires to translate the definition of a stable state
 in a set of ASP rules.
 The first step consists of eliminating all processes involved in a self-hit;
-the others are recorded by the predicate \texttt{shownProcess}:
+the others are recorded by the predicate \texttt{"shownProcess"}:
 \begin{lstlisting}
@@
 \end{lstlisting}
-\modMF{%
 The previous line creates an many candidate answer sets as there are possible states
@@
 indeed, any solution that satisfy the body of a constraint will be removed from the answer set.
-}%
 Regarding our problem, a state is eliminated if there exists an action linking two processes:
@@
 We can check that this model contains only one fixed point: $\PHstate{b_0, z_2, a_0}$.
-Indeed, there is no action between each two of the processes contained is this state;
-moreover, no other state verifies this property.
+Indeed, there is no action between each two of the processes contained is this state so no evolution is possible from this state. 
+For this example no other state verifies this property.
 
 If we execute the ASP program detailed below,
-alongside with the description of the PS model given in example~\ref{ex:asp-ph},
+alongside with the description of the PH model given in example~\ref{ex:asp-ph},
 we obtain one answer set, which matches the expected result:

### 2015-01-26 intro.tex
update

@@
 A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.
-This particular structure makes the formal analysis of BRNs with hundreds of components tractable \modMF{by using approached methods computing approximations of the dynamics~\cite{PMR12-MSCS}}.
+This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using approached methods computing approximations of the dynamics~\cite{PMR12-MSCS}.
 Furthermore, PH is suitable, according to the precision of the available information, to model BRNs with different levels of abstraction by capturing the most general dynamics.
@@
 The objectives of the work presented in this paper are the following.
-Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points \cite{wuensche1998genomic} \annotMF{Est-ce que cette ref est pertinente ?
-Elle me semble plutôt axée sur la notion d'attracteur}).
-\modifEB{Je l’ai mise car en effet je pense que le point fixe est un cas spécial des attracteurs.}
+Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points which is a specific attractor with only one state \cite{wuensche1998genomic}).
 For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fix points.
@@
 
-\modMF{%
 Section \ref{defs} defines the paradigm of Answer Set Programming,
@@
 and in section \ref{ccl} we conclude and talk about future work.
-}%

### 2015-01-26 reachability.tex
update

@@
 \subsection{Future states identification}
-\reformuler{From an initial known state, a PH network can evolve into several new states after a few steps.}
-\modMF{%
+Thanks to its actions, a PH network is considered like a dynamic model or a dynamic biological network.
+In fact starting from an initial state the network can envolve into several new states.
+Our objectif is to try to predict the next new states as well as to  see if other states are reachable and if it's the case to determine the path to reach them.
 Firstly, we focus on the evolutions of the model in a limited number of steps.
-}%
-The predicate \texttt{time(0..n)} sets the number of steps we want to play.
-\modMF{%
+So we define the predicate \texttt{time(0..n)} which sets the number of steps we want to play.
 The value of \texttt{n} can be arbitrarily chosen.
@@
 to list the processes included in this state:
-}%
 \begin{lstlisting}
@@
 when the processes $\texttt{A}_\texttt{I}$ and $\texttt{B}_\texttt{J}$ are active at step \texttt{T}.
-\annotMF{Peut-être écrire : $\texttt{A}_\texttt{I}$ et $\texttt{B}_\texttt{J}$} \repEB{c'est fait}
 It is also needed to enforce the strictly asynchronous dynamic
@@
 which means that in sort \texttt{B}, the active process changes from index $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
-\annotMF{Peut-être écrire : $\texttt{B}_\texttt{J}$ et $\texttt{B}_\texttt{K}$} \repEB{c'est fait}
-\modMF{%
 The cardinality rule of line \ref{e2} then
@@
 the asynchronous dynamics of the PH.
-}%
 \begin{lstlisting}
@@
 \end{lstlisting}
-\annotMF{N'est-il pas possible de réunir ces deux règles avec des bornes 1--1 de la façon suivante : 1 \{activeFromTo(B,J,K,T)\} 1 $\leftarrow$ ...
-?
-Sinon, voir à peut-être enlever la cardinalité pour une règle plus simple à comprendre.} \repEB{ En fait non ce n'est pas possible vu que la première est une règle normale et la deuxième est une contrainte.D'ailleurs si je mets comme tu me demande je n'aurai pas le meme résultat.
-En effet la première règle donne l'ensemble de tous les changements possibles pour un instant T  en plus il faut que ça soit un ensemble pour trouver tous les chemins.
-Je peux les enlever pour la clarté mais pour le resultat des deux scripts n'est pas le même}
 
@@
 
-\modMF{%
 The overall result of the pieces of program presented in this subsection
@@
 and starting from a given initial state.
-}%
 
@@
 \end{lstlisting}
-\modMF{%
 Then, the literal \texttt{satisfiable(F, T)} 
@@
 as defined in the rule of line \ref{c2}.
-}%
-\modifEB{Else the answer will be eliminated by the constraint \ref{c2-a} which verify if all processes of the goal are satisfied.}
-\todo{Il manque la contrainte qui élimine les mauvais parcours !} \repEB{Je les ai pas mises car ils s'agit de 3 lignes et encore plus d'éxplication et je ne vois pas si c'est important de les mettre car il faudrait les expliquer aussi}
+Else the answer will be eliminated by the constraint \ref{c2-a} which verifies if all processes of the goal are satisfied.
+\todo{Détailler la contraite 21 ou l'enlever} 
 \begin{lstlisting}
@@
 It is a main disadvantage because a search in $N$ steps will find no solution
-if the shortest path to the goal requires $K$ steps with $K > N$. \modifEB{It also may lengthen the resolution if the shortest path requires $n$ steps with $n << N$}
-\annotMF{Est-ce que ça ne risque pas aussi de rallonger anormalement la résolution si le plus court chemin est en $n$ étapes avec $n << N$ ?} \repEB{Si ça risque...}
+if the shortest path to the goal requires $K$ steps with $K > N$. It also may lengthen the resolution if the shortest path requires $n$ steps with $n << N$.
 One solution is to use an incremental computation mode,
-which is especially tackled by the incremental solver \textsc{iclingo}~\cite{gebser2008user}. \modifEB{In fact \textsc{iclingo} further includes the following guidelines:}
+which is especially tackled by the incremental solver \textsc{iclingo}~\cite{gebser2008user}. In fact \textsc{iclingo} further includes the following guidelines:
 
@@
 \texttt{\#base.}  \\
-... (fix code for definitions) ... \\
+... (static code for definitions) ... \\
 \texttt{\#cumulative constant.}  \\
@@
 In the following our constant is denoted by \texttt{"t"} indicate the number of a step in the evolution of the network.
-
-\annotMF{Peut-on citer le manuel d'utilisation de Clingo comme référence ?} \repEB{Effectivement la réference \cite{gebser2008user} est un manuel d'utilisation pour Clingo ainsi que IClingo}
-\todo{Expliquer brièvement en quoi consiste la résolution incrémentale en ASP, notamment les trois parties distinctes : définitions, corps d'itération, fin d'itération.}
-\modMF{%
 Thus, the obtained program is almost identical,
@@
 that will be incremented throughout the execution (body iteration).
-}%
 In each step \texttt{t}, the program computes the playable actions \texttt{playableAction(A,I,B,J,K,t)}, the possible bounces \texttt{activeFromTo(B,J,K,t-1)}
@@
 in the same way than previously.
-Regarding the part of local steps, \modifEB{that is, all rules, facts, and integrity constraints computed in one step are dismissed before the next incremental step},
-\todo{Définir “the part of local steps”}
+Regarding the part of local steps, that is, all rules, facts, and integrity constraints computed in one step are dismissed before the next incremental step,
 we use a special constraint (line \ref{c4})

### 2015-01-26 PH.tex
revise dynamical properties of PH

@@
 that is, of the form: $\PHfrappe{a_i}{a_i}{a_k}$.
+
+\modMF{%
+The PH is therefore a restriction of synchronous automata, where each transition
+changes the local state of exactly one automaton,
+and is triggered by the local states of at most two distinct automata.
+This restriction in the form of the actions was chosen to permit
+the development of efficient static analysis methods
+based on abstract interpretation~\cite{PMR12-MSCS}.
+}%
 
@@
 Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
-We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$ is \emph{playable in state $s$} if and only if $a_i \in s$ and $b_j \in s$ (\ie $\PHget{s}{a} = a_i$ and $\PHget{s}{b}=b_j$).
+We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$
+is \emph{playable in state $s$} if and only if
+$a_i \in s$ and $b_j \in s$ (\ie $\PHget{s}{a} = a_i$ and $\PHget{s}{b}=b_j$).
+The resulting state after playing $h$ in $s$
+is called a \emph{successor} of $s$ and
+is denoted by $(s \play h)$,
+where $\PHget{(s \play h)}{b} = b_k$ and
+$\forall c \in \PHs, c \neq b \Rightarrow \PHget{(s \play h)}{c}=\PHget{s}{c}$.
 \end{definition}
-The resulting state after playing an action $h$ in a state $s$ where it is playable is denoted by $(s \play h)$, where $\PHget{(s \play h)}{b} = b_k$ and $\forall c \in \PHs, c \neq b, \PHget{(s \play h)}{c}=\PHget{s}{c}$.
 
-\subsection*{Dynamic properties}
+\subsection*{Dynamical properties}
 
 The study of the dynamics of biological networks was the focus of many works, explaining the diversity of network modelings and the different methods developed in order to check dynamic properties.
-In this paper we focus on 2 main properties: the stable states and the reachability.
-We formally define these properties in the following and explain how they could be verified in a PH network.
+In this paper we focus on 2 main properties: the stable states and the reachability,
+which were already formalized and tackled by other methods in~\cite{PMR10-TCSB,PMR12-MSCS}.
+In the following, we consider a PH model $(\PHs,\PHl,\PHa)$,
+and we formally define these properties
+and explain how they could be verified on such a network.
 
-The dynamics of a network or its evolution is specified by its actions.
-Indeed, when an action $h = \PHfrappe{a_i}{b_j}{b_k}$ is played at a state $s$,
-the network evolves to a resulting next state $s'$ where exactly one active process
-is modified, namely: the active process $b_j$ changes into $b_k$.
+The notion of \emph{fixed point}, also called \emph{stable state},
+is given in definition~\ref{def:fixpoint}.
+A fixed point is a state which has no successor.
+Such states have a particular interest as they denote states in which the model
+stays indefinitely,
+and the existence of several of these states denotes a switch in the dynamics.
 
-\todo{Glu pour la définition \ref{def:NextState}}
-
-\begin{definition}[Next states]
-\label{def:NextState}
-Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
-one of its states. The set of the next possible states for $s$ are computed as follows:
-\begin{center}
-$next(s) = \{s[b_k /b_j ]| \exists (a_i , b_j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa\}$
-\end{center}
-\todo{Définir la notation $s[b_k /b_j ]$}
-\annotMF{Est-ce que $next(s)$ est vraiment utile ?
-On ne s'en ressert pas après.} \repEB{Je crois que si, pour le définition du point fixe et pour comprendre l'évolution du réseau d'un état à un autre}
+\begin{definition}[Fixed point]
+\label{def:fixpoint}
+  A state $s \in \PHl$ is called a \emph{fixed point}
+  (or equivalently \emph{stable state})
+  if and only if is has no successors.
+  In other words, $s$ is a fixed point if an only if no action is playable in this state:
+  \[\forall \PHfrappe{a_i}{b_j}{b_k} \in \PHa, a_i \notin s \vee b_j \notin s \enspace.\]
 \end{definition}
 
-A sequence of successively playable actions ($h_0 \play h_1 \play h_2...$)
-from an initial state $s_0$ is called a \emph{scenario in $s_0$}.
-We note $\Sce(s_0)$ the set of all scenarios in $s_0$.
+A finer and more interesting dynamical property consists in
+the notion of \emph{reachability property}.
+Such a property, defined in definition~\ref{def:reachability},
+states that starting from a given initial state, it is possible
+to reach a given goal, that is, a state that contains a process
+or a set of processes.
+Checking such a dynamical property is considered difficult
+as in usual model-checking techniques,
+it is required to build (a part of) the state graph,
+which has an exponential complexity.
 
-\todo{Glu pour la définition \ref{def:FixPoint1} et le lemme \ref{def:FixPoint2}}
-
-\begin{definition}[Stable state or fixed point]
-\label{def:FixPoint1}
-Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
-$s \in \PHl$ be state;
-$s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
-\end{definition}
-
-\begin{definition}
-[Lemma: Stable state or fix point]
-\label{def:FixPoint2}
-Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
-$s \in \PHl$ be a state;
-$s$ is a stable state for $PH$ if and only if there is no playable action in state $s$: 
-\begin{center}
-$\forall h \in \PHa$, $frappeur(h) \notin s \vee cible(h) \notin s$
-\end{center}
-\end{definition}
-
-\todo{Glu pour la définition \ref{def:reachability}}
+In the following, if $s \in \PHl$ is a state,
+we call \emph{scenario in $s$}
+any sequence of actions that are successively playable in $s$.
+We also note $\Sce(s)$ the set of all scenarios in $s$.
+Moreover, we denote by $\Proc = \bigcup_{a \in \PHs} \PHl_a$
+the set of all process in $\PH$.
 
@@
 \label{def:reachability}
+  If $s \in \PHl$ is a state and $A \subseteq \Proc$ is a set of processes,
+  we denote by $\mathcal{P}(s, A)$ the following \emph{reachability property}:
+  \[\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a} = a_i
+    \enspace.\]
+\end{definition}
 
- If $s \in \PHl$ is a state and $A \subseteq \Proc$ with $\Proc$ is a set of processes that we find at most one process of each sort $\{a_i, b_j,..., z_k\}$ 
- we note $\mathcal{P}(s, A)$ the \emph{the reachability property}:
- \begin{center}
-  $\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$
- \end{center}
-\todo{Il faut définir $\Proc$}
-\end{definition}
+The rest of the paper focuses on the resolution of the previous issues
+with the use of ASP.
+The enumeration of all fixed points of a PH model will be tackled in
+section~\ref{fixpoint}
+and the verification of a reachability property will be the subject of
+section~\ref{dynamics}.

### 2015-01-27 intro.tex
Add model-checking oriented introduction

@@
 \todo{à revoir} \\
+For more than a decade, model-checking and SAT-solvers developments have been strongly connected \cite{biere1999symbolic}.
+Among different reasons, the main issue is related to the different problems with BDD-based symbolic approaches, which are:
+\begin{itemize}
+\item BDD-based structures require some canonical form to be defined so that a total order can be defined; 
+\item They often become too large;
+\item Variable ordering is crucial to the performances of the approaches.  
+\end{itemize}
+
+Behind the use of SAT (respectively ASP) to perform model-checking, there is the intuitive idea that the existence of a trace satisfying a given property expressed in LTL \cite{biere1999symbolic} or CTL (first reduced to universal properties \cite{penczek2002bounded}, then extended to other properties) can be reduced to the satisfiability of an equivalent logic formula (respectively a logic program).
+Although there have been some works in the field of unbounded model-checking \cite{mcmillan2002applying}, it generally requires to consider bounded traces, which gave birth to the concept of Bounded Model-Checking (BMC).
+Its principle is following: given a discrete model $M$ (which, equipped with a semantics, leads to a transition system), a property $\varphi$ and $k \in \mathbb{N}$, does $M$ allow a counterexample to $\varphi$ of $k$ (or fewer) transitions?
+
+Bounded model-checking has been the subject of numerous researches.
+On the one hand, it also suffers from some limitations.
+First, it leads to incomplete analysis: if the SAT solver proves that the problem given as input is unsatisfiable for a length $k$, it only proves there are no counterexamples of length $k$.
+Finding bounds on the length is difficult and worst case is exponential.
+On the other hand, SAT is a viable alternative to BDD-based symbolic model-checking.
+It is also an efficient approach for debugging, i.e., quickly find counterexamples of minimal length.
+
+Recently, the rise of ASP has also given birth to some innovative approaches, taking profit of the performances of ASP to process efficient model-checking algorithms.
+In \cite{rocca2013inference}, the authors proposed an approach to perform LTL and CTL model-checking through ASP, with the goal to infer and learn (synchronous) Boolean networks.
+This method has been refined in the context of (asynchronous) Thomas' biological regulatory networks \cite{roccaasp}.
+
 As regulatory phenomena play a crucial role in biological systems, they need to

### 2015-01-27 intro.tex
minor corrections to the introduction

@@
-\todo{à revoir} \\
 For more than a decade, model-checking and SAT-solvers developments have been strongly connected \cite{biere1999symbolic}.
-Among different reasons, the main issue is related to the different problems with BDD-based symbolic approaches, which are:
+Among different reasons, the main issue is related to the different problems with BDD-based symbolic approaches, that is:
 \begin{itemize}
-\item BDD-based structures require some canonical form to be defined so that a total order can be defined; 
+\item BDD-based structures require some canonical form to be given so that a total order can be defined; 
 \item They often become too large;
@@
 On the other hand, SAT is a viable alternative to BDD-based symbolic model-checking.
-It is also an efficient approach for debugging, i.e., quickly find counterexamples of minimal length.
+It is also an efficient approach for debugging, \ie quickly find counterexamples of minimal length.
 
@@
 The objectives of the work presented in this paper are the following.
-Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points which is a specific attractor with only one state \cite{wuensche1998genomic}).
-For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fix points.
-The second phase of our work consists in computing the dynamics.
-It consists in determining from a known initial state the possible next states of the PH model.
-Then we verify if we can reach a specific state for one or several component (genes or proteins).
+Firstly, we show that starting from a PH model, it is possible to find all possible
+fixed points (also called stable states)
+which are specific attractors with only one state~\cite{wuensche1998genomic}).
+For this, we perform an exhaustive search of the possible states and then check which ones are fixed points.
+The second phase of our work consists in computing the dynamics by determining, from a given initial state, the possible next states of the PH model.
+This also permits to check if a goal is reachable for a given component or set of components in the system.
 The results are ensured to respect strictly the PH dynamics.
@@
 We then use ASP in section \ref{fixpoint} to tackle the enumeration
-of all fix points in a model,
+of all fixed points in a model,
 and in section \ref{dynamics} in order to compute the dynamics of a model

### 2015-01-27 comparaison.tex
improvements to section 'Comparison' (Pint + limitations)

@@
 \annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
+\annotMF{Je propose plutôt : “ASP for CTL model-checking” qui pourrait se simplifier en “ASP-CTL” dans le tableau.}
 Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
@@
 is a library gathering tools and translators related to the PH.
-It should be noted that the only reachability analysis developed so far on the Process Hitting
-was implemented in the software Pint, and consists in an approximation:
-by avoiding to compute the state-transition graph, it is ensured to be really efficient,
-at the cost of possibly terminating without being conclusive (although this is rare).
+It should be noted that Pint contains the only reachability analysis
+developed so far natively for the Process Hitting,
+before the method proposed in this paper.
+It consists in an approximation that avoids to compute the state-transition graph;
+It is thus ensured to be really efficient, which explains the fastest results,
+but at the cost of possibly terminating without being conclusive (although this is rare).
+However, it is not designed for goals consisting of a full state, which
+may trigger an exponential process to reach an answer,
+which explains the high computing times for some of them.
+\todo{Ref vers le tableau}
 Moreover, in the case of a positive answer,
@@
 but only outputs its conclusion.
-In the first example \textsc{PINT} did not give an answer because of the number of goals is greater than 5.
 \end{itemize}
 
-\subsection{Limitation}
+\subsection{Strengths and limitations of our method}
 
-We developed new methods to check dynamical properties,
+In the previous sections,
+we developed new methods to check dynamical properties,
 namely identifying stable states and finding all possible paths to reach a given goal.
@@
 It is still possible, however, to limit the number of iterations to an arbitrary
-maximum which will be eventually reached in the case of an endless loop. This is possible with the option \texttt{"--imax=n"}.
-\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
+maximum which will be eventually reached in the case of an endless loop.
+This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iclingo},
+where \texttt{n} is the maximum number of steps.
 %%
@@
 %%
-Some interesting values can be given to this integer \texttt{n},
-which sets the maximum number of steps.
-For example, the total number of states is a maximum which is obvious,
+For example, the total number of states is an obvious maximum,
 as it will never be exceeded by a minimum path,
-but too hight to be very interesting.
-The total number of sorts can also be an interesting value,
+but it is too hight to be very interesting.
+The total number of sorts is a more interesting value,
 under the hypothesis that each one will change its active process at most once,
-which is often the case for Boolean networks,
-or, by using the same reasoning, the total number of processes.
+which is often the case for Boolean networks;
+or, with a similar reasoning, the total number of processes can be chosen.
 In this case, however, a termination with no solution cannot be considered as a formal
@@
 to the goal.
+\todo{Y a-t-il moyen de les différencier de chemins corrects?}

### 2015-01-28 ASP.tex
improvement of ASP + update

@@
 This paradigm allows the programmer to specify what the problem to be solved is, and not how to solve it.
-ASP programs are written in AnsProlog* (short for "Answer Set Programming in Logic" with superscript:~*).
+ASP programs are written in AnsProlog* \cite{sureshkumar2006ansprolog} (short for "Answer Set Programming in Logic" with superscript:~*).
 These programs are composed of a set of \emph{facts} and a set of \emph{rules}, as defined below, from which other facts can be derived.
@@
 
-\subsubsection{Logic program}
-An ASP program is composed of rules, each one being an ordered pair~\cite{baral2003knowledge}:
+\subsubsection{The syntax and semantic of ASP}
+\label{sectionSyntaxeASP}
+\begin{enumerate}
+
+\item \textbf{The rules:}
+As previously mentioned, a rule has the following form:
+\texttt{
 \begin{equation} \label{eq1ASP}
@@
 \end{equation}
-with $head$ and $body$ two sets of literals.
-
-This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving.
-
-\subsubsection{The syntax and semantic of ASP}
-\label{sectionSyntaxeASP}
-\begin{enumerate}
-\item \textbf{The alphabet}
-
-\noindent{ According to \cite{baral2003knowledge}, the alphabet of the axioms (or simply called the \emph{alphabet}) of ASP is composed of seven classes of symbols:
-\begin{itemize}
-\item Variables,
-\item Constants,
-\item Symbols of functions,
-\item Symbols of predicates,
-\item Connectors,	
-\item Symbols of punctuation,
-\item The special symbol $\bot$
-\end{itemize}
 }
-All these classes vary from one alphabet to another, but the sets of $ 5^{th} $ and $ 6^{th} $ classes (Connectors and symbols of punctuation) are defined as follows:
-\begin{itemize}
-  \item Connectors with \{$\neg$, or, $\leftarrow$, not, ‘,’\}
-  \item Symbols of punctuation with \{‘(’, ‘)’, ‘.’\}
-\end{itemize}
-
-\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
-\repEB{à revoir}
-The other classes remain constant as they meet some informal agreement.
-In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
-The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
-String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).
-
-\annotMF{Je ne suis pas sûr que l'explication sur les prédicats soit pertinente. Si on la garde, il faudrait la reformuler.}
-What seems a little fuzzy is the concept of predicate. Indeed the word predicate can be an innovation of the new grammar. Let's consider the following sentence: 
-\begin{tabbing}
- \textit{Socrates was an athenian.} 
-\end{tabbing}
-\textit{Socrate} is the subject and \textit{was an athenian} is the predicat. This sentence can be noted in ASP by :
-\begin{tabbing}
- \texttt{athenian(Socrate) $\leftarrow ~ \top$. }\hspace{2cm} with \texttt{$\top$}  is the symbol of True.
-\end{tabbing}
-
-\item \textbf{The rules}
-As previously mentioned, a rule has the following form:
-\texttt{
-\begin{equation*} \label{eq3ASP}
- A ~ \leftarrow ~ B^+, ~ \ASPnot B^-.
-\end{equation*}
-}
-
-where the part on the left of the arrow is called $ head$
-and the right part is called $body$.
-This rule is then equivalent to:
-\texttt{
-\begin{equation*} \label{eq2ASP}
+This rule is equivalent to:
+\begin{equation} \label{eq2ASP}
 l_{0} \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
-\end{equation*}
-}
+\end{equation}
 And this rule states:
@@
     \begin{array}{ll}
-        \text{If} ~ \texttt{$l_{1},...,l_{m}$}  ~ \text{are \textbf{true}}\\
-        \text{and} ~ \texttt{$l_{m+1},...,l_{n}$} ~ \text{are \textbf{false}}
+        \text{If} ~ l_{1},...,l_{m}  ~ \text{are \textbf{true}}\\
+        \text{and} ~ l_{m+1},...,l_{n} ~ \text{are \textbf{false}}
     \end{array}
-\right \} \text{Then} ~ \texttt{$ l_{0} $} ~ \text{is } \textbf{true} 
+\right \} \text{Then} ~ l_{0} ~ \text{is } \textbf{true} 
 \]
-with \texttt{$l_{i}$} literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
-Therefore, $head$ is also called the \emph{conclusion} of the rule,
-and \texttt{body} is its \emph{premise}.
+with $l_{i}$ literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
+Therefore, "$head$" is also called the \emph{conclusion} of the rule,
+and "$body$" is its \emph{premise}.
 
@@
 \begin{itemize}
-\item A rule is said \textbf{constant} or \emph{ground} if all of its literals are constants;
+\item A  \textbf{ground} rule, is a rule  where all its literals are constants;
 \item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
-  It can be written without the central arrow ($\leftarrow$) or with a body equal to $\top$:
-  \texttt{
+  It can be written without the central arrow ($\leftarrow$) or with a $body$ equals to $\top$, and $\top$  is the symbol of True.
 \begin{equation} 
@@
 \end{equation}
-}
 
@@
 
+\item \textbf{The alphabet}
+
+According to \cite{baral2003knowledge}, the alphabet of ASP is composed of seven classes of symbols: Variables, start with a capital letter.
+Constants, symbols of functions, symbols of predicates, these 3 classes follow the same rule but they begin with lowercase letter.
+Connectors: $\neg$, or, $\leftarrow$, not, ‘,’.
+Symbols of punctuation :(‘(’, ‘)’, ‘.’) and finally the special symbols : $\bot$ and $\top$.
+
+The word predicate can be an innovation of the new grammar. Let's consider the following sentence to more explain it: 
+\begin{tabbing}
+ \textit{tweety is a bird.} 
+\end{tabbing}
+"\textit{tweety}" is a constant and "\textit{is a bird}" is a predicat. This sentence can be noted in ASP by :
+\begin{tabbing}
+ bird(tweety) $\leftarrow ~ \top$. \hspace{1cm} or \hspace{2cm} $bird(tweety) .$
+\end{tabbing} 
 \end{enumerate}
 
-\subsubsection{Modeling of a problem with ASP}
+\subsubsection{Modeling and solving of a problem with ASP: }
 
@@
 It concerns all systems we seek to control.
-A model has two main features \cite{Glimpse}:
-\begin{itemize}
-\item it is a simplification of a given system,
-\item it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties... 
-\end{itemize}
+A model has two main features \cite{Glimpse}: In fact it is a simplification of a given system and it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties...
 This concept model can address another angle issues related to representation process.
+Logical ASP programs follow the strategy of “generate and test”.
+Indeed, the modeling process can be regarded as a special form of representation.
+First, the system should be presented  with facts.
+Second it have to be explain with rules.
+Third, it should be possible to generate all possibilities using cardinalities.
+Finally, a filter will be done thanking to the constraints and only the sets of predicates satisfying all these conditions will be displayed which are considered "answer sets".
 
-\begin{enumerate}
-\item\textbf{Modeling steps}
-The modeling process can be regarded as a special form of representation whose operations are detailed in \cite{baral2003knowledge}.
-Logical ASP programs follow the strategy of “generate and test”.
-This strategy includes four steps:
-\begin{itemize}
-\item Enumerate with facts;
-\item Explain with rules;
-\item Generate all the possibilities with cardinalities; and finally,
-\item Filter with constraints.
-\end{itemize}
-
-\item \textbf{Solving an ASP program}
-The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
+We denote that the solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
 that usually work together:
@@
 
-\reformuler{
-The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations.
-Both systems employ grounding algorithms that work quickly and simplify the program.
-}
-For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.
-
-\end{enumerate}
+For the work presented in this paper, we worked with \texttt{Clingo} (verion 3) which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-28 PH.tex
improvement of ASP + update

@@
 
-\modMF{%
 The PH is therefore a restriction of synchronous automata, where each transition
@@
 based on abstract interpretation~\cite{PMR12-MSCS}.
-}%
 

### 2015-01-28 comparaison.tex
improvement of ASP + update

@@
 \textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
-\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al.
-\cite{roccaasp} which was developed also on ASP.
-\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
-\annotMF{Je propose plutôt : “ASP for CTL model-checking” qui pourrait se simplifier en “ASP-CTL” dans le tableau.}
-Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
+\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method for CTL model-checking of Rocca's et al.
+\cite{roccaasp} which was developed also on ASP but for transition states networks.
+Each method use a specific kind of representation\footnote{The transalation from PH to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
 logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
 Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
-state transition network for Rocca's et al. method
+state transitions network for Rocca's et al. method
 and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models.
@@
   \hline
-   Model&  \#sorts &\#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
+   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-CTL} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
   \hline
@@
   \hline
-  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & $ \lbrack 5mn ; \infty \rbrack$ & 4m27.93s \\
+  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & out & 4m27.93s \\
   \hline
@@
 \end{tabular}
-\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
+\caption{Compared performances of Rocca et al.
+method denoted by \textsc{ASP-CTL}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
 For each test, this table gives the considered model,
@@
 full states, because of its approach which consists in computing
-all the state-transition graph
-\todo{Seulement la partie accessible ou l'ensemble du graphe ?} \repEB{Je pense que oui mais je n'arrive pas à le m'assurer car il m'a fallut les états initiaux pour tester l'atteignabilité à ces états stables}
-and then check if there exists a path between two given states.
+all the state-transition graph and then check if there exists a path between two given states.
 Therefore, it was not possible to perform reachability checks on partial states.
@@
 may trigger an exponential process to reach an answer,
-which explains the high computing times for some of them.
-\todo{Ref vers le tableau}
+which explains the high computing times for some of them \ref{tab:reachability}
 Moreover, in the case of a positive answer,

### 2015-01-28 intro.tex
improvement of ASP + update

--- 2015-01-26 intro.tex
@@
-\todo{à revoir} \\
 As regulatory phenomena play a crucial role in biological systems, they need to
@@
 The objectives of the work presented in this paper are the following.
-Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points which is a specific attractor with only one state \cite{wuensche1998genomic}).
-For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fix points.
+Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points which is a specific attractor with only one state \cite{wuensche1998genomic}).
+For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fixed points.
 The second phase of our work consists in computing the dynamics.
@@
 The benefit of our approach is that it makes possible to get the minimal paths to reach our goal(s) also we can verify if it is possible after a given number of steps.
-%Furthermore, we claim that we are able to deal with large biological networks (42 sorts).
-
-\annotMF{Je pense que dans toute l'intro, il faut remettre ASP au centre car c'est bien lui qui rend notre contribution particulière.
-Il ne faudrait pas séparer Objectifs/Contributions car les deux se recoupent, ou trouver un moyen de parler d'ASP avant de présenter le travail sur les points fixes et la dynamique.}
 

### 2015-01-28 reachability.tex
improvement of ASP + update

@@
 “Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
-\annotMF{En fin de rédaction, il faudrait voir à peut-être décaler les définitions de la section “Dynamic properties” afin de faciliter la lecture.}
-\repEB{Ok}
 We now want to adapt the code of the dynamics computation of previous section in order to resolve this reachability problem.

### 2015-01-28 cc.tex
conclusion

In this paper, we gave a new method to compute some dynamical properties
on Process Hitting models, a subclass of synchronous automata.
The main interest of our method is the use of Answer Set Programming,
a declarative programming paradigm which benefits from powerful solvers.
We first focused on the enumeration of the fixed points of a model,
which is tackled simply on such models given their particular form.
We also considered the reachability problem, that is,
checking if it is possible to reach a state with a given property
from a given initial state,
which thus corresponds to an $\mathsf{EF}$ operator in CTL logic.
Our analysis is thus exhaustive, but can be limited to a number of steps,
for which the dynamics of the model from the given initial state is computed.
We gave an implementation of these problems into Answer Set Programming,
and applied them to several biological examples of various sizes, up to
40 biological components.
Our results showed that our implementation is faster and deals with bigger models
than other approaches, especially \textsc{LibDDD} which is a symbolic model-checker.

Our work could benefit from several extension.
Of course, the set of applicable models can be extended,
for example with the addition
of priorities or neutralizing edges,
or by considering synchronous dynamics or other representations
such as Thomas modeling~\cite{BernotSemBRN}.
However, the range of the analysis can also be extended,
by searching instead the set of initial states
allowing to reach a given goal,
or extending the method to universal properties (like the $\mathsf{AF}$ operator).
Finally, the research of attractors in a more general fashion
(such as cyclic or complex attractors)
would be of major interest to fully understand the behavior of models.

### 2015-01-28 comparaison.tex
update intermediate version
update intermediate version

@@
 \cite{roccaasp} which was developed also on ASP but for transition states networks.
-Each method use a specific kind of representation\footnote{The transalation from PH to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
+Each method use a specific kind of representation\footnote{The transalation from PH from or to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
 logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
@@
   \hline
-  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
+  TTR & 12 &42 & $2^{19}$ & full state & \todo{XX} & Inconc & 0m1.151s &  0m1.001s & 0m1.90s \\
   \hline
-  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
+  ERBB & 20 &152 & $2^{70}$ & full state & \todo{XX}  & - &1m55.38s & 2m31.64s & 0m11.84s \\
   \hline
-  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
+  ERBB & 20 &152 & $2^{70}$ & partial state & 0m2.61s & 0m0.027s &1m54.96s & - & 0m5.02s \\
   \hline
-  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & out & 4m27.93s \\
+  TCR & 40 &156 & $2^{73}$ & full state & - & Inconc & out & out & 4m27.93s \\
   \hline
@@
 
-\todo{Diviser en 2 tableaux}
-
-\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau} \repEB{En fait pour la méthode de Rocca le script qu'il m'ont envoyé génére quelques erreurs lors de la compilation.
-pour libddd y'avait pas l'option (--fixpoint) ne génère rien comme résultat et GINsim répond très rapidement mais il faudrait calculer tout le graphe de transition avant.
-}
+\todo{Diviser en 2 tableaux ou c'est bon??}
 
@@
 
-\item \textbf{Rocca's method} offers the possibility to model-check CTL properties
+\item \textbf{ASP-CTL} \todo{à revoir} offers the possibility to model-check CTL properties
 of state-transition networks.
@@
 Given our implementation, if the step \texttt{n} is reached,
-the computation stops and displays paths that are not valid executions
-to the goal.
-\todo{Y a-t-il moyen de les différencier de chemins corrects?}
+the computation stops and there is no path to be displayed because there is no path verifies the goals after \texttt{n} changes.

### 2015-01-28 fixPoint.tex
update intermediate version
update intermediate version

@@
 Before writing any piece of scripts on ASP to analyse the dynamic of the network,
-we first need to translat the concerned PH network in ASP  \footnote{You can find our new methods, translation from PH to ASP and the dynamic verification on line: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
+we first need to translat the concerned PH network in ASP  \footnote{You can find our new methods for the dynamic verification and also the translation from PH to ASP on line: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
 
@@
 \begin{lstlisting}
-sort(X) :- process(X,) %\label{ASPsort}
+sort(X) :- process(X,I) %\label{ASPsort}
 process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
@@
 \subsection{Search of fixed points}
-\todo{ Parler du point fixe et ses interrêts. \\
-Décrire la nouvelle méthode qui permet de déterminer les états stables d'un réseau }
 

### 2015-01-29 comparaison.tex
new update
update and new description of the method of Rocca  et al.

@@
   \hline
-   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-CTL} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
+   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-thomas} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
   \hline
-  TTR & 12 &42 & $2^{19}$ & full state & \todo{XX} & Inconc & 0m1.151s &  0m1.001s & 0m1.90s \\
+  TTR & 12 &42 & $2^{19}$ & full state & \todo{XX} & 0m.00.97s & 0m1.151s &  0m1.001s & 0m1.90s \\
   \hline
-  ERBB & 20 &152 & $2^{70}$ & full state & \todo{XX}  & - &1m55.38s & 2m31.64s & 0m11.84s \\
+  ERBB & 20 &152 & $2^{70}$ & full state & \todo{XX}  & out &1m55.38s & 2m31.64s & 0m11.84s \\
   \hline
@@
 \caption{Compared performances of Rocca et al.
-method denoted by \textsc{ASP-CTL}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
+method denoted by \textsc{ASP-thomas}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
 For each test, this table gives the considered model,
@@
 
-\todo{Diviser en 2 tableaux ou c'est bon??}
-
 \begin{itemize}
 
-\item \textbf{ASP-CTL} \todo{à revoir} offers the possibility to model-check CTL properties
-of state-transition networks.
-We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
-Therefore, using this method first requires an exponential computation to
-express the state-transition graph of the studied model into ASP.
-Indeed, for the tadpole tail resorption model comprising only 42 processes,
-this computation took 3 minutes, which was not included in the solving time
-reported in table~\ref{tab:reachability}.
+\item \textbf{ASP-THOMAS} offers the possibility to model-check CTL properties
+of Thomas networks.
+We note that in general, presenting such a network needs so much time because there is no converters for thomas networks to ASP.
+regardless of the complexity of the presentation of the graph it is clear that this approach gives a very quick result when compared with others (for the example TCR we didn't do the test).
+Indeed it also proves ASP is a good choice to run the graphs and check these properties.
 Afterwards, this method is able to check any kind of CTL formula
-(and not only the EF form that we focused on in this paper),
-but this checking may be time-consuming.
+(and not only the “$\mathsf{EF}$” form that we focused on in this paper) but the inconvinient thet they need to provide a number of steps greater than or equal to the number of steps after which the objectives are achieved.
+But this number is not always estimable.
 

### 2015-01-30 intro.tex
new update
proofread intro + definitions

@@
 For more than a decade, model-checking and SAT-solvers developments have been strongly connected \cite{biere1999symbolic}.
-Among different reasons, the main issue is related to the different problems with BDD-based symbolic approaches, that is:
+Among different reasons, the main ones are related to some of the drawbacks of Binary Decision Diagrams (BDD) -based symbolic approaches, that are:
 \begin{itemize}
@@
 In \cite{rocca2013inference}, the authors proposed an approach to perform LTL and CTL model-checking through ASP, with the goal to infer and learn (synchronous) Boolean networks.
-This method has been refined in the context of (asynchronous) Thomas' biological regulatory networks \cite{roccaasp}.
+This method has been refined in the context of (asynchronous) Thomas' biological regulatory networks in \cite{roccaasp}.
 
@@
 of either positive or negative mutual effects between the components. With the
-purpose of analyzing these systems, they are often modeled as graphs which make
+purpose of analyzing these systems, they are often modeled as graphs which makes
 it possible to determine the possible evolutions of all the interacting components of the system.
-Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels.
+Thus, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels.
 A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.
@@
 
-Our contribution is from the results that allowed to determine the stable states, we propose to evaluate the benefits of the Answer Set Programming (ASP) \cite{baral2003knowledge} to compute them.
-ASP has been proven efficient to tackle models with a large number of components and parameters.
-Our aim here is to assess its potential w.r.t.
-the computation of some dynamical properties of the PH model.
-In this paper, we show that ASP turns out to be effective for these enumerative searches which justifies its use.
-The benefit of our approach is that it makes possible to get the minimal paths to reach our goal(s) also we can verify if it is possible after a given number of steps.
+The particularity of our contribution relies in the use of Answer Set Programming
+(ASP)~\cite{baral2003knowledge}
+to compute the result of these searches.
+This declarative programming framework has been proven efficient
+to tackle models with a large number of components and parameters.
+Our aim here is to assess its potential w.r.t.\ the computation
+of some dynamical properties of PH models.
+In this paper, we show that ASP turns out to be effective
+for these enumerative searches, which justifies its use.
+In some cases, the computed dynamics can also be bounded by a give
+number of evolution steps.
 

### 2015-01-30 fixPoint.tex
proofread fixpoint + reachability

@@
-The study of fixed points (and, more generally, basins of attraction) provides an important understanding of the different behaviors of a Biological Regulatory Network BRN~\cite{wuensche1998genomic}.
+The study of fixed points (and, more generally, basins of attraction) provides an important understanding of the different behaviors of a Biological Regulatory Network (BRN)~\cite{wuensche1998genomic}.
 Indeed, a system will always eventually end in a basin of attraction,
@@
 
-In the following, we consider a Process Hitting $(\Sigma, \PHl, \PHh)$.
-It has been shown that a state $s \in L$ is a fixed point or a stable state of the Process Hitting if and only if $s$ has no next state~\cite{PMR10-TCSB} \ie if there is no playable action in this state.
+In the following, we consider a Process Hitting $\PH = (\Sigma, \PHl, \PHh)$.
+It has been shown that a state $s \in L$ is a fixed point (or stable state) of the Process Hitting model if and only if $s$ has no next state, \ie if there is no playable action in this state~\cite{PMR10-TCSB}.
 Therefore, a stable state in a Process Hitting network is a state so that
 every process does not hit or is not hit by another process in the same state.
-We note that given this result, processes involved in a self-hit (a process hit its self) cannot be part of a stable state.
+We note that given this result, processes involved in a self-hit (an action whose hitter and target processes are the same) cannot be part of a stable state.
 
 \subsection{Process Hitting translation in ASP}
-Before writing any piece of scripts on ASP to analyse the dynamic of the network,
-we first need to translat the concerned PH network in ASP  \footnote{You can find our new methods for the dynamic verification and also the translation from PH to ASP on line: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
+Before analyzing the dynamics of the network,
+we first need to translate the concerned PH network in ASP\footnote{All programs, including this translation and the methds describes in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
 
@@
 We will see in example \ref{ex:asp-ph} how a PH network is defined with these predicates.
-We note that all the ASP scripts are
 
@@
 \begin{lstlisting}
+process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
 sort(X) :- process(X,I) %\label{ASPsort}
-process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
 action("a",0,"b",1,0). action("a",1,"a",1,0). action("b",1,"z",0,2). %\label{actions1}
@@
 \end{lstlisting}
-Line \ref{ASPsort} enumerates every sort of network.
 In line \ref{ASPprocess} we create the list of processes corresponding to each sort,
@@
 \texttt{process("z", 0)}, \texttt{process("z", 1)}, \texttt{process("z", 2)}.
+Line \ref{ASPsort} enumerates every sort of the network from the previous information.
 Finally, all the actions of the network are defined in lines \ref{actions1} and \ref{actions2};
@@
 The first step consists of eliminating all processes involved in a self-hit;
-the others are recorded by the predicate \texttt{"shownProcess"}:
+the others are recorded by the predicate \texttt{shownProcess}:
 \begin{lstlisting}
@@
 \end{lstlisting}
-The previous line creates an many candidate answer sets as there are possible states
+The previous line creates an many potential answer sets as there are possible states
 to take into account.
-The last step consists of filtering each state that is not a fixed points,
-or, in other words, eliminating all candidate answer sets that do not
-represent a fixed point.
+Finally, the last step consists of filtering any state that is not a fixed point,
+or, in other words, eliminating all candidate answer sets in which an action could be played.
 For this, we use a constraint:
-indeed, any solution that satisfy the body of a constraint will be removed from the answer set.
-Regarding our problem, a state is eliminated if there exists an action linking two processes:
+any solution that satisfy the body of this constraint will be removed from the answer set.
+Regarding our problem, a state is eliminated if there exists an action linking two selected processes:
 \begin{lstlisting}
@@
 
-\begin{example}
+\begin{example}[Fixed points enumeration]
 The PH model of figure \pref{fig:ph} contains 3 sorts:
-$a$ and $b$ have 2 processes and $z$ has 3, therefore, the whole model has $2*2*3 = 12$ states (whether they can be reached or not from a given initial state).
+$a$ and $b$ have 2 processes and $z$ has 3; therefore, the whole model has $2*2*3 = 12$ states (whether they can be reached or not from a given initial state).
 We can check that this model contains only one fixed point: $\PHstate{b_0, z_2, a_0}$.
 Indeed, there is no action between each two of the processes contained is this state so no evolution is possible from this state. 
-For this example no other state verifies this property.
+In this example, no other state verifies this property.
 

### 2015-01-30 ASP.tex
new update

@@
 This paradigm allows the programmer to specify what the problem to be solved is, and not how to solve it.
-ASP programs are written in AnsProlog* \cite{sureshkumar2006ansprolog} (short for "Answer Set Programming in Logic" with superscript:~*).
+ASP programs are written in AnsProlog* \cite{sureshkumar2006ansprolog} (short for “Answer Set Programming in Logic”).
 These programs are composed of a set of \emph{facts} and a set of \emph{rules}, as defined below, from which other facts can be derived.
@@
 
-\subsubsection{The syntax and semantic of ASP}
+\subsubsection{Syntax and semantics of ASP}
 \label{sectionSyntaxeASP}
@@
 
+\item \textbf{The alphabet}
+
+According to \cite{baral2003knowledge}, the alphabet of ASP is composed of seven classes of symbols:
+\begin{itemize}
+  \item \emph{constants}, symbols of \emph{functions} and symbols of \emph{predicates} are composed of letters and start with a lowercase letter,
+  \item \emph{variables} follow the same rule but start with a capital letter,
+  \item \emph{connectors} that are “$\neg$”, “$\textbf{or}$”, “$\leftarrow$”, “$\textbf{not}$” and “$,$”,
+  \item \emph{punctuation} that can be “(”, “)” and “.”,
+  \item the special symbols “$\bot$” and “$\top$”.
+\end{itemize}
+
+The notion of predicate allows to qualify constants in this grammar.
+For example, the property “\textit{Tweety is a bird}”
+can be expressed by the predicate name \textit{bird}
+around the constant \textit{tweety} into the following literal:
+\textit{bird(tweety)}.
+Such literal can them be used as premise or conclusion in a rule as defined below.
+
 \item \textbf{The rules:}
-As previously mentioned, a rule has the following form:
-\texttt{
+A rule has the following form:
 \begin{equation} \label{eq1ASP}
@@
 \end{equation}
-}
-This rule is equivalent to:
+which is equivalent to:
 \begin{equation} \label{eq2ASP}
@@
 \end{equation}
-And this rule states:
+where $l_{i}$ are literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
+This rule states:
 \[
@@
     \end{array}
-\right \} \text{Then} ~ l_{0} ~ \text{is } \textbf{true} 
+\right \} \text{Then} ~ l_{0} ~ \text{is } \textbf{true}.
 \]
-with $l_{i}$ literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
-Therefore, "$head$" is also called the \emph{conclusion} of the rule,
-and "$body$" is its \emph{premise}.
+
+Therefore, “$head$” is also called the \emph{conclusion} of the rule,
+and “$body$” is its \emph{premise}.
+% An \emph{answer set} of a program (that is, a set of rules)
+% is a minimal set of literals that satisfies all the rules.
 
@@
 \begin{itemize}
-\item A  \textbf{ground} rule, is a rule  where all its literals are constants;
-\item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
-  It can be written without the central arrow ($\leftarrow$) or with a $body$ equals to $\top$, and $\top$  is the symbol of True.
+\item A \emph{ground} rule is a rule where all the literals are constants,
+  and thus contains no variables;
+
+\item A \emph{fact}, also called \emph{reality}, is a rule with an empty body.
+  It can be written without the central arrow ($\leftarrow$) or with a $body$ equal to $\top$:
 \begin{equation} 
@@
 
-\item A \textbf{constraint} is a rule where the head equals false ($\bot$).
+\item A \emph{constraint} is a rule where the head equals $\bot$.
   In this case, the head is often not represented and the constraint will be written generally as:
@@
 \end{equation}
-
-We say that a set of literals $X$ violates the constraint \eqref{eq6ASP} if $\{l_{1},...,l_{m}\} \subseteq X$ and $\{l_{m+1},...,l_{n}\} \nsubseteq X$.
-If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.
+A set of literals $X$ \emph{violates} the constraint \eqref{eq6ASP} if $\{l_{1},...,l_{m}\} \subseteq X$ and $\{l_{m+1},...,l_{n}\} \nsubseteq X$.
+Thus, $X$ cannot be an answer set of any program containing this constraint.
+% If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.
 
@@
 with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$).
-This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$.
-But that the number of this atoms $q_i$ should be between $l$ and $m$.
-That means that it is possible to find the same atom repeated many times at one answer set.\\
-In addition the literals $q_i$ can also appear in the body of the rule,
-and $l$ and $u$ are the lower and upper bounds of the number of these literals
-that can appear in the answer sets.
-%cardinality of the subsets of $ \{q_{1}, ... , q_{m}\}$ that satisfy the answer sets.
-Therefore, a cardinality constraint is satisfied in an answer set $X$, if the number of atoms of $\{q_{1}, ... , q_{m}\}$ belonging to $X$ is between $l$ and $u$,
-or, in other words, if:
+Such a cardinality means that the answer set $X$ contains
+at least $l$ and at most $u$ atoms in the set $\{q_{1}, ... , q_{m}\}$,
+or, in other words:
 \begin{equation}
@@
 \end{equation}
-where $\cap$ is the symbol of the intersection of sets
-and $|A|$ is the cardinality of the set $A$.
+where $\cap$ is the symbol of sets intersection
+and $|A|$ denotes the cardinality of set $A$.
+Thus, if such a cardinality is found in the head of a rule,
+then it directly constrains the resulting answer set.
+If it is found in the body of a rule,
+then it is simply true if the above condition is satisfied.
 In the following, if they are not explicitly given,
@@
 
-\item \textbf{The alphabet}
-
-According to \cite{baral2003knowledge}, the alphabet of ASP is composed of seven classes of symbols: Variables, start with a capital letter.
-Constants, symbols of functions, symbols of predicates, these 3 classes follow the same rule but they begin with lowercase letter.
-Connectors: $\neg$, or, $\leftarrow$, not, ‘,’.
-Symbols of punctuation :(‘(’, ‘)’, ‘.’) and finally the special symbols : $\bot$ and $\top$.
-
-The word predicate can be an innovation of the new grammar. Let's consider the following sentence to more explain it: 
-\begin{tabbing}
- \textit{tweety is a bird.} 
-\end{tabbing}
-"\textit{tweety}" is a constant and "\textit{is a bird}" is a predicat. This sentence can be noted in ASP by :
-\begin{tabbing}
- bird(tweety) $\leftarrow ~ \top$. \hspace{1cm} or \hspace{2cm} $bird(tweety) .$
-\end{tabbing} 
 \end{enumerate}
 
-\subsubsection{Modeling and solving of a problem with ASP: }
+\subsubsection{Modeling and solving of a problem with ASP}
 
-We can consider that the construction of models is one of the fundamental components of the scientific process.
+The construction of models is one of the fundamental components of the scientific process.
 It concerns all systems we seek to control.
-A model has two main features \cite{Glimpse}: In fact it is a simplification of a given system and it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties...
-This concept model can address another angle issues related to representation process.
-Logical ASP programs follow the strategy of “generate and test”.
-Indeed, the modeling process can be regarded as a special form of representation.
-First, the system should be presented  with facts.
-Second it have to be explain with rules.
-Third, it should be possible to generate all possibilities using cardinalities.
-Finally, a filter will be done thanking to the constraints and only the sets of predicates satisfying all these conditions will be displayed which are considered "answer sets".
+A model has two main features \cite{Glimpse}:
+it is a simplification of a given system
+and it allows an action on the system like resolving problems, searching for a path, identifying states and verifying properties...
+This concept of model can address another angle issues related to the representation process.
+the programs written in ASP follow the strategy of “generate and test”.
+Indeed, the modeling process can be regarded as a special form of representation:
+\begin{itemize}
+  \item First, the system should be presented with facts.
+  \item Second, properties of the model can be explained with rules.
+  \item Third, all candidate answers can be generated, usually using cardinalities.
+  \item Finally, constraints allow to filter the answers and only the sets of predicates satisfying all these conditions will be returned which and considered as \emph{answer sets}.
+\end{itemize}
 
-We denote that the solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
+Therefore, in practice, the solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
 that usually work together:
 first the grounder is used to remove variables in order to achieve an equivalent but constant program,
-then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
+then the solver computes all answer sets for this stabilized program~\cite{Vladimir,AnsPrologAPE}.
 Amongst all available grounders for ASP, we can name
@@
 
-For the work presented in this paper, we worked with \texttt{Clingo} (verion 3) which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.
+For the work presented in this paper, we worked with \texttt{Clingo} (version 3) which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-30 PH.tex

@@
 stays indefinitely,
-and the existence of several of these states denotes a switch in the dynamics.
+and the existence of several of these states denotes a switch in the dynamics~\cite{wuensche1998genomic}.
 

### 2015-01-30 reachability.tex
update intermediate version
update intermediate version
new update
update and new description of the method of Rocca  et al.

@@
 Then we tackle the reachability question: are there scenarios in a given initial state
-that allow to reach a given goal (under the form of a set of processes)?
+that allow to reach a given goal?
 
 \subsection{Future states identification}
-Thanks to its actions, a PH network is considered like a dynamic model or a dynamic biological network.
-In fact starting from an initial state the network can envolve into several new states.
-Our objectif is to try to predict the next new states as well as to  see if other states are reachable and if it's the case to determine the path to reach them.
-Firstly, we focus on the evolutions of the model in a limited number of steps.
-So we define the predicate \texttt{time(0..n)} which sets the number of steps we want to play.
-The value of \texttt{n} can be arbitrarily chosen.
-For example, \texttt{time(0..10)} will be used to compute the 11 first steps,
+In the previous section, enumerating the fixed points did not require to
+encode the full dynamics of PH, but only a condition.
+In this section, we thus implement the dynamics of the PH into ASP,
+which will then permit to search for the paths allowing to reach
+
+Firstly, we focus on the evolutions of models in a limited number of steps.
+We therefore define the predicate \texttt{time(0..n)} which sets the number of steps we want to play.
+The value of \texttt{n} can be arbitrarily chosen;
+for example, \texttt{time(0..10)} will be used to compute the 11 first steps,
 including the initial state.
@@
 identifying the future states requires to first identify the playable actions for each state.
-We recall that an action is playable in a state when both its hitter and target are active in this state (see Definition \ref{def:playableAction}).
+We remind that an action is playable in a state when both its hitter and target are active in this state (see Definition \ref{def:playableAction}).
 Therefore, we define an ASP predicate \texttt{playableAction(A,I,B,J,K,T)} that is true
@@
 We thus represent the change of the active process of a sort
-by the predicate \texttt{"activeFromTo(B,J,K,T)"}
+by the predicate \texttt{activeFromTo(B,J,K,T)}
 which means that in sort \texttt{B}, the active process changes from index $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
 The cardinality rule of line \ref{e2} then
-created as many answer sets as there are possible evolutions from the current step,
+creates as many answer sets as there are possible evolutions from the current step,
 thus recreating all possible branchings in the possible evolutions of the model.
@@
 two steps, by using the constraint of line \ref{e3}.
-Thus, the remaining scenarios of the answer set all strictly follow
+Thus, the remaining scenarios of the answer sets all strictly follow
 the asynchronous dynamics of the PH.
@@
           J!=K, time(T). %\label{e2}
-:- 2{ activeFromTo(B,J,K,T)}, time(T). %\label{e3}
+:- 2 {activeFromTo(B,J,K,T)}, time(T). %\label{e3}
 \end{lstlisting}
@@
 \subsection{Reachability verification}
-In this section, we focus on the reachability of set of processes which corresponds to the reachability question (see definition \ref{def:reachability}):
+In this section, we focus on the reachability of a set of processes which corresponds to the reachability property (see definition \ref{def:reachability}):
 “Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
@@
 as defined in the rule of line \ref{c2}.
-Else the answer will be eliminated by the constraint \ref{c2-a} which verifies if all processes of the goal are satisfied.
-\todo{Détailler la contraite 21 ou l'enlever} 
+Else the answer will be eliminated by a constraint (not detailed here) which verifies if all processes of the goal are satisfied.
 \begin{lstlisting}
 satisfiable(F, T) :-  goal(F), instate(F, T). %\label{c2}
-:- X={ satisfaible(F, T+1) : goal(F) }, getNbreGoals(Y), X=Y, time(T). %\label{c2-a}
 \end{lstlisting}
@@
 It is a main disadvantage because a search in $N$ steps will find no solution
-if the shortest path to the goal requires $K$ steps with $K > N$. It also may lengthen the resolution if the shortest path requires $n$ steps with $n << N$.
-One solution is to use an incremental computation mode,
-which is especially tackled by the incremental solver \textsc{iclingo}~\cite{gebser2008user}. In fact \textsc{iclingo} further includes the following guidelines:
+if the shortest path to the goal requires $K$ steps with $K > N$.
+It may also needlessly lengthen the resolution if the shortest path requires $n$ steps with $n << N$.
+One solution is then to use an incremental computation mode,
+which is especially tackled by the incremental solver \textsc{iClingo}~\cite{gebser2008user}.
+The syntax of \textsc{iClingo} separates the program in 3 parts.
+The \texttt{\#base} part contains only non-incremental elements
+and is thus used to declare general rules
+that do not depend on the time step (such as the data related to the model).
+The body iteration is written in the
+\texttt{\#cumulative} and \texttt{\#volatile} parts
+which are computed at each incremental step,
+the first part comprising rules depending on the time step,
+and the second containing a constraint that stops the iteration when needed.
+The step number is no given by a variable but by a constant placeholder
+called “\texttt{t}” in the following.
 
-\begin{tabbing}
-\texttt{\#base.}  \\
-... (static code for definitions) ... \\
-\texttt{\#cumulative constant.}  \\
-... (code depending on constant: body iteration) ... \\
-\texttt{\#volatile constant.} \\
-... (code for testing the satisfiyability property : stop computing iteration) ...
-\end{tabbing}
-
-Via \texttt{"\#base."} the subsequent part of a logic program is declared as static, that is, it is processed only once at the beginning of an incremental computation.
-In contrast, \texttt{"\#cumulative constant."} and \texttt{"\#volatile constant."}  are used to declare a (symbolic) constant as a placeholder for incremental step numbers.
-In the following our constant is denoted by \texttt{"t"} indicate the number of a step in the evolution of the network.
-Thus, the obtained program is almost identical,
-except that step numbers \texttt{N}
-are replaced by constant placeholders \texttt{"t"}
-that will be incremented throughout the execution (body iteration).
+When using this new syntax, the obtained program is almost identical
+to what was presented before,
+except that step numbers \texttt{T}
+are replaced by the constant placeholder \texttt{t}.
 In each step \texttt{t}, the program computes the playable actions \texttt{playableAction(A,I,B,J,K,t)}, the possible bounces \texttt{activeFromTo(B,J,K,t-1)}
 and the new states \texttt{instate(activeProcess(A,I),t+1)}
-in the same way than previously.
-Regarding the part of local steps, that is, all rules, facts, and integrity constraints computed in one step are dismissed before the next incremental step,
-we use a special constraint (line \ref{c4})
-which states that the program should continue to the next step
-while it is not satisfiable,
-so that all answers that do not meet the goals are eliminated:
+in the \texttt{\#cumulative} part
+the same way than previously, but only for the current step.
+The solver then compares its current answer sets with
+the \texttt{t}-dependent constraint given in the \texttt{\#volatile} part.
+Regarding our implementation, this constraint is given in line \ref{c4}
+and simply states that all goals have to be met.
+If this constraint invalidates all current answer sets,
+the computation continues in the next iteration in order to reach a valid answer set.
+As soon as some answer sets meet the constraints,
+they are returned and the computation stops.
 \begin{lstlisting}

### 2015-01-30 PH.tex
justification of the PH

@@
 
+The PH was chosen for several reasons.
+First, it is a general framework that,
+although it was mainly used for biological networks,
+allows to represent any kind of dynamical model,
+and converters to several other representations are available (see section~\ref{comparison}).
+Although an efficient dynamical analysis already exists for this framework,
+based on an approximation of the dynamics,
+it is interesting to identify its limits
+and compare them to the approached we present later in this paper.
+Finally, the particular form of the actions in a PH model allow
+to easily represent them in ASP,
+with one fact per action, as described in the next section.
+Other representations may have required supplementary complexity;
+for instance, a labeling would be required
+if actions could be triggered by a variable number of processes.
+
 \subsection*{Dynamical properties}

### 2015-01-30 comparaison.tex
Laurent results
proofread comparison

@@
 We have chosen the following tools, that are detailed below: 
-\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
-\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method for CTL model-checking of Rocca's et al.
-\cite{roccaasp} which was developed also on ASP but for transition states networks.
-Each method use a specific kind of representation\footnote{The transalation from PH from or to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
-logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
-Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
-state transitions network for Rocca's et al. method
-and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models.
-It available online \url{http://loicpauleve.name/pint/} } as well as for our method.
+\textsc{GINsim}\footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
+\textsc{LibDDD}\footnote{\textsc{LibDDD} is a library for symbolic model-checking of CTL \& LTL properties: \url{http://www.lip6.fr/}}
+(Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
+\textsc{Pint}\footnote{\textsc{Pint} is a library that implements tools for the parsing, simulation, formal analysis, and translation of PH models: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
+and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
+which was developed also in ASP but for transition states networks.
+Each method uses a specific kind of representation\footnote{When available, we used the converters included into \textsc{Pint} for these translations.}:
+Thomas models (a particular kind of logical regulatory networks) for \textsc{GINsim},
+instantiable transition systems for \textsc{LibDDD},
+state transition networks for the method of Rocca \textit{et al.}
+and Process Hitting (PH) for \textsc{Pint} as well as for our method.
 
@@
 and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.
-
 These models were chosen to be of different sizes:
@@
 the original number of biological components, due to the use of
-“cooperative sorts”, which allow to model Booelan gates but do not necessarily
+“cooperative sorts”, which allow to model Boolean gates but do not necessarily
 have a biological meaning.
@@
 ensuring the conservation of the dynamical properties.
-All these translations are available with \textsc{PINT} 
 All results alongside with more detailed specifications of the models
@@
   \hline
-   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-thomas} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
+   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-Thomas} & \textsc{Pint} & \textsc{LibDDD} & \textsc{GINsim} & \textsc{ASP-PH} \\
   \hline
@@
 \caption{Compared performances of Rocca et al.
-method denoted by \textsc{ASP-thomas}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
+method denoted by \textsc{ASP-thomas}, \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
 For each test, this table gives the considered model,
@@
 
-\item \textbf{ASP-THOMAS} offers the possibility to model-check CTL properties
+\item \textbf{ASP-Thomas} offers the possibility to model-check CTL properties
 of Thomas networks.
-We note that in general, presenting such a network needs so much time because there is no converters for thomas networks to ASP.
-regardless of the complexity of the presentation of the graph it is clear that this approach gives a very quick result when compared with others (for the example TCR we didn't do the test).
-Indeed it also proves ASP is a good choice to run the graphs and check these properties.
-Afterwards, this method is able to check any kind of CTL formula
-(and not only the “$\mathsf{EF}$” form that we focused on in this paper) but the inconvinient thet they need to provide a number of steps greater than or equal to the number of steps after which the objectives are achieved.
-But this number is not always estimable.
+The translations to Thomas modeling in ASP was made by hand.
+It is clear that this approach gives a very quick result when compared with others
+(although we could not perform tests on the TCR example).
+Indeed it also proves that ASP is a good choice to run the dynamics of a model and check reachability properties.
+Furthermore, this method is able to check any kind of CTL formula
+(and not only the “$\mathsf{EF}$” form that we focused on in this paper)
+but it is required to provide a maximum number of steps
+for which the dynamics will be computed, which mays be difficult to estimate
+(see the discussion in section~\ref{limitations}).
 
@@
 of gene interaction networks.
-It allows to compute all reachable stable states instantly;
+It allows to compute all reachable stable states from a given initial state instantly;
 however, it is not possible to compute all stable states independently from the initial state.
-Regarding the reachability problem, \textsc{GINSim} only allows to check
-full states, because of its approach which consists in computing
-all the state-transition graph and then check if there exists a path between two given states.
+Regarding the reachability problem, \textsc{GINsim} only allows to check the reachability of
+full states, because of its approach consists in computing
+all the state-transition graph and then search for a path between the two given states.
 Therefore, it was not possible to perform reachability checks on partial states.
@@
 
-\item \textbf{\textsc{libddd}}
+\item \textbf{\textsc{LibDDD}}
 is a library for symbolic model-checking of CTL \& LTL properties.
@@
 which is then stored under the form of a binary decision diagram for a more efficient analysis.
-
-This computation explains why \textsc{libddd} takes more time to respond,
+This computation explains why \textsc{LibDDD} takes more time to respond,
 and does not answer in less than 12 minutes for the biggest example
 (which contains $2^{73}$ states).
-Finally, \textsc{libddd} is not able to compute the stable states of a network.
+Finally, \textsc{LibDDD} is not able to compute the stable states of a network.
 
-\item \textbf{PINT}
-is a library gathering tools and translators related to the PH.
-It should be noted that Pint contains the only reachability analysis
+\item \textbf{\textsc{Pint}}
+is a library gathering tools and converters related to the PH.
+It should be noted that \textsc{Pint} contains the only reachability analysis
 developed so far natively for the Process Hitting,
@@
 It consists in an approximation that avoids to compute the state-transition graph;
-It is thus ensured to be really efficient, which explains the fastest results,
-but at the cost of possibly terminating without being conclusive (although this is rare).
-However, it is not designed for goals consisting of a full state, which
-may trigger an exponential process to reach an answer,
-which explains the high computing times for some of them \ref{tab:reachability}
+t is thus ensured to be really efficient, which explains the fastest results,
+but at the cost of possibly terminating without being conclusive.
+However, it is not designed for goals consisting of many processes,
+which are more likely to trigger an inconclusive response,
+or an exponential research for conclusive sub-solutions.
+This explains the high computation times for some tests in table \ref{tab:reachability}.
 Moreover, in the case of a positive answer,
@@
 \subsection{Strengths and limitations of our method}
+\label{limitations}
 
 In the previous sections,
-we developed new methods to check dynamical properties,
+we developed new methods in ASP to check dynamical properties,
 namely identifying stable states and finding all possible paths to reach a given goal.
-Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
+Compared to some other methods describes above
+(\textsc{GINsim}, \textsc{LibDDD} and the method proposed by Rocca \textit{et al.}),
 our method is relatively faster and also permits to study larger networks
-(up to $2^{73}$ states in our tests) despite some inconclusive cases.
-Indeed, our ASP program is non-conclusive in the cases where the given goal is not
+(up to $2^{73}$ states in our tests).
+
+However, our ASP program mays still be non-conclusive
+in the cases where the given goal is not
 reachable and the model contains loops in its dynamics
@@
 maximum which will be eventually reached in the case of an endless loop.
-This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iclingo},
+This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iClingo},
 where \texttt{n} is the maximum number of steps.
@@
 %%
+
+Several values can be given to this parameter.
 For example, the total number of states is an obvious maximum,
@@
 or, with a similar reasoning, the total number of processes can be chosen.
-In this case, however, a termination with no solution cannot be considered as a formal
+In these cases, however, a termination with no solution cannot be considered as a formal
 negative answer, unless one can prove that the chosen value \texttt{n}

### 2015-01-30 ASP.tex
fixes in section ASP
style of the tools names

@@
   \item \emph{variables} follow the same rule but start with a capital letter,
-  \item \emph{connectors} that are “$\neg$”, “$\textbf{or}$”, “$\leftarrow$”, “$\textbf{not}$” and “$,$”,
+  \item \emph{connectors} that are “$\leftarrow$”, “$\textbf{not}$” and “$,$”,
   \item \emph{punctuation} that can be “(”, “)” and “.”,
@@
 \left.
-    \begin{array}{ll}
-        \text{If} ~ l_{1},...,l_{m}  ~ \text{are \textbf{true}}\\
-        \text{and} ~ l_{m+1},...,l_{n} ~ \text{are \textbf{false}}
+    \begin{array}{rcl}
+        \text{If } & l_{1},...,l_{m}  & \text{ \textbf{can} be proved true}\\
+        \text{and } & l_{m+1},...,l_{n} & \text{ \textbf{cannot} be proved true}
     \end{array}
@@
 Amongst all available grounders for ASP, we can name
-\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
+\textsc{Gringo}, \textsc{DLV} and \textsc{LParse},
 and the solvers include
-\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...
+\textsc{SModels}, \textsc{DLV}, \textsc{CModels}, \textsc{Clasp}...
 
-For the work presented in this paper, we worked with \texttt{Clingo} (version 3) which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.
+For the work presented in this paper, we worked with \textsc{Clingo} (version 3) which is a combination of the grounder \textsc{Gringo} and the solver \textsc{Clasp}.

### 2015-01-30 comparaison.tex
label of table 1

@@
   \hline
-  TTR & 12 &42 & $2^{19}$ & full state & \todo{XX} & 0m.00.97s & 0m1.151s &  0m1.001s & 0m1.90s \\
+  TTR & 12 &42 & $2^{19}$ & full state & 0m7.21s & 0m.00.97s & 0m1.151s &  0m1.001s & 0m1.90s \\
   \hline
-  ERBB & 20 &152 & $2^{70}$ & full state & \todo{XX}  & out &1m55.38s & 2m31.64s & 0m11.84s \\
+  ERBB & 20 &152 & $2^{70}$ & full state & 0m2.448s & out &1m55.38s & 2m31.64s & 0m11.84s \\
   \hline
@@
 \end{tabular}
-\caption{Compared performances of Rocca et al.
-method denoted by \textsc{ASP-thomas}, \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
-For each test, this table gives the considered model,
-the number of its processes and its states, the type of goal
+\caption{Compared performances of several methods to compute reachability analyses.
+The method of Rocca \textit{et al.} (denoted by \textsc{ASP-Thomas}), \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method \textsc{ASP-PH} presented in this paper.
+For each test, this table gives the name of the considered model,
+its number of sorts, processes and states, the type of goal
 (either a whole state or a sub-state)
 and the computation time of the different methods used for the tests,
-where “out” marks an execution taking too much time or memory
-and “-” indicates that is not possible to do the test.
+where “out” marks an execution taking too much time or memory,
+“-” indicates that is not possible to do the test,
+and “Inconc” states that the method terminates without a response.
 }

### 2015-02-02 comparaison.tex
fix table comparison
split table comparison

--- 2015-01-30 comparaison.tex
@@
 
-\begin{figure}[htp]
+\begin{table}[ht]
 \begin{center}
-\label{tab:reachability}
 \noindent%
-\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
+\begin{tabular}{|l|c||c|c|c||c|c|}
 \hline
-   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-Thomas} & \textsc{Pint} & \textsc{LibDDD} & \textsc{GINsim} & \textsc{ASP-PH} \\
+  \multicolumn{2}{|c||}{Models} & \multicolumn{3}{c||}{PH representation} & \multicolumn{2}{c|}{Fixed points enumeration} \\
 \hline
-  TTR & 12 &42 & $2^{19}$ & full state & \todo{XX} & 0m.00.97s & 0m1.151s &  0m1.001s & 0m1.90s \\
+  Name & Components & Sorts & Processes & States & computation time & Nbr of results \\
 \hline
-  ERBB & 20 &152 & $2^{70}$ & full state & \todo{XX}  & out &1m55.38s & 2m31.64s & 0m11.84s \\
 \hline
-  ERBB & 20 &152 & $2^{70}$ & partial state & 0m2.61s & 0m0.027s &1m54.96s & - & 0m5.02s \\
+  TTR  & 8  & 12 & 42  & $2^{19}$ & \todo{FP} & \todo{nbrFP} \\
 \hline
-  TCR & 40 &156 & $2^{73}$ & full state & - & Inconc & out & out & 4m27.93s \\
+  ERBB & 20 & 42 & 152 & $2^{70}$ & \todo{FP} & \todo{nbrFP} \\
 \hline
-   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
+  TCR  & 40 & 54 & 156 & $2^{73}$ & \todo{FP} & \todo{nbrFP} \\
 \hline
 \end{tabular}
-\caption{Compared performances of Rocca et al.
-method denoted by \textsc{ASP-thomas}, \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
-For each test, this table gives the considered model,
-the number of its processes and its states, the type of goal
-(either a whole state or a sub-state)
-and the computation time of the different methods used for the tests,
-where “out” marks an execution taking too much time or memory
-and “-” indicates that is not possible to do the test.
+\caption{\label{tab:models}%
+Description of the models used in our tests and results of our fixed point enumeration.
+Each model is referred to by its short name, where
+TTR stands for the tadpole tail resorption model~\cite{khalis2009smbionet},
+ERBB for the receptor-regulated G1/S transition of the same name~\cite{Samaga2009}
+and TCR for the T-cell receptor signaling network~\cite{Klamt06}.
+For each of them, this table gives the number of biological components
+in the original representation,
+and the number of sorts, the number of processes
+and the number of states in the PH model.
+Finally, the last column gives the computation time for the enumeration of all fixed points
+and the number of results returned.
 }
 \end{center}
-\end{figure}
+\end{table}
+
+\begin{table}[ht]
+\begin{center}
+\noindent%
+\begin{tabular}{|l|l|c||c|c|c|c|c|}
+\hline
+  \multicolumn{3}{|c||}{Experiments} & \multicolumn{5}{c|}{Results} \\
+\hline
+  & Model & Target & \textsc{ASP-Thomas} & \textsc{Pint} & \textsc{LibDDD} & \textsc{GINsim} & \textsc{ASP-PH} \\
+\hline
+\hline
+  \#1 & TTR & full state & 0m7.21s & 0m.00.97s & 0m1.151s &  0m1.001s & 0m1.90s \\
+\hline
+  \#2 & ERBB & full state & 0m2.448s & out &1m55.38s & 2m31.64s & 0m11.84s \\
+\hline
+  \#3 & ERBB & partial state & 0m2.61s & 0m0.027s &1m54.96s & - & 0m5.02s \\
+\hline
+  \#4 & TCR & full state & - & Inconc & out & out & 4m27.93s \\
+\hline
+  \#5 & TCR & partial state & - & 0m0.014s & out & - & 1m35.080s \\
+\hline
+\end{tabular}
+\caption{\label{tab:reachability}%
+Compared performances of several methods to compute reachability analyses.
+The method of Rocca \textit{et al.} (denoted by \textsc{ASP-Thomas}), \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method presented in this paper, called \textsc{ASP-PH}.
+For each test, this table gives the short name of the considered model,
+as given in table~\ref{tab:models},
+the type of goal (either a whole state or a sub-state)
+and the computation time of the different methods used for the tests,
+where “out” marks an execution taking too much time or memory,
+\mbox{“~-~”} indicates that is not possible to do the test,
+and “Inconc” states that the method terminates without a response.
+}
+\end{center}
+\end{table}
 

### 2015-02-03 comparaison.tex
add thanks to Laurent Trilling + refs to experiments + typos

@@
 and compare it to other existing approaches.
-All computations were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
+All computations except one were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
 
@@
 
-\item \textbf{ASP-Thomas} offers the possibility to model-check CTL properties
+\item \textbf{ASP-Thomas}\footnote{These tests have been performed on another computer. The authors wish to thank Laurent Trilling for his help.}
+offers the possibility to model-check CTL properties
 of Thomas networks.
@@
 It is clear that this approach gives a very quick result when compared with others
-(although we could not perform tests on the TCR example).
+\todo{although we could not perform tests on the TCR example}.
 Indeed it also proves that ASP is a good choice to run the dynamics of a model and check reachability properties.
@@
 Regarding the reachability problem, \textsc{GINsim} only allows to check the reachability of
-full states, because of its approach consists in computing
+full states, because its approach consists in computing
 all the state-transition graph and then search for a path between the two given states.
-Therefore, it was not possible to perform reachability checks on partial states.
+Therefore, it was not possible to perform reachability checks on partial states
+(experiments \#3 \& \#5).
 Small state-transition graphs can also be displayed by this tool.
@@
 This computation explains why \textsc{LibDDD} takes more time to respond,
-and does not answer in less than 12 minutes for the biggest example
-(which contains $2^{73}$ states).
+and gets out of memory in about 12 minutes for the biggest example
+which contains $2^{73}$ states
+(experiments \#4 \& \#5).
 Finally, \textsc{LibDDD} is not able to compute the stable states of a network.
@@
 It consists in an approximation that avoids to compute the state-transition graph;
-t is thus ensured to be really efficient, which explains the fastest results,
+it is thus ensured to be really efficient, which explains the fastest results,
 but at the cost of possibly terminating without being conclusive.
 However, it is not designed for goals consisting of many processes,
-which are more likely to trigger an inconclusive response,
-or an exponential research for conclusive sub-solutions.
+which are more likely to trigger an inconclusive response
+(such as for experiment \#4),
+or an exponential research in sub-solutions
+(such as for experiment \#2).
 This explains the high computation times for some tests in table \ref{tab:reachability}.

### 2015-02-04 comparaison.tex
Morgan’s comments

@@
 \subsection{Evaluation}
-To evaluate the effectiveness of our new approach,
+To assess the efficiency of our new approach,
 we position ourselves with respect to existing methods dealing with different biological models.
@@
 from small (12 components) to large (40 components).
-We note however that the PH models considered may contain more sorts than
+We note however that the considered PH models may contain more sorts than
 the original number of biological components, due to the use of
@@
 have a biological meaning.
-The different model representations required have been obtained by translations
+The different model representations that are required for performing these benchmarks have been obtained by translations
 from the PH
@@
 The methods and the results provided by each of them are detailed in the following.
-The overall results shows that our method is efficient in computing reachability
+The overall results show that our method is efficient in computing reachability
 from a given initial state;
-furthermore, it sometimes provides more informations that the other existing ones.
+furthermore, it sometimes provides more information than the other existing ones.
 
@@
 \hline
-  TTR  & 8  & 12 & 42  & $2^{19}$ & \todo{FP} & \todo{nbrFP} \\
+  TTR  & 8  & 12 & 42  & $2^{19}$ & 0.004s & 0 \\
 \hline
-  ERBB & 20 & 42 & 152 & $2^{70}$ & \todo{FP} & \todo{nbrFP} \\
+  ERBB & 20 & 42 & 152 & $2^{70}$ & 0.017s & 3 \\
 \hline
-  TCR  & 40 & 54 & 156 & $2^{73}$ & \todo{FP} & \todo{nbrFP} \\
+  TCR  & 40 & 54 & 156 & $2^{73}$ & 0.021s & 1 \\
 \hline
@@
 \hline
-  \#4 & TCR & full state & - & Inconc & out & out & 4m27.93s \\
+  \#4 & TCR & full state & - & Inconc & out & out & 6m27.93s \\
 \hline
@@
 \caption{\label{tab:reachability}%
-Compared performances of several methods to compute reachability analyses.
+Compared performances of several methods to compute reachability analyses:
 The method of Rocca \textit{et al.} (denoted by \textsc{ASP-Thomas}), \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method presented in this paper, called \textsc{ASP-PH}.
@@
 of Thomas networks.
-The translations to Thomas modeling in ASP was made by hand.
+The translations to Thomas modeling in ASP were made by hand.
 It is clear that this approach gives a very quick result when compared with others
@@
 
-However, our ASP program mays still be non-conclusive
+However, our ASP program may still be non-conclusive
 in the cases where the given goal is not

### 2015-02-04 cc.tex
affiliation + ack
Olivier remarques

@@
 In this paper, we gave a new method to compute some dynamical properties
-on Process Hitting models, a subclass of synchronous automata.
+on Process Hitting models, a subclass of asynchronous automata.
 The main interest of our method is the use of Answer Set Programming,
@@
 
-Our work could benefit from several extension.
+Our work could benefit from several extensions.
 Of course, the set of applicable models can be extended,
@@
 would be of major interest to fully understand the behavior of models.
+
+{\small
+\paragraph{Acknowledgement}
+The European Research Council has provided financial support
+under the European Community's Seventh Framework Programme (FP7/2007--2013)~/
+ERC grant agreement no.~259267.
+}

### 2015-02-04 comparaison.tex
Olivier remarques

@@
 
-For this comparative study, we focus biological network of different sizes:
+For this comparative study, we focus on biological network of different sizes:
 a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
@@
 Compared to some other methods describes above
-(\textsc{GINsim}, \textsc{LibDDD} and the method proposed by Rocca \textit{et al.}),
-our method is relatively faster and also permits to study larger networks
+(\textsc{GINsim}, \textsc{LibDDD} and the method proposed by Rocca \textit{et al.}) noted by \textsc{ASP-Thomas} our method is relatively faster and also permits to study larger networks
 (up to $2^{73}$ states in our tests).

### 2015-02-04 fixPoint.tex
Olivier remarques

@@
 \begin{example}[Fixed points enumeration]
-The PH model of figure \pref{fig:ph} contains 3 sorts:
+The PH model of \pref{fig:ph} contains 3 sorts:
 $a$ and $b$ have 2 processes and $z$ has 3; therefore, the whole model has $2*2*3 = 12$ states (whether they can be reached or not from a given initial state).
 We can check that this model contains only one fixed point: $\PHstate{b_0, z_2, a_0}$.
-Indeed, there is no action between each two of the processes contained is this state so no evolution is possible from this state. 
+Indeed, there is no action between each two of the processes contained is this state so no execution is possible from this state. 
 In this example, no other state verifies this property.

### 2015-02-04 intro.tex
Olivier remarques

@@
 A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.
-This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using approached methods computing approximations of the dynamics~\cite{PMR12-MSCS}.
+This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using abstract methods computing approximations of the dynamics~\cite{PMR12-MSCS}.
 Furthermore, PH is suitable, according to the precision of the available information, to model BRNs with different levels of abstraction by capturing the most general dynamics.
@@
 and the Process Hitting framework.
-We then use ASP in section \ref{fixpoint} to tackle the enumeration
+Then we use ASP in section \ref{fixpoint} to tackle the enumeration
 of all fixed points in a model,

### 2015-02-04 reachability.tex
Olivier remarques

@@
 
-In this section, firstly present how to determine the possible evolution of a PH model after a finite number of steps with an ASP program.
-Then we tackle the reachability question: are there scenarios in a given initial state
+In this section, we present at first how to determine the possible behaviour in a PH model after a finite number of steps with an ASP program.
+Then we tackle the reachability question: are there scenarios from a given initial state
 that allow to reach a given goal?

### 2015-02-05 comparaison.tex
n influence
update

@@
 We have chosen the following tools, that are detailed below: 
-\textsc{GINsim}\footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
-\textsc{LibDDD}\footnote{\textsc{LibDDD} is a library for symbolic model-checking of CTL \& LTL properties: \url{http://www.lip6.fr/}}
+\textsc{GINsim}\footnote{\textsc{GINsim} \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
+\textsc{LibDDD}\footnote{\textsc{LibDDD} \url{http://move.lip6.fr/software/DDD/}}
 (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
-\textsc{Pint}\footnote{\textsc{Pint} is a library that implements tools for the parsing, simulation, formal analysis, and translation of PH models: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
+\textsc{Pint}\footnote{\textsc{Pint} \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
 and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
@@
 The translations to Thomas modeling in ASP were made by hand.
-It is clear that this approach gives a very quick result when compared with others
-\todo{although we could not perform tests on the TCR example}.
+It is clear that this approach gives a very quick result when compared with others although we could not perform tests on the TCR example.
 Indeed it also proves that ASP is a good choice to run the dynamics of a model and check reachability properties.
@@
 Compared to some other methods describes above
-(\textsc{GINsim}, \textsc{LibDDD} and the method proposed by Rocca \textit{et al.}) noted by \textsc{ASP-Thomas} our method is relatively faster and also permits to study larger networks
+(\textsc{GINsim}, \textsc{LibDDD} and the method proposed by Rocca \textit{et al.} noted by \textsc{ASP-Thomas}) our method is relatively faster and also permits to study larger networks
 (up to $2^{73}$ states in our tests).
@@
 the computation stops and there is no path to be displayed because there is no path verifies the goals after \texttt{n} changes.
+Indeed this permits to get a final result, which is \texttt{unsatisfable} property after \texttt{n} steps so with this way we eliminate the unlimitaded exection time.

### 2015-02-09 PH.tex
Morgan’s comments
submitted version

@@
 Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 5 \emph{actions} that model the dynamic behavior are depicted by pairs of arrows in solid and dotted lines.
-The grayed processes stand for the possible initial state: $\PHstate{a_1, v_0, z_1}$.
+The grayed processes stand for the possible initial state: $\PHstate{a_1, b_0, z_1}$.
 }
@@
   (or equivalently \emph{stable state})
-  if and only if is has no successors.
+  if and only if it has no successors.
   In other words, $s$ is a fixed point if an only if no action is playable in this state:

### 2015-02-09 comparaison.tex
submitted version

@@
 and compare it to other existing approaches.
-All computations except one were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
+All computations except one ( \textsc{ASP-Thomas} ) were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
 
@@
 and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
-which was developed also in ASP but for transition states networks.
+which was developed also in ASP but for states transitions networks.
 Each method uses a specific kind of representation\footnote{When available, we used the converters included into \textsc{Pint} for these translations.}:
@@
 \hline
-  \#1 & TTR & full state & 0m7.21s & 0m.00.97s & 0m1.151s &  0m1.001s & 0m1.90s \\
+  \#1 & TTR & full state & 0m7.21s & 0m0.97s & 0m1.15s &  0m2.05s & 0m1.90s \\
 \hline
-  \#2 & ERBB & full state & 0m2.448s & out &1m55.38s & 2m31.64s & 0m11.84s \\
+  \#2 & ERBB & full state & 0m2.45s & out &1m55.38s & 2m31.64s & 0m11.84s \\
 \hline
-  \#3 & ERBB & partial state & 0m2.61s & 0m0.027s &1m54.96s & - & 0m5.02s \\
+  \#3 & ERBB & partial state & 0m2.61s & 0m0.03s &1m54.96s & - & 0m5.02s \\
 \hline
-  \#4 & TCR & full state & - & Inconc & out & out & 6m27.93s \\
+  \#4 & TCR & full state & 0m7.61s& Inconc & out & out & 6m27.93s \\
 \hline
-  \#5 & TCR & partial state & - & 0m0.014s & out & - & 1m35.080s \\
+  \#5 & TCR & partial state & 0m2.12s & 0m0.02s & out & - & 1m35.08s \\
 \hline
@@
 
-\item \textbf{ASP-Thomas}\footnote{These tests have been performed on another computer. The authors wish to thank Laurent Trilling for his help.}
-offers the possibility to model-check CTL properties
-of Thomas networks.
-The translations to Thomas modeling in ASP were made by hand.
-It is clear that this approach gives a very quick result when compared with others although we could not perform tests on the TCR example.
-Indeed it also proves that ASP is a good choice to run the dynamics of a model and check reachability properties.
+\item \textbf{ASP-Thomas}\footnote{These tests have been performed on another computer: dual core , 2.13 GHZ with 1.8 GB RAM. The authors wish to thank Laurent Trilling for his help.}
+offers the possibility to model-check CTL properties of Thomas networks. 
+There is no automatic way that allows the modelisation of thomas networks in ASP.
+That is why it is necessary to study the entire network than to present it in ASP.
+But this modelisation takes so much time to be done regarding the complex representation of thomas networks.
+Comparing with our method, we use the PH model which its actions are more easy to be presented in ASP, one fact per action.
+In addition, the method "ASP-Thomas" requires to provide a maximum number of steps
+for which the dynamics will be computed, which mays be difficult to be predicted.
+However it is clear that this approach gives a very quick result when compared with others.
+Indeed it also show that ASP is a good choice to run the dynamics of a model and check reachability properties. 
 Furthermore, this method is able to check any kind of CTL formula
-(and not only the “$\mathsf{EF}$” form that we focused on in this paper)
-but it is required to provide a maximum number of steps
-for which the dynamics will be computed, which mays be difficult to estimate
+(and not only the “$\mathsf{EF}$” form that we focused on in this paper).
 (see the discussion in section~\ref{limitations}).
@@
 Moreover, in the case of a positive answer,
-it currently does not return the execution the path achieving the desired reachability,
+it currently does not return the execution of the path achieving the desired reachability,
 but only outputs its conclusion.
@@
 Compared to some other methods describes above
-(\textsc{GINsim}, \textsc{LibDDD} and the method proposed by Rocca \textit{et al.} noted by \textsc{ASP-Thomas}) our method is relatively faster and also permits to study larger networks
+(\textsc{GINsim}, \textsc{LibDDD} and the method \textsc{ASP-Thomas}) our method is relatively faster and also permits to study larger networks
 (up to $2^{73}$ states in our tests).
-
+We study the networks which are modeled in Process Hitting.
+This new fomalisim for network modelisation is a restriction of synchronous automata so it allows to represent any kind of dynamical model.
+Moreover it is easy to implement the dynamics of the PH into ASP.
 However, our ASP program may still be non-conclusive
@@
 Given our implementation, if the step \texttt{n} is reached,
-the computation stops and there is no path to be displayed because there is no path verifies the goals after \texttt{n} changes.
-Indeed this permits to get a final result, which is \texttt{unsatisfable} property after \texttt{n} steps so with this way we eliminate the unlimitaded exection time.
+the computation stops and there is no path to be displayed because there is no path which verifies the goals after \texttt{n} changes.
+Indeed this permits to get a final result, which is \texttt{unsatisfable} property after \texttt{n} steps so with this way we eliminate the unlimitaded execution time.

### 2015-02-09 fixPoint.tex
update
submitted version

@@
 Before analyzing the dynamics of the network,
-we first need to translate the concerned PH network in ASP\footnote{All programs, including this translation and the methds describes in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
+we first need to translate the concerned PH network in ASP\footnote{All programs, including this translation and the methods describes in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
 
@@
 \begin{lstlisting}
-1 { selectProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
+1 { selecedtProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
 \end{lstlisting}
-The previous line creates an many potential answer sets as there are possible states
+The previous line creates as many potential answer sets as there are possible states
 to take into account.
@@
 \begin{lstlisting}
-:- hit(A,I,B,J), selectProcess(A, I), selectProcess(B, J), A != B. %\label{contraintFix}
+:- hit(A,I,B,J), selectedProcess(A, I), selectedProcess(B, J), A != B. %\label{contraintFix}
 \end{lstlisting}

### 2015-02-09 reachability.tex
submitted version

@@
 In this section, we thus implement the dynamics of the PH into ASP,
-which will then permit to search for the paths allowing to reach
+which will then permit to search for the paths allowing to reach the goals.
 
@@
 and the second containing a constraint that stops the iteration when needed.
-The step number is no given by a variable but by a constant placeholder
+The step number is not given by a variable but by a constant placeholder
 called “\texttt{t}” in the following.

### 2015-04-06 PH.tex
After CAV reviewers

@@
 \begin{example*}
-\pref{fig:ph} represents a PH $(\PHs,\PHl,\PHa)$ with three sorts
+\pref{fig:ph} represents a $\PH$ $(\PHs,\PHl,\PHa)$ with three sorts
 ($\PHs = \{a, b, c\}$) and:
 $\PHl_a = \{a_0, a_1\}$,
-$\PHl_b = \{b_0, b_1, b_2\}$,
-$\PHl_c = \{c_0, c_1\}$.
+$\PHl_b = \{b_0, b_1\}$,
+$\PHl_z = \{z_0, z_1, z_2\}$.
 \begin{figure}[ht]
@@
 \label{def:playableAction}
-Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
+Let $\PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ a state of $PH$.
 We say that the action $h = \PHfrappe{a_i}{b_j}{b_k} \in \PHa$

### 2015-04-06 comparaison.tex
After CAV reviewers

@@
 offers the possibility to model-check CTL properties of Thomas networks. 
-There is no automatic way that allows the modelisation of thomas networks in ASP.
+There is no automatic way that allows the modelisation of Thomas networks in ASP.
 That is why it is necessary to study the entire network than to present it in ASP.
-But this modelisation takes so much time to be done regarding the complex representation of thomas networks.
+But this modelisation takes so much time to be done regarding the complex representation of Thomas networks.
 Comparing with our method, we use the PH model which its actions are more easy to be presented in ASP, one fact per action.

### 2015-04-06 fixPoint.tex
After CAV reviewers

@@
 for example, the first predicate \texttt{action("a",0,"b",1,0)} represents the action
-$\PHfrappe{a_1}{b_1}{b_0}$.
+$\PHfrappe{a_0}{b_1}{b_0}$.
 \end{example}
@@
 \begin{lstlisting}
-1 { selecedtProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
+1 { selectedProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
 \end{lstlisting}

### 2015-04-06 intro.tex
submitted version
After CAV reviewers

@@
 
-Behind the use of SAT (respectively ASP) to perform model-checking, there is the intuitive idea that the existence of a trace satisfying a given property expressed in LTL \cite{biere1999symbolic} or CTL (first reduced to universal properties \cite{penczek2002bounded}, then extended to other properties) can be reduced to the satisfiability of an equivalent logic formula (respectively a logic program).
+Behind the use of SAT, respectively ASP, to perform model-checking, there is the intuitive idea that the existence of a trace satisfying a given property expressed in LTL \cite{biere1999symbolic} or CTL can be reduced to the satisfiability of an equivalent logic formula, respectively a logic program.
+But first it should be reduced to universal properties \cite{penczek2002bounded}, then extended to other properties.
 Although there have been some works in the field of unbounded model-checking \cite{mcmillan2002applying}, it generally requires to consider bounded traces, which gave birth to the concept of Bounded Model-Checking (BMC).
@@
 for these enumerative searches, which justifies its use.
-In some cases, the computed dynamics can also be bounded by a give
+In some cases, the computed dynamics can also be bounded by a given
 number of evolution steps.

### 2015-04-06 reachability.tex
After CAV reviewers

@@
 which means that in sort \texttt{B}, the active process changes from index $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
-The cardinality rule of line \ref{e2} then
-creates as many answer sets as there are possible evolutions from the current step,
-thus recreating all possible branchings in the possible evolutions of the model.
+The cardinality rule of line \ref{e2a} - \ref{e2}
+creates the set of the predicates as many as there are possible evolutions from the current step,
+thus recreating all possible branchings in the possible evolutions of the model so many answer sets.
 This allows to filter all scenarios where two actions have been played between
@@
 \begin{lstlisting}
-{activeFromTo(B,J,K,T)} :-  playableAction(A,I,B,J,K,T),
+{activeFromTo(B,J,K,T)} :-  playableAction(A,I,B,J,K,T), %\label{e2a}
           instate(activeProcess(A,I),T), instate(activeProcess(B,J),T),

### 2015-04-20 main.tex
changement pour ICLP-15

% TLP2esam.tex / sample pages for TLP
% v2.11, released 6-nov-2002

\documentclass{tlp}
%\usepackage{aopmath}

% Packages
\usepackage{latexsym}
\usepackage{epic,eepic}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage[dvips]{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{amssymb}

%Package CAV
\usepackage[english]{babel}
\usepackage{stmaryrd} % Maths (crochets doubles)
\usepackage{url}     % Mise en forme + liens pour URLs
%\usepackage{array}   % Tableaux évolués
\usepackage{setspace}

% Tunning
%------------------
\usepackage[12pt]{moresize}
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end if" text
%\setlength{\belowcaptionskip}{10.0pt}
%\setlength{\abovecaptionskip}{10.0pt}
%\setlength{\textfloatsep}{10.0pt}
%\setlength{\floatsep}{10.0pt}
% TUNNING
\setstretch{0.94}
%\fontsize{3.85mm}{3.85mm}\selectfont

\newcommand{\pf}[1]{\langle#1\rangle}
\newcommand{\noproof}{\hfill $qed$}
\newcommand{\ignore}[1]{}

\definecolor{gray50}{gray}{0.15}
\graphicspath{{pictures/}}

\newtheorem{definition}{Definition} % [section]
\newtheorem{example}{Example} % [section]
\newcommand{\pivot}[1]{\mathbin{\, {#1} \,}}
\newcommand{\Pivot}[1]{\mathbin{\; {#1} \;}}
\let\from=\leftarrow

%Commande perso
\newcommand{\ie}{i.e.,\ }
\newcommand{\eg}{e.g.,\ }
\newcommand{\resp}{resp.\ }

\usepackage{listings}
% Définition du langage ASP
\lstdefinelanguage{ASP}{%
literate={:-}{$\leftarrow$}1,
morekeywords={not},
numbers=left,
escapeinside={\%}{\^^M}
}
% Définition des styles de tous les listings du document
\lstset{
language=ASP,
basicstyle=\small\ttfamily,
columns=fullflexible,
keywordstyle=\bfseries,
firstnumber=last,
keepspaces=true
}
\renewcommand{\thelstnumber}{\the\value{lstnumber}}
%% fin définition

% Styles ASP inline
\newcommand{\ASPnot}{\textbf{not}~}

% les inputs files
\input{macros/macros}
\input{macros/macros-ph}
\input{macros/tikzstyles2.tex}
%\input{macros/macros-abstr}

\begin{document}
\bibliographystyle{acmtrans}

\long\def\comment#1{}

\title[Exhaustive analysis of biological regulatory networks using ASP]{Exhaustive dynamic analysis of reachability and stable states
of biological regulatory networks modeled in Process Hitting
using Answer Set Programming}

\author[E. Ben Abdallah, M. Folschette, O. Roux, M. Magnin]
{Emna Ben Abdallah\\
$^1$LUNAM Université, \'Ecole Centrale de Nantes, \\
 IRCCyN UMR CNRS 6597\\ (Institut de Recherche en Communications et Cybern\'etique de Nantes),\\
  1 rue de la Noë, 44321 Nantes, France.\\
\email{emna.ben-abdallah@irccyn.ec-nantes.fr},\\
%E-mail: 
\and Maxime Folschette$^{1,2}$\\
$^2$School of Electrical Engineering and Computer Science,\\
University of Kassel, Germany \\
\email{maxime.folschette@uni-kassel.de},\\
\and
Olivier Roux$^1$ \\
LUNAM Université, \'Ecole Centrale de Nantes, IRCCyN \\
\email{olivier.roux@irccyn.ec-nantes.fr},\\
\and Morgan Magnin$^{1,3}$ \\
$^3$National Institute of Informatics, \\
2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.\\
\email{morgan.magnin@irccyn.ec-nantes.fr},\\
}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{\textbf{10} (3):}
\jdate{April 2015}
\setcounter{page}{1}
\pubyear{2015}

\maketitle

\label{firstpage}
%
\begin{abstract}
In this paper, we propose a new logical approach to perform model-checking on a restriction of synchronous automata networks, namely Process Hitting. The Process Hitting framework is notably suitable, but not limited, to model and analyze efficiently biological regulatory networks. 
The originality of our model-checking approach relies in the use of Answer Set Programming to consider exhaustive enumeration of all possible solutions of a property with respect to a given execution length.  Our work here focuses on state reachability properties (which are equivalent to the CTL “$\mathsf{EF}$” operator with paths limited to a given length) on the one hand, on the identification of fixed points on the other hand.
The merits of our methods is illustrated by applying them to
biological examples of various sizes
and comparing the results with some existing approaches.
It turns out that our approach succeeds in processing large models
with a high number of components and interactions.

	\keywords{Process Hitting, Answer Set Programming, stable states, fixed points, reachability}
\end{abstract}

\section{Introduction}
\input{parts/intro}

\section{Preliminary definitions}
\label{defs}
%\input{ASP}
\subsection{Process Hitting}
\input{parts/PH}

\section{Fixed point enumeration}
\label{fixpoint}
\input{parts/fixPoint}

\section{Dynamical analysis}
\label{dynamics}
\input{parts/reachability}

\section{Comparative performance analysis}
\label{comparison}
\input{parts/comparaison}

%\vspace{-1.0cm}

\section{Conclusion and future directions}
\label{ccl}
\input{parts/cc}

%\clearpage
\bibliographystyle{acmtrans}
\bibliography{biblio}

\clearpage
\appendix

%\ssmall
\input{parts/ASP}

\end{document}

### 2015-04-20 parts/PH.tex
changement pour ICLP-15

--- 2015-04-06 PH.tex
@@
-\pref{def:PH} introduces the Process Hitting~(PH)~\cite{PMR10-TCSB}
+\ref{def:PH} introduces the Process Hitting~(PH)~\cite{PMR10-TCSB}
 which allows to model a finite number of local levels,
@@
 
-\begin{example*}
-\pref{fig:ph} represents a $\PH$ $(\PHs,\PHl,\PHa)$ with three sorts
+\begin{example}
+\ref{fig:ph} represents a $\PH$ $(\PHs,\PHl,\PHa)$ with three sorts
 ($\PHs = \{a, b, c\}$) and:
@@
 \end{figure}
-\end{example*}
+\end{example}
 A state of the networks is a set of active processes containing a single process of each sort.

### 2015-04-20 parts/cc.tex
changement pour ICLP-15

--- 2015-02-04 cc.tex
@@
 on Process Hitting models, a subclass of asynchronous automata.
-The main interest of our method is the use of Answer Set Programming,
+The main interest of our method is the use of ASP,
 a declarative programming paradigm which benefits from powerful solvers.
@@
 for which the dynamics of the model from the given initial state is computed.
-We gave an implementation of these problems into Answer Set Programming,
+We gave an implementation of these problems into ASP,
 and applied them to several biological examples of various sizes, up to

### 2015-04-20 parts/comparaison.tex
changement pour ICLP-15

--- 2015-04-06 comparaison.tex
@@
 
-\begin{table}[ht]
-\begin{center}
-\noindent%
-\begin{tabular}{|l|c||c|c|c||c|c|}
-\hline
-  \multicolumn{2}{|c||}{Models} & \multicolumn{3}{c||}{PH representation} & \multicolumn{2}{c|}{Fixed points enumeration} \\
-\hline
-  Name & Components & Sorts & Processes & States & computation time & Nbr of results \\
-\hline
-\hline
-  TTR  & 8  & 12 & 42  & $2^{19}$ & 0.004s & 0 \\
-\hline
-  ERBB & 20 & 42 & 152 & $2^{70}$ & 0.017s & 3 \\
-\hline
-  TCR  & 40 & 54 & 156 & $2^{73}$ & 0.021s & 1 \\
-\hline
-\end{tabular}
+\begin{table}
+
+\includegraphics[width=\textwidth]{images/table1.pdf}
 \caption{\label{tab:models}%
@@
 Each model is referred to by its short name, where
-TTR stands for the tadpole tail resorption model~\cite{khalis2009smbionet},
-ERBB for the receptor-regulated G1/S transition of the same name~\cite{Samaga2009}
-and TCR for the T-cell receptor signaling network~\cite{Klamt06}.
+TTR stands for the tadpole tail resorption model% ~\cite{khalis2009smbionet},
+ERBB for the receptor-regulated G1/S transition of the same name% ~\cite{Samaga2009}
+and TCR for the T-cell receptor signaling network %\cite{Klamt06}.
 For each of them, this table gives the number of biological components
@@
 }
-\end{center}
 \end{table}
@@
 \begin{table}[ht]
-\begin{center}
-\noindent%
-\begin{tabular}{|l|l|c||c|c|c|c|c|}
-\hline
-  \multicolumn{3}{|c||}{Experiments} & \multicolumn{5}{c|}{Results} \\
-\hline
-  & Model & Target & \textsc{ASP-Thomas} & \textsc{Pint} & \textsc{LibDDD} & \textsc{GINsim} & \textsc{ASP-PH} \\
-\hline
-\hline
-  \#1 & TTR & full state & 0m7.21s & 0m0.97s & 0m1.15s &  0m2.05s & 0m1.90s \\
-\hline
-  \#2 & ERBB & full state & 0m2.45s & out &1m55.38s & 2m31.64s & 0m11.84s \\
-\hline
-  \#3 & ERBB & partial state & 0m2.61s & 0m0.03s &1m54.96s & - & 0m5.02s \\
-\hline
-  \#4 & TCR & full state & 0m7.61s& Inconc & out & out & 6m27.93s \\
-\hline
-  \#5 & TCR & partial state & 0m2.12s & 0m0.02s & out & - & 1m35.08s \\
-\hline
-\end{tabular}
+\includegraphics[width=\textwidth]{images/table2}
 \caption{\label{tab:reachability}%

### 2015-04-20 parts/fixPoint.tex
changement pour ICLP-15

--- 2015-04-06 fixPoint.tex
@@
 Before analyzing the dynamics of the network,
-we first need to translate the concerned PH network in ASP\footnote{All programs, including this translation and the methods describes in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
+we first need to translate the concerned PH network in ASP \footnote{All programs, including this translation and the methods describes in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
 
@@
 \begin{example}[Fixed points enumeration]
-The PH model of \pref{fig:ph} contains 3 sorts:
+The PH model of \ref{fig:ph} contains 3 sorts:
 $a$ and $b$ have 2 processes and $z$ has 3; therefore, the whole model has $2*2*3 = 12$ states (whether they can be reached or not from a given initial state).

### 2015-04-20 parts/intro.tex
changement pour ICLP-15

--- 2015-04-06 intro.tex
@@
 A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.
-This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using abstract methods computing approximations of the dynamics~\cite{PMR12-MSCS}.
+This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using abstract methods computing approximations of the dynamics \cite{PMR12-MSCS}.
 Furthermore, PH is suitable, according to the precision of the available information, to model BRNs with different levels of abstraction by capturing the most general dynamics.
@@
 fixed points (also called stable states)
-which are specific attractors with only one state~\cite{wuensche1998genomic}).
+which are specific attractors with only one state% \cite{wuensche1998genomic}).
 For this, we perform an exhaustive search of the possible states and then check which ones are fixed points.
@@
 The particularity of our contribution relies in the use of Answer Set Programming
-(ASP)~\cite{baral2003knowledge}
+(ASP)% \cite{baral2003knowledge}
 to compute the result of these searches.
@@
 
-Section \ref{defs} defines the paradigm of Answer Set Programming,
-and the Process Hitting framework.
-Then we use ASP in section \ref{fixpoint} to tackle the enumeration
-of all fixed points in a model,
-and in section \ref{dynamics} in order to compute the dynamics of a model
-and check reachability properties.
-Finally, a comparison of the methods developed in this paper is
-proposed in section \ref{comparison},
-and in section \ref{ccl} we conclude and talk about future work.
+%Section \ref{defs} defines the paradigm of Answer Set Programming, and the Process Hitting framework.
+Then we use ASP in section \ref{fixpoint} to tackle the enumeration of all fixed points in a model, and in section \ref{dynamics} in order to compute the dynamics of a model and check reachability properties.
+Finally, a comparison of the methods developed in this paper is proposed in section \ref{comparison}, and in section \ref{ccl} we conclude and talk about future work.

### 2015-04-27 parts/reachability.tex
syntaxe + pas de versionnage du PDF

--- 2015-04-06 reachability.tex
@@
 which means that in sort \texttt{B}, the active process changes from index $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
-The cardinality rule of line \ref{e2a} - \ref{e2}
-creates the set of the predicates as many as there are possible evolutions from the current step,
-thus recreating all possible branchings in the possible evolutions of the model so many answer sets.
+The cardinality rule of line \ref{e2a} -- \ref{e2}
+creates a set of as many predicates as there are possible evolutions from the current step,
+thus recreating all possible branchings in the possible evolutions of the model into as many potential answer sets.
 This allows to filter all scenarios where two actions have been played between

### 2015-04-28 main.tex
relecture complète

@@
 
+% Pretty ref
+\usepackage{prettyref}
+\newrefformat{def}{Definition~\ref{#1}}
+\newrefformat{fig}{Figure~\ref{#1}}
+\newrefformat{tab}{Table~\ref{#1}}
+%\newrefformat{pro}{Property~\ref{#1}}
+%\newrefformat{pps}{Proposition~\ref{#1}}
+%\newrefformat{lem}{Lemma~\ref{#1}}
+\newrefformat{th}{Theorem~\ref{#1}}
+\newrefformat{sec}{Section~\ref{#1}}
+%\newrefformat{subsec}{Subsect.~\ref{#1}}
+%\newrefformat{suppl}{Appendix~\ref{#1}}
+\newrefformat{ex}{Example~\ref{#1}}
+%\newrefformat{eq}{Eq.~\eqref{#1}}
+\def\pref{\prettyref}
+
 %Commande perso
@@
 \begin{document}
-\bibliographystyle{acmtrans}
+%\bibliographystyle{acmtrans}
 
@@
 The originality of our model-checking approach relies in the use of Answer Set Programming to consider exhaustive enumeration of all possible solutions of a property with respect to a given execution length.
-Our work here focuses on state reachability properties (which are equivalent to the CTL “$\mathsf{EF}$” operator with paths limited to a given length) on the one hand, on the identification of fixed points on the other hand.
+Our work here focuses on state reachability properties (which are equivalent to the CTL ``$\mathsf{EF}$'' operator with paths limited to a given length) on the one hand, on the identification of fixed points on the other hand.
 The merits of our methods is illustrated by applying them to
@@
 \section{Preliminary definitions}
-\label{defs}
-%\input{ASP}
+\label{sec:defs}
+%\input{parts/ASP}
 \subsection{Process Hitting}
@@
 \section{Fixed point enumeration}
-\label{fixpoint}
+\label{sec:fixpoint}
 \input{parts/fixPoint}
@@
 \section{Dynamical analysis}
-\label{dynamics}
+\label{sec:dynamics}
 \input{parts/reachability}
@@
 \section{Comparative performance analysis}
-\label{comparison}
+\label{sec:comparison}
 \input{parts/comparaison}
@@
 \section{Conclusion and future directions}
-\label{ccl}
+\label{sec:ccl}
 \input{parts/cc}

### 2015-04-28 parts/ASP.tex
changement pour ICLP-15
relecture complète

--- 2015-01-30 ASP.tex
@@
-In order to develop these methods we have used two main frameworks. The first is both a programming paradigm and a declarative programming language: the Answer Set Programming (ASP).
-The second framework is the Process Hitting (PH) a previously introduced formalism for the representation of biological regulatory networks.
+\section*{Answer Set Programming}
 
-\subsection{Answer Set Programming }
-
-\emph{Answer Set Programming} (ASP) is a declarative programming paradigm with semantics known as the \emph{semantics of answer sets}.
+We develop in this appendix the syntax and semantics of \emph{Answer Set Programming} (ASP) that is used in our work to solve declarative programs.
+% 
+% In order to develop these methods we have used two main frameworks. The first is both a programming paradigm and a declarative programming language: the Answer Set Programming (ASP).
+% The second framework is the Process Hitting (PH) a previously introduced formalism for the representation of biological regulatory networks.
+%
+%\subsection{Answer Set Programming }
+%
+ASP is a declarative programming paradigm with semantics known as the \emph{semantics of answer sets}.
 This paradigm allows the programmer to specify what the problem to be solved is, and not how to solve it.
-ASP programs are written in AnsProlog* \cite{sureshkumar2006ansprolog} (short for “Answer Set Programming in Logic”).
+ASP programs are written in AnsProlog* \cite{sureshkumar2006ansprolog} (short for ``Answer Set Programming in Logic'').
 These programs are composed of a set of \emph{facts} and a set of \emph{rules}, as defined below, from which other facts can be derived.
 A consistent set of facts that can be derived from a program using the rules is called \emph{answer set} for the program.
-The sets of possible responses to an AnsProlog* program are calculated with a program called a solver.
+The sets of possible responses to an AnsProlog* program are calculated with a program called \emph{solver}.
 
-\subsubsection{Syntax and semantics of ASP}
+\subsection*{Syntax and semantics of ASP}
 \label{sectionSyntaxeASP}
@@
   \item \emph{variables} follow the same rule but start with a capital letter,
-  \item \emph{connectors} that are “$\leftarrow$”, “$\textbf{not}$” and “$,$”,
-  \item \emph{punctuation} that can be “(”, “)” and “.”,
-  \item the special symbols “$\bot$” and “$\top$”.
+  \item \emph{connectors} that are ``$\leftarrow$'', ``$\textbf{not}$'' and ``$,$'',
+  \item \emph{punctuation} that can be ``('', ``)'' and ``.'',
+  \item the special symbols ``$\bot$'' and ``$\top$''.
 \end{itemize}
@@
 The notion of predicate allows to qualify constants in this grammar.
-For example, the property “\textit{Tweety is a bird}”
+For example, the property ``\textit{Tweety is a bird}''
 can be expressed by the predicate name \textit{bird}
@@
 \textit{bird(tweety)}.
-Such literal can them be used as premise or conclusion in a rule as defined below.
+Such literal can then be used as premise or conclusion in a rule as defined below.
 
-\item \textbf{The rules:}
+\item \textbf{The rules}
+
 A rule has the following form:
@@
     \begin{array}{rcl}
-        \text{If } & l_{1},...,l_{m}  & \text{ \textbf{can} be proved true}\\
-        \text{and } & l_{m+1},...,l_{n} & \text{ \textbf{cannot} be proved true}
+        \text{if } & l_{1},...,l_{m}  & \text{ \textbf{can} be proven true}\\
+        \text{and } & l_{m+1},...,l_{n} & \text{ \textbf{cannot} be proven true}
     \end{array}
-\right \} \text{Then} ~ l_{0} ~ \text{is } \textbf{true}.
+\right \} \text{ then} ~ l_{0} ~ \text{is } \textbf{true}.
 \]
 
-Therefore, “$head$” is also called the \emph{conclusion} of the rule,
-and “$body$” is its \emph{premise}.
+Therefore, ``$head$'' is also called the \emph{conclusion} of the rule,
+and ``$body$'' is its \emph{premise}.
 % An \emph{answer set} of a program (that is, a set of rules)
@@
 \end{equation}
-A set of literals $X$ \emph{violates} the constraint \eqref{eq6ASP} if $\{l_{1},...,l_{m}\} \subseteq X$ and $\{l_{m+1},...,l_{n}\} \nsubseteq X$.
-Thus, $X$ cannot be an answer set of any program containing this constraint.
+A set of literals $X$ \emph{violates} a constraint if
+its body is true, and thus cannot be an answer set
+(as $\bot$ cannot be part of an answer set).
+%$\{l_{1},...,l_{m}\} \subseteq X$ and $\{l_{m+1},...,l_{n}\} \nsubseteq X$.
+%Thus, $X$ cannot be an answer set of any program containing this constraint.
 % If a program $\pi$ contains such a constraint $r$, then $X$ is an answer set of $\pi$ if and only if $X$ is an \textit{answer set} of $\pi \setminus \{r\}$ and $X$ does not violate $r$.
@@
 
-\subsubsection{Modeling and solving of a problem with ASP}
+\subsection*{Modeling and solving of a problem with ASP}
 
@@
 This concept of model can address another angle issues related to the representation process.
-the programs written in ASP follow the strategy of “generate and test”.
+The programs written in ASP follow the strategy of ``generate and test''.
 Indeed, the modeling process can be regarded as a special form of representation:
@@
 
-For the work presented in this paper, we worked with \textsc{Clingo} (version 3) which is a combination of the grounder \textsc{Gringo} and the solver \textsc{Clasp}.
+For the work presented in this paper, we have worked with \textsc{Clingo} (version 3) which is a combination of the grounder \textsc{Gringo} and the solver \textsc{Clasp}.

### 2015-04-28 parts/PH.tex
relecture complète

--- 2015-04-06 PH.tex
@@
 An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
-“$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$”,
+``$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$'',
 where $a_i$, $b_j$, $b_k$ are processes of sorts $a$ and $b$,
@@
 
-\begin{example*}
-\pref{fig:ph} represents a $\PH$ $(\PHs,\PHl,\PHa)$ with three sorts
+\begin{example}
+\pref{fig:ph} represents a PH $(\PHs,\PHl,\PHa)$ with three sorts
 ($\PHs = \{a, b, c\}$) and:
@@
 A PH model example with three sorts: $a$, $b$ and $z$ ($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2).
-Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 5 \emph{actions} that model the dynamic behavior are depicted by pairs of arrows in solid and dotted lines.
-The grayed processes stand for the possible initial state: $\PHstate{a_1, b_0, z_1}$.
+Boxes represent the sorts (network components), circles represent the processes (component levels), and the 5 actions that model the dynamic behavior are depicted by pairs of arrows in solid and dotted lines.
+The grayed processes stand for the possible initial state $\PHstate{a_1, b_0, z_1}$.
 }
 \end{figure}
-\end{example*}
+\end{example}
 A state of the networks is a set of active processes containing a single process of each sort.
@@
 allows to represent any kind of dynamical model,
-and converters to several other representations are available (see section~\ref{comparison}).
+and converters to several other representations are available (see \pref{sec:comparison}).
 Although an efficient dynamical analysis already exists for this framework,
@@
 it is interesting to identify its limits
-and compare them to the approached we present later in this paper.
+and compare them to the approach we present later in this paper.
 Finally, the particular form of the actions in a PH model allow
@@
 The notion of \emph{fixed point}, also called \emph{stable state},
-is given in definition~\ref{def:fixpoint}.
+is given in \pref{def:fixpoint}.
 A fixed point is a state which has no successor.
@@
 A finer and more interesting dynamical property consists in
-the notion of \emph{reachability property}.
-Such a property, defined in definition~\ref{def:reachability},
+the notion of \emph{reachability}.
+Such a property, defined in \pref{def:reachability},
 states that starting from a given initial state, it is possible
@@
   we denote by $\mathcal{P}(s, A)$ the following \emph{reachability property}:
-  \[\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a} = a_i
+  \[\mathcal{P}(s, A) \equiv \exists \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a} = a_i
     \enspace.\]
@@
 The enumeration of all fixed points of a PH model will be tackled in
-section~\ref{fixpoint}
+\pref{sec:fixpoint}
 and the verification of a reachability property will be the subject of
-section~\ref{dynamics}.
+\pref{sec:dynamics}.
+The semantics of ASP is given in Appendix.

### 2015-04-28 parts/comparaison.tex
relecture complète

@@
 and compare it to other existing approaches.
-All computations except one ( \textsc{ASP-Thomas} ) were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
+All computations (except the one called \textsc{ASP-Thomas}) were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.
 
@@
 We have chosen the following tools, that are detailed below: 
-\textsc{GINsim}\footnote{\textsc{GINsim} \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
-\textsc{LibDDD}\footnote{\textsc{LibDDD} \url{http://move.lip6.fr/software/DDD/}}
+\textsc{GINsim}\footnote{\textsc{GINsim}: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
+\textsc{LibDDD}\footnote{\textsc{LibDDD}: \url{http://move.lip6.fr/software/DDD/}}
 (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
-\textsc{Pint}\footnote{\textsc{Pint} \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
+\textsc{Pint}\footnote{\textsc{Pint}: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
 and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
@@
 the original number of biological components, due to the use of
-“cooperative sorts”, which allow to model Boolean gates but do not necessarily
+``cooperative sorts'', which allow to model Boolean gates but do not necessarily
 have a biological meaning.
@@
 ensuring the conservation of the dynamical properties.
-All results alongside with more detailed specifications of the models
-are given in table \ref{tab:reachability}.
+The specification of the models and the results of our stable state enumeration
+are summed up in \pref{tab:models}.
+The time performance is roughly the same than the SAT implementation
+that comes with \textsc{Pint}.
+The results for several methods regarding reachability properties
+are summarized in \pref{tab:reachability}.
 The methods and the results provided by each of them are detailed in the following.
@@
 Each model is referred to by its short name, where
-TTR stands for the tadpole tail resorption model% ~\cite{khalis2009smbionet},
-ERBB for the receptor-regulated G1/S transition of the same name% ~\cite{Samaga2009}
-and TCR for the T-cell receptor signaling network %\cite{Klamt06}.
+TTR stands for the tadpole tail resorption model, % ~\cite{khalis2009smbionet},
+ERBB for the receptor-regulated G1/S transition of the same name % ~\cite{Samaga2009}
+and TCR for the T-cell receptor signaling network. %\cite{Klamt06}.
 For each of them, this table gives the number of biological components
@@
 Compared performances of several methods to compute reachability analyses:
-The method of Rocca \textit{et al.} (denoted by \textsc{ASP-Thomas}), \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method presented in this paper, called \textsc{ASP-PH}.
+the method of Rocca \textit{et al.} (denoted by \textsc{ASP-Thomas}), \textsc{Pint}, \textsc{LibDDD}, \textsc{GINsim} and our new method presented in this paper, called \textsc{ASP-PH}.
 For each test, this table gives the short name of the considered model,
@@
 and the computation time of the different methods used for the tests,
-where “out” marks an execution taking too much time or memory,
-\mbox{“~-~”} indicates that is not possible to do the test,
-and “Inconc” states that the method terminates without a response.
+where ``out'' marks an execution taking too much time or memory,
+\mbox{``~-~''} indicates that is not possible to do the test,
+and ``Inconc'' states that the method terminates without a response.
 }
@@
 
-\item \textbf{ASP-Thomas}\footnote{These tests have been performed on another computer: dual core , 2.13 GHZ with 1.8 GB RAM. The authors wish to thank Laurent Trilling for his help.}
+\item \textbf{\textsc{ASP-Thomas}}\footnote{These tests have been performed on another computer: dual core 2.13GHZ with 1.8GB RAM. The authors wish to thank Laurent Trilling for his help.}
 offers the possibility to model-check CTL properties of Thomas networks. 
@@
 But this modelisation takes so much time to be done regarding the complex representation of Thomas networks.
-Comparing with our method, we use the PH model which its actions are more easy to be presented in ASP, one fact per action.
-In addition, the method "ASP-Thomas" requires to provide a maximum number of steps
-for which the dynamics will be computed, which mays be difficult to be predicted.
-However it is clear that this approach gives a very quick result when compared with others.
-Indeed it also show that ASP is a good choice to run the dynamics of a model and check reachability properties. 
-Furthermore, this method is able to check any kind of CTL formula
-(and not only the “$\mathsf{EF}$” form that we focused on in this paper).
-(see the discussion in section~\ref{limitations}).
+Comparing with our method, we use the PH model whose actions are more easy to be represent in ASP, with one fact per action.
+In addition, the method ``\textsc{ASP-Thomas}'' requires to provide a maximum number of steps
+for which the dynamics will be computed, which may be difficult to be predicted.
+However it is clear that this approach terminates very quickly when compared with others.
+Indeed it also shows that ASP is a good choice to run the dynamics of a model and check reachability properties.
+Furthermore, this method is able to check any kind of CTL formula,
+and not only the ``$\mathsf{EF}$'' form that we focused on in this paper
+(see the discussion in \pref{sec:limitations}).
 
@@
 full states, because its approach consists in computing
-all the state-transition graph and then search for a path between the two given states.
+(part of) the state-transition graph and then searching for a path between the two given states.
 Therefore, it was not possible to perform reachability checks on partial states
@@
 (such as for experiment \#2).
-This explains the high computation times for some tests in table \ref{tab:reachability}.
+This explains the high computation times for some tests in \pref{tab:reachability}.
 Moreover, in the case of a positive answer,
@@
 \subsection{Strengths and limitations of our method}
-\label{limitations}
+\label{sec:limitations}
 
@@
 Compared to some other methods describes above
-(\textsc{GINsim}, \textsc{LibDDD} and the method \textsc{ASP-Thomas}) our method is relatively faster and also permits to study larger networks
+%and the method \textsc{ASP-Thomas}
+(\textsc{GINsim} and \textsc{LibDDD}) our method is relatively faster and also permits to study larger networks
 (up to $2^{73}$ states in our tests).
-We study the networks which are modeled in Process Hitting.
-This new fomalisim for network modelisation is a restriction of synchronous automata so it allows to represent any kind of dynamical model.
+We exclusively study networks modeled in Process Hitting.
+This new formalism for network modeling is a restriction of synchronous automata and thus allows to represent any kind of dynamical model.
 Moreover it is easy to implement the dynamics of the PH into ASP.
@@
 maximum which will be eventually reached in the case of an endless loop.
-This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iClingo},
+This is possible with the option ``\texttt{-{}-imax=n}'' of \textsc{iClingo},
 where \texttt{n} is the maximum number of steps.
@@
 
-Several values can be given to this parameter.
+Several values can be given to this parameter \texttt{n}.
 For example, the total number of states is an obvious maximum,
@@
 negative answer, unless one can prove that the chosen value \texttt{n}
-is bigger than the longest possible path in the state-transition graph.
-Given our implementation, if the step \texttt{n} is reached,
-the computation stops and there is no path to be displayed because there is no path which verifies the goals after \texttt{n} changes.
+is bigger than the longest possible path without loop in the state-transition graph.
+Given our implementation, if the step \texttt{n} gives no valid path,
+the computation stops with an \texttt{unsatisfiable} response.
+%and there is no path to be displayed because there is no path which verifies the goals after \texttt{n} changes.
 Indeed this permits to get a final result, which is \texttt{unsatisfable} property after \texttt{n} steps so with this way we eliminate the unlimitaded execution time.

### 2015-04-28 parts/fixPoint.tex
relecture complète

--- 2015-04-06 fixPoint.tex
@@
 and this may depend on biological switch or other complex phenomena.
-A fixed point is a state of the BRN in which it is not possible any more to have new changes;
+A fixed point is a state of the BRN in which it is not possible to have any new dynamic evolutions anymore;
 in other words, it is a basin of attraction that is composed of only one state.
@@
 In the following, we consider a Process Hitting $\PH = (\Sigma, \PHl, \PHh)$.
-It has been shown that a state $s \in L$ is a fixed point (or stable state) of the Process Hitting model if and only if $s$ has no next state, \ie if there is no playable action in this state~\cite{PMR10-TCSB}.
-Therefore, a stable state in a Process Hitting network is a state so that
-every process does not hit or is not hit by another process in the same state.
+%It has been shown that a
+A state $s \in L$ is a fixed point (or stable state) of the Process Hitting model if and only if it has no successor state, \ie if there is no playable action in $s$.
+%this state~\cite{PMR10-TCSB}.
+Therefore, it has been shown~\cite{PMR10-TCSB} that
+a stable state in a Process Hitting network is a state so that
+every active process does not hit or is not hit by another process in the same state.
 We note that given this result, processes involved in a self-hit (an action whose hitter and target processes are the same) cannot be part of a stable state.
@@
 Before analyzing the dynamics of the network,
-we first need to translate the concerned PH network in ASP\footnote{All programs, including this translation and the methods describes in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
-
+we first need to translate the concerned PH network into ASP \footnote{All programs, including this translation and the methods described in the following, are available online at: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.
 To do this we use the following self-describing predicates:
-\texttt{"sort"} to define sorts, \texttt{"process"} for the processes and \texttt{"action"} for the network actions.
-We will see in example \ref{ex:asp-ph} how a PH network is defined with these predicates.
+\texttt{sort} to define sorts, \texttt{process} for the processes and \texttt{action} for the network actions.
+\pref{ex:asp-ph} shows how a PH network is defined with these predicates.
 
@@
 \label{ex:asp-ph}
-The representation of the PH network of figure \ref{fig:ph} in ASP is the following:
+The representation of the PH network of \pref{fig:ph} in ASP is the following:
 \begin{lstlisting}
 process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
-sort(X) :- process(X,I) %\label{ASPsort}
+sort(X) :- process(X,_). %\label{ASPsort}
 action("a",0,"b",1,0). action("a",1,"a",1,0). action("b",1,"z",0,2). %\label{actions1}
@@
 this specific predicate will in fact expand into the three following predicates:
-\texttt{process("z", 0)}, \texttt{process("z", 1)}, \texttt{process("z", 2)}.
+\begin{lstlisting}[numbers=none]
+process("z", 0). process("z", 1). process("z", 2).
+\end{lstlisting}
 Line \ref{ASPsort} enumerates every sort of the network from the previous information.
+The underscore is a placeholder for any value.
 Finally, all the actions of the network are defined in lines \ref{actions1} and \ref{actions2};
@@
 \begin{lstlisting}
-hiddenProcess(A,I) :- action(A,I,B,J,K), A=B, process(A,I),process(B,J),
-                      process(B,K). %\label{hiddenProcess}
+hiddenProcess(A,I) :- action(A,I,B,J,K), A=B. %\label{hiddenProcess}
 shownProcess(A,I) :- not hiddenProcess(A,I), process(A,I). %\label{shownProcess}
@@
 \begin{lstlisting}
-1 { selectedProcess(A,I) : showProcess(A,I) } 1 :- sort(A).
+1 { selectedProcess(A,I) : shownProcess(A,I) } 1 :- sort(A).
 \end{lstlisting}
@@
 \begin{lstlisting}
-:- hit(A,I,B,J), selectedProcess(A, I), selectedProcess(B, J), A != B. %\label{contraintFix}
+:- action(A,I,B,J,_), selectedProcess(A,I),
+    selectedProcess(B,J), A!=B. %\label{contraintFix}
 \end{lstlisting}
@@
 If we execute the ASP program detailed below,
-alongside with the description of the PH model given in example~\ref{ex:asp-ph},
+alongside with the description of the PH model given in \pref{ex:asp-ph},
 we obtain one answer set, which matches the expected result:
 \begin{lstlisting}[numbers=none]
-Answer 1 : fixProcess(a, 0), fixProcess(b, 0), fixProcess(z, 2)
+Answer 1 : selectedProcess(a, 0), selectedProcess(b, 0),
+            selectedProcess(z, 2)
 \end{lstlisting}

### 2015-04-28 parts/intro.tex
relecture complète

@@
 Bounded model-checking has been the subject of numerous researches.
-On the one hand, it also suffers from some limitations.
-First, it leads to incomplete analysis: if the SAT solver proves that the problem given as input is unsatisfiable for a length $k$, it only proves there are no counterexamples of length $k$.
+On the one hand, it suffers from some limitations.
+Indeed, it leads to incomplete analysis: if the SAT solver proves that the problem given as input is unsatisfiable for a length $k$, it only proves there are no counterexamples of length $k$.
 Finding bounds on the length is difficult and worst case is exponential.
@@
 it possible to determine the possible evolutions of all the interacting components of the system.
-Thus, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels.
-A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system.
+Thus, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the ``Process Hitting'' (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels.
+A PH describes, in an atomic manner, the possible evolutions of a ``process'' (representing one component at one level) triggered by the hit of at most one other ``process'' in the system.
 This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using abstract methods computing approximations of the dynamics \cite{PMR12-MSCS}.
@@
 fixed points (also called stable states)
-which are specific attractors with only one state% \cite{wuensche1998genomic}).
+which are specific attractors with exactly one state. % \cite{wuensche1998genomic}).
 For this, we perform an exhaustive search of the possible states and then check which ones are fixed points.
@@
 The particularity of our contribution relies in the use of Answer Set Programming
-(ASP)% \cite{baral2003knowledge}
+(ASP) % \cite{baral2003knowledge}
 to compute the result of these searches.
@@
 
-%Section \ref{defs} defines the paradigm of Answer Set Programming, and the Process Hitting framework.
-Then we use ASP in section \ref{fixpoint} to tackle the enumeration of all fixed points in a model, and in section \ref{dynamics} in order to compute the dynamics of a model and check reachability properties.
-Finally, a comparison of the methods developed in this paper is proposed in section \ref{comparison}, and in section \ref{ccl} we conclude and talk about future work.
+\pref{sec:defs} defines
+%the paradigm of Answer Set Programming, and
+the Process Hitting framework.
+Then we use ASP in \pref{sec:fixpoint} to tackle the enumeration of all fixed points in a model,
+and in \pref{sec:dynamics} in order to compute the dynamics of a model and check reachability properties.
+Finally, a comparison of the methods developed in this paper is proposed in \pref{sec:comparison},
+and \pref{sec:ccl} allows to conclude and talk about future work.
+The definition and use of ASP is briefly explained in Appendix.

### 2015-04-28 parts/reachability.tex
relecture complète

--- 2015-04-06 reachability.tex
@@
 In this section, we thus implement the dynamics of the PH into ASP,
-which will then permit to search for the paths allowing to reach the goals.
+which will then permit to search for the paths allowing to reach some given goals.
 
@@
 The value of \texttt{n} can be arbitrarily chosen;
-for example, \texttt{time(0..10)} will be used to compute the 11 first steps,
+for example, \texttt{time(0..10)} will compute the 11 first steps,
 including the initial state.
-In order to specify such an initial state, we add several facts of the following form
-to list the processes included in this state:
+In order to specify such an initial state, we add several facts
+to make a list of all processes included in this state:
 \begin{lstlisting}
-init(activeProcess("a",0)).
+init(activeProcess("a",0)). init(activeProcess("b",0)).
+init(activeProcess("z",0)).
 \end{lstlisting}
-where \texttt{"a"} is the name of the sort and \texttt{"0"} the index of the active process.
-The dynamics of a network is described by its actions,
-identifying the future states requires to first identify the playable actions for each state.
-We remind that an action is playable in a state when both its hitter and target are active in this state (see Definition \ref{def:playableAction}).
+%where \texttt{"a"} is the name of the sort and \texttt{"0"} the index of the active process.
+The dynamics of a network is described by its actions;
+therefore, identifying the future states requires to first identify the playable actions for each state.
+We remind that an action is playable in a state when both its hitter and target are active in this state (see \pref{def:playableAction}).
 Therefore, we define an ASP predicate \texttt{playableAction(A,I,B,J,K,T)} that is true
@@
 by the predicate \texttt{activeFromTo(B,J,K,T)}
-which means that in sort \texttt{B}, the active process changes from index $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
-The cardinality rule of line \ref{e2a} - \ref{e2}
-creates the set of the predicates as many as there are possible evolutions from the current step,
-thus recreating all possible branchings in the possible evolutions of the model so many answer sets.
-This allows to filter all scenarios where two actions have been played between
+which means that in sort \texttt{B}, the active process can change from $\texttt{B}_\texttt{J}$ to $\texttt{B}_\texttt{K}$ between steps \texttt{T} and \texttt{T+1}.
+The cardinality rule of lines \ref{e2a} -- \ref{e2}
+creates a set of as many predicates as there are possible evolutions from the current step,
+thus reproducing all possible branchings in the possible evolutions of the model in the form of as many potential answer sets.
+This allows to remove all scenarios where two or more actions have been played between
 two steps, by using the constraint of line \ref{e3}.
-Thus, the remaining scenarios of the answer sets all strictly follow
+Thus, the remaining scenarios contained in the answer sets all strictly follow
 the asynchronous dynamics of the PH.
@@
 {activeFromTo(B,J,K,T)} :- playableAction(A,I,B,J,K,T), %\label{e2a}
-          instate(activeProcess(A,I),T), instate(activeProcess(B,J),T),
-          J!=K, time(T). %\label{e2}
+          instate(activeProcess(A,I),T),
+          instate(activeProcess(B,J),T), J!=K, time(T). %\label{e2}
 :- 2 {activeFromTo(B,J,K,T)}, time(T). %\label{e3}
@@
 resulting from the predicate \texttt{activeFromTo} (line \ref{e4})
-as well as all the unchanged processes that correspond to the other sorts (line \ref{e5} and \ref{e5-a}).
+as well as all the unchanged processes that correspond to the other sorts (lines \ref{e5} and \ref{e5-a}).
 
@@
 possible evolution in \texttt{n} time steps,
-and starting from a given initial state.
+starting from a given initial state.
 
 \subsection{Reachability verification}
-In this section, we focus on the reachability of a set of processes which corresponds to the reachability property (see definition \ref{def:reachability}):
-“Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
-We now want to adapt the code of the dynamics computation of previous section in order to resolve this reachability problem.
-For this, we first use a predicate to list the processes we want to reach, called \texttt{goal}, and we add as many rules of the following form as there are objective processes:
+In this section, we focus on the reachability of a set of processes which corresponds to the reachability property (see \pref{def:reachability}):
+``Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?''
+We now want to use the code of the dynamics computation of the previous section in order to solve this reachability problem.
+For this, we first use a predicate \texttt{goal} to list the processes we want to reach and we add as many rules of the following form as there are objective processes:
 \begin{lstlisting}
@@
 However, the limitation of the method above is that the user has to decide upstream
-the number of computed steps that should be sufficient.
+the number of computed steps.
+%that should be sufficient.
 It is a main disadvantage because a search in $N$ steps will find no solution
@@
 that do not depend on the time step (such as the data related to the model).
-The body iteration is written in the
-\texttt{\#cumulative} and \texttt{\#volatile} parts
-which are computed at each incremental step,
-the first part comprising rules depending on the time step,
-and the second containing a constraint that stops the iteration when needed.
+The body iteration is then written in the
+\texttt{\#cumulative} and \texttt{\#volatile} parts,
+which are computed at each incremental step.
+The first part comprises rules depending on the time step,
+and the second contains a constraint that stops the iteration when needed.
 The step number is not given by a variable but by a constant placeholder
-called “\texttt{t}” in the following.
+called ``\texttt{t}'' in the following.
 
@@
 are replaced by the constant placeholder \texttt{t}.
-In each step \texttt{t}, the program computes the playable actions \texttt{playableAction(A,I,B,J,K,t)}, the possible bounces \texttt{activeFromTo(B,J,K,t-1)}
-and the new states \texttt{instate(activeProcess(A,I),t+1)}
+In each step \texttt{t}, the program computes:
+\begin{itemize}
+  \item the playable actions \texttt{playableAction(A,I,B,J,K,t)},
+  \item the possible bounces \texttt{activeFromTo(B,J,K,t-1)}
+  \item the new states \texttt{instate(activeProcess(A,I),t+1)}
+\end{itemize}
 in the \texttt{\#cumulative} part

### 2015-04-29 parts/comparaison.tex
reformulations + versions logicielles

@@
 We have chosen the following tools, that are detailed below: 
-\textsc{GINsim}\footnote{\textsc{GINsim}: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
-\textsc{LibDDD}\footnote{\textsc{LibDDD}: \url{http://move.lip6.fr/software/DDD/}}
+\textsc{GINsim}\footnote{\textsc{GINsim} version 2.4 alpha: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
+\textsc{LibDDD}\footnote{\textsc{LibDDD} version 1.8: \url{http://move.lip6.fr/software/DDD/}}
 (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
-\textsc{Pint}\footnote{\textsc{Pint}: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
+\textsc{Pint}\footnote{\textsc{Pint} version 2015-02-11: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
 and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
@@
 offers the possibility to model-check CTL properties of Thomas networks. 
-There is no automatic way that allows the modelisation of Thomas networks in ASP.
-That is why it is necessary to study the entire network than to present it in ASP.
-But this modelisation takes so much time to be done regarding the complex representation of Thomas networks.
-Comparing with our method, we use the PH model whose actions are more easy to be represent in ASP, with one fact per action.
+There is however no automatic way that allows the modeling of Thomas networks in ASP, which currently has to be made by hand and requires labeling.
+As for our method, we use the PH whose actions are easily represented in ASP with one fact per action.
 In addition, the method ``\textsc{ASP-Thomas}'' requires to provide a maximum number of steps

### 2015-04-29 parts/comparaison.tex
modif de la version de pint

@@
 (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
-\textsc{Pint}\footnote{\textsc{Pint} version 2015-02-11: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
+\textsc{Pint}\footnote{\textsc{Pint} version 2015-11-14: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
 and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
@@
 in the cases where the given goal is not
-reachable and the model contains loops in its dynamics
-(which happens in almost every biological model).
+reachable and the model contains loops in its dynamics.
 In this case, the program will compute all infinite paths in these loops,
@@
 maximum which will be eventually reached in the case of an endless loop.
-This is possible with the option ``\texttt{-{}-imax=n}'' of \textsc{iClingo},
+This is possible with the option "\texttt{-{}-imax=n}'' of \textsc{iClingo},
 where \texttt{n} is the maximum number of steps.
